---
title: "Depression and suicidality as costly signals of need"
author: "Michael R. Gaffney, Kai Adams, Kristen L. Syme, and Edward H. Hagen"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
always_allow_html: true
output: 
  bookdown::html_document2:
    number_sections: false
bibliography: MTBib.bib
csl: apa-6th-edition.csl
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, warning = F, error = F, message = F)
# options(warn = 2)
library(equatiomatic)
library(knitr)
library(kableExtra)
library(glue)
library(ggplot2)
library(patchwork)
library(hagenutils)

tar_load(
    c(
    'd',
    'signalingdata2018b',
    'signaling2020',
    'pilot_results',
    'power_curve',
    'power_curve2',
    'models',
    'models_prereg',
    'effect_sizes',
    'effect_sizes2',
    'effectsPlots', 
    'effectsPlots_US',
    'effectsPlots_India',
    'ecdf_plot',
    'plot_T1Belief_T1Action',
    'plot_emotions',
    'plot_raw_beliefs',
    'plot_raw_actions',
    'plot_raw_belief_action2d',
    'med_prereg',
    'med1',
    'med1b',
    'med4',
    'med4b',
    'm1_boot',
    'm3_boot',
    'm1b_boot',
    'm3b_boot',
    'dfT1emotion_vignette',
    'dfT1emotion_signal',
    'dfT2emotion_signal'
    )
)

```

# Abstract

Major depression and suicide are generally viewed as pathological responses to stress that elicit aversive responses from others. Alternatively, depression and suicide attempts might be the result of adaptations for reliably signaling need and bargaining for increased support from reluctant others. To test this hypothesis, we examined responses to  emotional signals in a vignette study in which the existence of conflicts of interest and private information about one's need were emphasized in four scenarios involving claims of extreme need. Although the effect of signal type varied by vignette, costlier signals like depression and suicide attempts resulted in greater perceptions of need and increased likelihood of support compared to simple verbal requests and crying without further symptoms in our sample of 1,240 participants recruited from Amazon Mechanical Turk. Additionally, the effect of signal type on likelihood of support was largely mediated by the effect of signaling on believe in need as predicted. Our results suggest that depression and suicide attempts, apparent human universals, are a relatively successful way of signaling honest need and eliciting support, consistent with costly signaling explanations.

# Introduction

Durkheim’s pioneering study of suicide [@durkheim1897] helped launch sociology as a quantitative discipline. Sociological investigations of suicide flourished through much of the 20th century, but were supplanted by medical and public health approaches as the biomedical model of mental health took hold in the late 1970's and early 1980's [@wray2011; @syme2020]. The biomedical model tends to frame suicide as a problem with the individual, e.g., as one involving psychiatric disturbance, drug addiction, or stress sensitivity [@wray2011]. This theoretical approach has made little progress: Rates of suicide, which vary widely by country, have remained relatively flat in countries for which there are long-term data, such as the US, UK, Japan, and Sweden [@owidsuicide].

Early on, anthropologists also tackled the problem of suicide as manifested in the small traditional societies in which they conducted their fieldwork. Suicide, they found, was commonly a form of protest, revenge, and/or appeal [e.g., @Malinowski:1932; @niehaus2012]. Some ethnographers, such as @Hezel:1987, conceptualized suicide as a form of anger, whereas others, such as @Giddens:1964, conceptualized suicide as a form of social pressure. @Counts:1980 emphasized the powerlessness of suicide victims. Common to almost all theoretical and empirical investigations of suicide is a focus on *completed* suicides, i.e., suicide deaths. @SymeEtAl2016, in contrast, argued that because the vast majority of suicidal behavior does not result in death, the proper focus should be on suicide ideation and suicide attempts, with death as an unintended consequence. See Figure \@ref(fig:cdc).

```{r cdc, fig.width=8, fig.cap="US Suicide attempt and completion rates by age and sex (2001-2019)."}
library(hrafsuicidedata)
cdc2 %>% 
  na.omit() %>% 
  ggplot(aes(Age, `Crude Rate`, colour = Sex, linetype = Type)) + 
  geom_line() +
  hagenutils::scale_color_binary() +
  guides(
    colour = guide_legend(title=NULL),
    linetype = guide_legend(title=NULL, reverse = T)
  ) +
  labs(x = '\nAge (years)', y = 'Annual rate\n(per 100,000)') +
  theme_minimal(15) +
  theme(axis.title.y = element_text(angle = 0, hjust=0))
```

Within anthropology, Raymond Firth was likely the first to conceputalize suicidality as a gamble to improve one's circumstances in the here and now. Based on observations that suicide attempts often followed social conflict or loss, and varied substantially in their likelihood of death, he argued that a sizable subset of the suicide attempts among the Tikopia were not meant to end in death but instead were a means to elicit attention, status, and immediate reintegration into the community following negative events [@firth1936; @Firth:1961]. 

Suicidality is closely linked to depression. Suicidal ideation is one of nine diagnostic criteria for a major depressive episode [MDE, @dsm5], it is associated with depression across cultures [@haroz2017], and depression is a major risk factor for suicidal behavior [@hawton2013]. Compared to suicide, however, there are fewer anthropological investigations of depression [@kleinman1985] despite frequent reports of depression or depression-like behavior in a wide range of societies [@haroz2017; @liu2020] and the potential for cross-cultural research to help understand depressive states [@haroz2017].

Evolutionary theories of depression tend to either ignore suicidality, or take it as evidence that depression is dysfunctional. Wakefield, for instance, who has devoted much of his career to arguing that most cases of depression are simply severe but normal sadness and grief, nevertheless takes suicidality to indicate pathology for some cases of depression [e.g., @wakefield2017]. @nettle2004 argues that the costs of depression, including suicidality, outweigh any benefits, and therefore adaptationist models of depression are implausible. @gilbert2006 sees suicidality as an example of a dysregulated escape defense [for reviews of evolutionary theories of depression, see @].

and anthropological explanations for low mood have become more common. In general, these approaches see sadness and minor depression as ways to motivate adaptive behavior changes in response to negative outcomes in one's environment (cite Nesse and others), mainly differing in how they view major depression. While some approaches present most forms of major depression as dysfunctional (inflammation and some of Nesse), others have extended the logic on the benefits of low mood and reduced behavior to more severe instances of sadness. Of the possible benefits, the ability of depression to help solve fitness-relevant in times of social conflict is the most emphasized, with depression being suggested to help by reallocating resources with the body for problem solving (cite ARH), reducing the costs of interacting with higher status individuals (cite), or serving as a means to both signal one's need and bargain for increased support (cite). 


As with suicide, research on depression since the 1980's has been dominated by biomedical approaches of limited success [@patten2016; @ormel2019]. 


and there is still substantial uncertainty about why both are as common as they are (cite). Likely due to  the suffering they cause, the vast majority of these approaches have been built on the assumption that they are the result of internal dysfunction (cite examples). However, a subset of anthropologists and evolutionary theorists have challenged this idea, suggesting that despite the subjective suffering they cause, they may function as a means to limit negative fitness outcomes in times of substantial adversity (Firth + others).



## Connection to Adversity

Likely the most reliable finding regarding depression and suicidality is that most cases are closely tied to major negative life events [@Hammen2005; @Mazure1998; @devries2011; @ellsberg2008]. Compared to non-depressed individuals, those with depression tend to report at least twice as many instances of negative events than non-depressed individuals [@Mazure1998] and more negative events than those with schizophrenia and bipolar depression across a wide range of studies [@Paykel1994]. In longitudinal studies, the onset of depression tends to quickly follow the negative event [@KendlerEtAl1999] (search for others) or coincide with periods where conflict is likely to increase prior to it (e.g., depression starting before a divorce rather than after; @rosenstrom2017). 

This strong connection between negative life events and depression along with the adversity and distress that likely accompanies these events has led to the suggestion that this relationship is causal [@KendlerEtAl1999]. Support for this comes from the fact that the relationship between negative events and depression holds true when looking solely at events outside of one's control, showing that the connection is unlikely to be driven solely by depressed individuals creating situations where negative events are likely to be common [@Hammen2005; @KendlerEtAl1999]. In addition to this, twin studies have shown that one's history of negative events is still a strong predictor of major depression when controlling for genetic similarity and that part of the heritability of depression stems from the heritability of negative events, like divorce and family conflict [@KendlerBaker2007; @KendlerEtAl1999].

Of the types of negative life events associated with depression, those which involve social adversity tend to be more strongly associated with depression [@Hammen2005; @Mazure1998]. The loss of a loved one, marital problems, bullying, and the existence of abusive relationships are all commonly cited risk factors for depression [@KendlerEtAl1995; @KendlerEtAl1999; @klomek2007; @klomek2019], with  assault (both sexual and non-sexual), in particular, greatly increasing one's risk of depression [@KendlerEtAl1995; @KendlerEtAl1999]. Among the Tsimane, conflict in general has also been significantly associated with depression, particularly among non-kin [@StieglitzEtAl2015a]. Other notable risk factors like financial problems, job loss, or severe or prolonged illness, might seem to lack this social element. However, the ability for others to provide meaningful support to limit the harm caused by such events [@SugiyamaSugiyama2003] combined with the ubiquity of conflicts of interests mean that many types of problems will eventually involve social conflict even if they did not start that way [@Hagen2003)].

As depression is the leading cause of suicide compared to mental disorders [@who2001] (need to find something more recent), it should not be surprising that suicidality also often coincides with social adversity. Grief over the loss of a loved one [@firth1936: @Firth:1961; @devries2011], bullying [@alavi2017; @butler2019; @ford2017; @wang2020], and instances of social transgression [@Firth:1961; @Malinowski:1932; @SymeHagen2018; @lester1997] have all been reported to lead to suicidal behavior in both industrialized and traditional societies. Of the recorded instances of suicide attempts in the Human Relations Area Files (HRAF), 77.4% of attempts were found to be linked to times of social conflict, 88.7% were linked to fitness threats in general, and 84.9% were linked to situations in which attempters were powerless to resolve their situation alone [@SymeEtAl2016]. As with depression, assault is an particularly powerful risk factor for suicidal behavior based on multiple large-scale multi-country studies [@devries2011; @ellsberg2008] and more localized examinations. For example, both sexual and non-sexual assault were also associated with more self-harm and suicidal behaviors among both indigenous Sami and majority Norwegian adolescents, with those who attempted suicides also reporting more than twice the number of negative life events than those who did not [@reigstad2017]. 

## Depression and suicide attempts as both signals and bargaining tools 

One possible reason for the connection between depression, suicide, and social adversity is that some instances of depression and suicide attempts serve as means to elicit increased support from others [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], particularly in times when an individual cannot solve substantial fitness-relevant challenges on their own [@Hagen2003; @SymeEtAl2016]. According to the bargaining models of depression and suicide, the idea is that the small groups humans lived in for much of our evolutionary history created many situations in which certain forms of cooperation were unlikely to be easily replaced by others while also allowing for harmful or unfair social relationships to develop [@Hagen2003; @SymeEtAl2016]. In these instances, one's monopoly on one or more types of cooperation could be leveraged to change how one is treated if others valued an individual's cooperation sufficiently.

One way this approach expects depression and suicidal behavior to lead to beneficial treatment is through their ability to reliably demonstrate that one’s need is genuine by making one’s private information about their condition public [@Hagen2003]. This information is necessary due to the ease in which signals can be faked [@Zahavi1975; @mokkonen2016] and the costs of accepting a deceptive signal as reliable [@dawkins1978; @johnstone1993; @Hasson1994], both of which have favored adaptations to spot or ignore fake signals in many species [@dawkins1991]. 

One way for honest signals to evolve is through costly signaling [@Zahavi1975]. The evolution of costly signals is possible whenever the signal in question is comparatively cheaper to produce for those who have the underlying quality being compared to those who lack it [@grafen1990]. In these instances, individuals with the quality maximize their fitness at greater levels of investment in the signal compared to those who lack the quality or possess it to a lesser degree [@zahavi1977]. The end result of this is expected to be a signal that either tends to not be worth faking or is impossible to fake to a large degree due to selection against fakers constraining the physiology behind the signal [@zahavi1977; @gangestad2005].

Both depression and suicidal behavior fit this criterion as they are both relatively costly for those not facing adversity than those who are. In the case of depression, the largescale reduction in one's normal routine of fitness enhancing behavior is relatively less costly for those whose situations result in a limited ability to obtain fitness-relevant benefits, with depression likely adding no additional costs in instances of extreme deprivation [@Hagen2003]. However, when it comes to suicide, the bulk of the costs associated with suicide attempts are only felt in times that result in death or severe bodily harm. For this reason, the way others view and respond to suicide attempts is likely to be strongly influenced by the probability that the attempt would result in death or severe injury [@Rosenthal1993]. That given, the previously mentioned logic still applies with increased risk of death being relatively more costly for those able to advance their fitness prospects in their day-to-day lives.

Beyond resulting in the cost differential that allows for costly signaling in the first place, the costs of depression and suicidal behavior may be particularly important in times of conflict or when the level of support needed to address one’s situation is substantial. As previously mentioned, there are likely few forms of social adversity that do not come with conflict. For example, breakups in both romantic and cooperative relationships are often preceded by conflict [@amato2007], and surely involve conflict when one side wants reconciliation and the other does not. Likewise, the death of a loved one may also be a source of conflict as the survivor may turn to others for more support who might not want to invest more in the survivor [@Hagen2003]. Regardless of the source of conflict, greater conflict should correspond with greater conflicts of interests, something which would favor greater investment in manipulating others through fake signals. In these instances, less costly signals like verbal requests, facial sadness, and crying are expected to be less reliable as their low costs can easily be made up for with increased social support, thereby favoring the use of costlier signals in those experiencing need they cannot resolve on their own. 

Beyond signaling need to those invested in the well-being of the signaler, depression and suicidal behavior may also help an individual bargain for increased support [@Hagen2003]. This is possible due to the reduced investment in others that frequently accompanies depression or the threat to future cooperation that comes from a suicide. In these instances, depression and suicide attempts are suggested to function analogously to a labor strike, sending the message that one’s need is genuine and that if the target of the signal wanted to continue receiving benefits from the depressed, he or she would need to increase support towards the signaler.

### Determining the impact of cues versus signals

Distinguishing between the information provided by signals versus cues is a major challenge for signaling approaches to behavior. A signal differs from a cue in that it is intentionally sent by an organism to alter the behavior or thoughts of other individuals, while a cue is any piece of information an organism uses to alter its behavior or physiological state that was created without the intent to do so [@MaynardSmithHarper1995]. Regarding the signals of interest, suicide attempts are more likely to take the form of pure, external signals than other symptoms of depression. This is because low mood's impact on behavior for non-signaling benefits likely conveys information as cues in addition to what might exist for signaling purposes. In other words, the signaling component of depression may come from the upregulation of symptoms beyond what would be optimal for non-signaling purposes rather than the creation of entirely new symptoms. Suicidality is likely an exception to this, due to the likelihood most suicide attempts likely do not have non-signaling functions (see @deCatanzaro1991 for a possible exception).

## Responses to depression and suicidal behavior 

For depression and suicide attempts to be worth the costs involved with signaling, substantial benefits in terms of social support are expected, either in the form of temporary but impactful support or long-term changes in how one is treated [@Hagen2003; @SymeEtAl2016]. This can take the form of reconciliations in failing relationships, increased fairness in cooperative relationships, or an increase of resources being directed towards an individual. Of all the benefits, a reduction in physical abuse may be one of the most important due to the threat of death or severe injury to oneself and one’s family.

Most studies examining responses to depression have been observational and focused on responses outside of changes in social support. In general, most report aversive responses and high levels of rejection towards the depressed, and depressed individuals are often linked to impaired social functioning [@EvraireDozois2011; @GotlibLee1989; @HirschfeldEtAl2000]. One explanation for this is that depressed individuals engage in behaviors that drive others away. In line with this reasoning, both excessive reassurance seeking and negative feedback seeking, which often accompany depression, have been shown to elicit negative emotional responses from others and increase the chance of rejection from others [@JoinerMetalsky1995; @StarrDavila2008]. Negative interactions are also seen in families as depressed children tend to report less support and more negative responses from parents than non-depressed children, although it has not been demonstrated that such interactions did not lead to depression in the first place [@MesserGross1995; @SchwartzEtAl2012].

Despite being looked for less frequently, beneficial responses to depressed individuals are also reported. This is seen in that depression has been shown to reduce the likelihood of aggression within families [@hops1987; @dadds1992], including those already experiencing high levels of conflict [@SheeberEtAl2001]. There is also limited evidence that depression can result in increased levels of support from others outside the family. For example, non-depressed individuals were shown to be more cooperative and exhibit more caretaking behavior towards their college depressed roommates over the course of their time living together [@HokansonEtAl1986]. Likewise, a study examining responses to confederates trained to appear ‘depressed’, ‘anxious’, or ‘normal’ found that individuals were more likely to provide advice and support to ‘depressed’ individuals than ‘anxious’ or ‘normal’ individuals despite the confederate in ‘depression’ condition being viewed more negatively than the other conditions and the ‘depression’ condition involving similar levels of distress compared to the ‘anxious’ condition [@stephens1987]. 

Unfortunately, there are few studies examining how individuals respond to suicide attempt survivors, with the majority examining responses after death. In those which exist, stigmatization is commonly reported within and outside families [@FreyEtAl2016; @ScoccoEtAl2012], with perceptions of survivors as being weak, selfish, mentally ill, and antisocial all being commonly reported [@BatterhamEtAl2013; @TzengLipson2004]. Although these types of stigmatization have been reported across relationships types, the likelihood of stigma has been found to differ based on who the survivor is to the person being questioned. This is seen in a study which found that individuals felt the most stigma from members of their preexisting social networks than mental health providers, with family members providing more stigmatization (57.1% of the time) than romantic partners (41.2% of the time) and close friends (28.2% of the time) [@FreyEtAl2016]. 

Although negative responses to suicide survivors are more often studied and reported, positive responses have also been found. Increased social support and changes to important relationships have been reported to follow suicide attempts with some indication that these effects may hold long term [@Stengel1956]. For example, a study of 100 women who survived suicide attempts found that in 75 cases individuals gained identifiable benefits through the attempt, with 41 individuals benefiting from reconciliations with others [@Lukianowicz1971]. Unlike many Western countries, where suicide is often viewed as pathological [@Hidaka2012], members of traditional societies have been reported to view suicide attempts as cries for help rather than mental illness [@Shostak1981]. For example, attempts have been described as ways of escaping unwanted marriage arrangements, persistent abuse, or a lack of support in obtaining mates by those engaging in the behavior and observers [@GutierrezdePinedaMuirden1948; @Hilger1957; @Karsten1935; @Tessmann1930; @Wilson1960], consistent with the idea that help is at least occasionally provided. In support of this, 30 out of 84 examples of suicidal behavior included in the HRAF resulted in positive changes for the survivor, even though suicide was rarely if ever the main topic [@SymeEtAl2016].

# Study aims and predictions

The primary aim of this study was to examine responses to various forms of emotional signaling in an experimental vignette study. We predicted that costlier signals would increase the participants' self-reported belief in the signaler's claims and the likelihood of them providing help, with this prediction entailing the following rank from low to high cost: verbal request, crying, mild-depression, depression, and suicide attempt. We also predicted that changes in the likelihood participants would provide support would be mediated by changes in the participants' belief in the signaler's need pre-signal to post-signal.

# Materials and methods

## Design

This study utilized a between-subjects pretest-posttest design to examine how four different emotional signals would influence (1) the degree participants believed a fictional character to be in need (*Belief*) and (2) the likelihood they would provide help (*Action*) relative to a simple verbal request (the control condition), in four different vignettes, for a total of 20 conditions. In this design, the outcomes are measured at pre-treatment (T1), participants are then randomized into either a control group or one or more treatment groups (one of the emotional signals), and the outcomes are measured again (T2). Regression models are used to determine the effect of the treatment conditions on the posttest outcome variables, relative to the control condition, controlling for pretest levels of the outcome variables. We also explored within-subjects effects of the signal on outcomes at T2 compared to T1.

## Vignettes

Four vignettes were used in this study in which both the presence of conflicts of interest and the existence of private information about one's need were emphasized. Due to the expectation that depression and suicide are likely too costly of signals for relatively minor instances of need, we designed this study to involve severe scenarios associated with suicide attempts both ethnographically and clinically (i.e, sexual assault, non-sexual assault, and thwarted marriage).

Participants in the US sample were randomly assigned to either the "basketball coach," "romantic partner," or "brother-in-law" vignettes. In the "basketball coach" scenario, the star player on a university basketball team comes to the participant (who is asked to imagine themselves as the athletic director at a university) and claims she was sexually assaulted by the coach. However, there are indications the player might be lying to get more playing time.

The "brother-in-law" vignette involved participants imagining that they let their sister, brother-in-law, and niece move in with the participant and their 15-year-old daughter after their sister's family lost their house in a fire. During this time, the participant's daughter becomes jealous of her cousin before accusing her uncle of sexually assaulting her.

The "romantic partner" vignette had participants imagine that they found a highly desirable partner after years of being single. However, their 13-year-old daughter, who has a history of trying to get in the way of their relationships is clearly not happy with this. After a period of sustained conflict, she accuses the romantic partner of physically assaulting her but cannot produce any evidence.

Those in the Indian sample only saw the "thwarted marriage vignette." In this scenario, the participant and their family is described as trying to arrange a dowry for their older daughter so she can marry a man she already loves. After the man's family demands more money, the participant's family tries to find a second man, who the older daughter claims to find unattractive. Due to the younger daughter in the story only being two years younger, any additional investment in the dowry will come at the younger daughter's expense.

The full vignettes are available in the SI.

## Signals

The sole randomized treatment in this study was the signal participants saw, with these signals being one of (1) the control condition: a verbal request without additional signaling, (2) crying, (3) mild-depression, (4) depression, and (5) a suicide attempt.

As many depression scales include measures of crying [for discussion on the relationship between crying and depression see @bylsma2020] and depression is a major risk factor for suicide [@BostwickPankratz2000; @Kessler2012], each signal was designed to build off of the less costly signals so that each signal contained elements of any signal it was costlier than. Although referred to as signals throughout the paper for brevity, we expect that the hypothesized signals of need, like many signals, will also provide information to others in the form of cues [For discussion on the evolution of cues from signals see: @biernaskie2018; @Tiokhin2016; @Steinkopf2015]. The full text for the signals are available in the SI.

## Lessons learned from a pilot study

The current study is a refinement of a vignette-based pilot study (N=`r nrow(signalingdata2018b)`) with very similar signals and outcomes. In the pilot study, there was a single vignette in which an imagined sister claims her child needs an expensive emergency medical procedure, and asks for a very large loan. We learned five major lessons from the pilot study that influenced the design of the current study. First, as predicted, belief in the sister's need was highest in the *Depression* condition and lowest in the *Verbal request* condition, giving us confidence to continue this research (see Figure \@ref(fig:piloteffects)). Second, to rule out that believability and willingness to help were simply artifacts of the sister's psychiatric distress, we included a "signaling" condition in which the sister exhibited schizophrenic symptoms. As predicted, believability and willingness to help in this condition were dramatically lower than in any other condition, ruling out this alternative explanation. We therefore did not include the schizophrenic condition in the current study. Third, baseline (T1) believability was very high (mean = `r signif(pilot_results$meanT1belief, 2)` and median = `r signif(pilot_results$medianT1belief, 2)` on a 0-100 point scale), leaving little scope for signals to increase believability. In the current study we therefore endeavored to provide details in the vignettes that would undermine the victim's believability at baseline. Fourth, contrary to our hypothesis, the *Suicide threat* and *Suicide attempt* signals reduced believability relative to *Depression*. We surmised that the sister's adversity -- a sick child -- was not sufficient to justify a suicide attempt, thus undermining her believability. In the current study the signaler herself is therefore the victim, and the forms of adversity -- physical and sexual assault and thwarted and forced marriage -- are those that are frequently associated with suicidality. Fifth, the content of the vignette matters, so instead of a single vignette we decided to use four different vignettes to help determine the extent to which signaling effects were sensitive to seemingly irrelevant details of the vignettes.

## Preregistration

We preregistered four OLS linear regression models and one mediation model (https://osf.io/g3s6n). The first two regression models tested the effects of our ordinal signal variable on Time 2 *Belief* and *Action*, controlling for T1 *Belief* and *Action*, respectively (a pre-test/post-test design):

`r extract_eq(models_prereg$Model$m1, intercept='beta', wrap=T)`
`r extract_eq(models_prereg$Model$m3, intercept='beta', wrap=T)`

The second two models added interaction terms between signal and T1 *Belief* and *Action*, with the rationale that the signals might have reduced effect for individuals who indicate high levels of belief and helping at baseline: 

`r extract_eq(models_prereg$Model$m1b, intercept='beta', wrap=T)`
`r extract_eq(models_prereg$Model$m3b, intercept='beta', wrap=T)`

The mediation model is in Figure \@ref(fig:medmodel):

```{r medmodel, fig.cap="Possible causal effects of the signal on helping behavior. A credible signal of need increases belief that the victim is telling the truth and needs help, which increase the likelihood of helping. The direct path from the signal to action represents other causal effects of the signal on observers, such as perceptions of the victim's emotional state (and other factors not measured in this study) that might alter observer behavior."}

knitr::include_graphics('paths.svg')
```

Specifically, the mediation and outcome models, respectively, are:

`r extract_eq(med_prereg$model.m, intercept='beta', wrap=T)`
`r extract_eq(med_prereg$model.y, intercept='beta', wrap=T)`

For reasons we explain in the *Statistical methods* section, we report the OLS models in the Supplementary Information (SI), and in the Results section instead report conceptually equivalent generalized linear regression models (GLMs).

## Power analysis

```{r}
pwr80_90 <- 
  power_curve %>% 
  group_by(Base) %>% 
  summarise(
    pwr80 = min(sample_size[power>=.8]), 
    pwr90 = min(sample_size[power>=.9])
  )
```

We used the pilot data to estimate the sample sizes needed to detect an effect of the *Depression* vs. *Verbal request* signal (two conditions) on *Belief* in the victim's need. Power was about 80% for a sample size of about `r pwr80_90[[2,2]]`, and was about 90% for a sample size of about `r pwr80_90[[2,3]]`. We also estimated the sample size needed to detect an effect of *Depression* vs. a *Control* condition that was meant to not include any signaling (but which we realized, post hoc, did include a verbal request). Power was about 80% for a sample size of about `r pwr80_90[[1,2]]`, and was about 90% for a sample size of about `r pwr80_90[[1,3]]`. See Figure \@ref(fig:power). Given our $1500 USD budget, we aimed for a sample size of 125 for two conditions, and 1250 for all conditions in the study. For more details, see the SI and XX.

## Sampling

Participants for this study were recruited from Amazon Mechanical Turk, a crowdsourcing platform that allows for the creation of Human Intelligence Tasks (HITs) that workers can complete for pay. As Amazon provides the infrastructure, it allows for a relatively low-cost way of collecting data for academic research at the cost of the data not being representative of any real population [@thomas2017]. Despite this limitation, it is more diverse than most convenience samples [@kennedy2020; @DworkinEtAl2016; @thomas2017], with turkers in the US mainly differing in being younger, more educated, and lower income than the greater U.S. population [@RossEtAl2010; @boas2020]. Indian samples are also likely to be younger and more educated than the general population but differ from US samples in that they are less diverse, have higher incomes relative to other Indians, and are more likely to come from regions with good internet access [@boas2020].
	
Overall, the quality of data provided by MTurk workers tends to resemble that of university sample pools [@NeckaEtAl2016; @robinson2019], with there being some indication that it might outperform university samples due to its previously mentioned diversity [@thomas2017] or the overall attentiveness of turkers [@HauserSchwarz2016]. For this reason, concerns about data quality come primarily from the threat of bot use or respondents faking their location to take surveys in a language they do not understand well, with there being little evidence of the former [@kennedy2020] and the risk of the latter able to be minimized through well designed attention checks, timed responses, or general study design [@aguinis2020; @huang2015; @kennedy2020; @thomas2017).
	
## Sample

All participants in the US sample were over 18, located in the United States, had a HIT approval rate of over 98%, and had successfully completed at least 100 HITs as anything less results in approval rates of 100% [@kennedy2020]. The Indian sample was the same except for only including participants who resided in India at the time of the study and the inclusion two participants who were included with a stricter HIT approval rate of over 99%.

To reduce the risk of low quality responses, we used a variety of attention checks, which we clearly labeled in an attempt to avoid demand characteristics. The first was shown immediately after the consent form and provided participants with a random word and asked them to enter the vowels in the order in which they are found in the word. Capitalization did not matter and no word contained repeat vowels. Failure to enter the correct vowels in order resulted in the participants being filtered out of the study. The second followed the vignette and involved asking three questions about the story that should be easy to answer for anyone paying attention. 

We also filtered out those who read the vignette faster than one-third the time it took MG to read it as fast as possible without concern for comprehension. This occurred immediately after participants finished the vignette.

Informed consent was given by proceeding with the study after reading the consent form, and all participants who passed the attention checks were paid $1 for their time, with us estimating it would take 4-8 minutes before data collection. This study was certified exempt by the Washington State University Human Research Protection Program (certification number: 18528-001).

## Survey

After accepting the HIT on MTurk, every participant took the survey using Qualtrics. In every instance, the elements of the survey were laid out in the following order: 1) the consent form, 2) the vowel attention check, 3) the main body of the vignette, 3) the pre-test baseline measures (T1), 4) the story content attention check, 5)  the signal portion of the vignette, 7) the post-test outcome measures (T2-T3), 8) an attention check that did not filter out participants (T4), 9) a set of demographic questions, and 10) an open feedback prompt. The full survey can be found in the SI.

### Baseline measures (T1)

In both samples, we recorded self-reports on how much participants believed the signaler to be telling the truth (T1Belief) and the likelihood of them helping the signaler as requested (T1Action). With the thwarted marriage vignette we also asked how they would split the money they had saved for the dowry between their daughters (T1Divide). In every instance, the order of the questions was randomized to avoid order effects [@KrosnickAlwin1987].

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. Due to findings that a slider's starting position may bias results [@liu2019], each T1 slider started fully to the left (0) (don't exactly remember the logic for left instead of click to start). For the exact wording of the question and the labels on the sliders, see table 1.

We also asked which emotions participants felt the signaler was experiencing using a multiple choice question in which they could select multiple options. This included emotions directly related to the signals (e.g., sad, depressed, and suicidal), states which suggest genuine need (traumatized and violated), and states which suggest deception (e.g., deviousness or jealousy). The complete list can be found in the SI.

### Post-treatment measures (T2-T4)

We asked questions identical to those asked at T1 as the main post-treatment variables of interest (T2). For the belief, action, and divide variables, the position of the slider starting where participants placed it at T1. For the T2 emotion variables this involved no changes from the T1 version. 

As both a validity check and a way to understand the degree participants would be willing to help if they believed the participant completely, we presented participants with strong evidence that the claims were true (T3Action). In the US sample, this involved video evidence of the event in question occurring or a similar event after the fact. With the thwarted marriage vignette, this involved the participant seeing the man their daughter wants to marry trash talk their daughter and their family. As a final attention check, we presented participants with the T3Action question but included instructions to ignore this and drag the slider all the way to the left (T4AttentionCheck). We did not exclude anyone from the study for failing this nor did we withhold payment due to the time participants had invested in the study at this point.

Variable | Vignette | Wording | Coding 
---------|----------|---------|-------
T1Belief | Basketball | How much do you believe the star player is telling the truth about the coach? | 0: Not at all; 100: Completely
T1Belief | Brother-in-law | How much do you believe your daughter is telling the truth about your brother-in-law? | 0: Not at all; 100: Completely
T1Belief | Romantic partner | How much do you believe your daughter is telling the truth about your partner? | 0: Not at all; 100: Completely
T1Belief | Thwarted Marriage | How much do you believe your older(bold) daughter is telling the truth about the second man not being a good match for her?  | 0: Not at all; 100: Completely
T1Action | Basketball | How likely are you to recommend that the president of the university should suspend the coach? | 0: Very unlikely; 100: Very likely
T1Action | Brother-in-law | How likely are you to kick your brother-in-law out? | 0: Very unlikely; 100: Very likely
T1Action | Romantic partner | How likely are you to kick your partner out? | 0: Very unlikely; 100: Very likely
T1Action | Thwarted Marriage | How likely are you to provide the large dowry so your older(bold) daughter can marry the man she is in love with?  | 0: Very unlikely; 100: Very likely
T1Divide | Thwarted Marriage | How would you plan on dividing the money between your daughters?  | 0: Much more for the younger daughter; 50: Even division; 100: Much more for the older daughter

Table: (\#tab:outcomes) Outcome variables in each vignette.

### Demographic Questions

The final part of the survey was a brief demographic questionnaire which examined: (1) the age, (2) sex, (3) number of siblings, (4) number of sons, (5) number of daughters, (6) current relationship status, (7) highest level of education, and (8) the annual household income of each participant.

## Statistical analyses

Our pre-test and post-test measures, *T1 & T2 Belief* and *T1 & T2 Action*, were all measured on a 0-100 point scale. A substantial number of participants rated their beliefs and actions as exactly 0 or exactly 100 at either T1 or T2. OLS linear regression models are not suitable for a closed and bounded distribution with so many values on the boundary because the residuals would not be normally distributed or have constant variance. Beta regression, a possible alternative, is only defined on the open interval, (0, 1), whereas our data had many values on the boundaries (a closed interval). A variant of beta regression, zero-one inflated beta, is inappropriate because it assumes different processes for the generation of data on the boundaries vs. the interior, and we had no reason to believe that our zeros and ones were generated by a different process than the rest of our observations. 

We therefore fit models of our two outcome variables (which we rescaled to the continuous interval [0, 1]) in four ways. First, we fit the preregistered OLS models (reported in the SI). Second, we fit fractional regression models, introduced in a seminal paper by @papke1996 for continuous outcomes on the interval [0, 1], which, in brief, involve fitting a quasi-likelihood version of logistic regression with robust standard errors. This method is implemented in the R `frm` package [@ramalho2019]. Third, we fit GLM logistic regression models, computing 95% CI's via bootstrapping, using the R `glmmboot` package created to model continuous outcomes on [0, 1] [@humphrey2020]. Finally, we fit GLM models with the quasibinomial family and logit link using the `glm` function in R stats library, which can accept continuous outcomes on [0, 1]. The estimated coefficients, standard errors, and confidence intervals produced by the latter three methods were *very* similar, and often identical. We therefore proceeded to fit and report models using `glm` with the quasibinomial family and logit link because model objects produced by this function are compatible with numerous other R packages for estimating and plotting effects.

Although two of our four preregistered OLS regression models included an interaction term, the rationale for it was less convincing for quasibinomial models using a logit link, which maps [0, 1] to (-$\infty$, $\infty$). The interaction would also increase the number of parameters to estimate, increasing the risk of overfitting. We therefore did not include it in the GLM models (but we still report OLS models with the interaction term).

Our intervention was *signal*, an ordinal variable with the following preregistered rank order: verbal request, crying, mild depression, depression, suicide attempt. We used the default `glm` settings, which code such variables as orthogonal polynomial contrasts. Our primary confirmatory regression models were as follows (where the signal subscripts indicate polynomial degree):

`r extract_eq(models$Model$m1, intercept='beta', wrap=T)`
`r extract_eq(models$Model$m3, intercept='beta', wrap=T)`

To test our hypothesis that the likelihood of acting to help the victim would be largely mediated by a signal's positive impact on the participant's belief in the victim's need, we used the `mediation` package [@tingley2014] to fit a mediation model for *Depression* vs. *Verbal request*. The mediation and outcome models, respectively, were:

`r extract_eq(med1$model.m, intercept='beta', wrap=T)`
`r extract_eq(med1$model.y, intercept='beta', wrap=T)`

## Data availibility

http://doi.org/10.5281/zenodo.4637904

# Results

```{r}
nationT0 <- table(signaling2020$sample)
removed <- sum(!signaling2020$CompleteSurvey)
removed_pct <- 100*signif(removed/nrow(signaling2020), 2)
nationT1 <- table(d$sample)
sexT1 <- table(d$Sex)
n_cond <- xtabs(~signal+vignette, d)
```

The study was started by N=`r nrow(signaling2020)` participants who clicked the link to Qualtrics (`r nationT0['American']` US and `r nationT0['Indian']` India). After removing participants who did not finish the survey, failed attention checks, or moved through the study at an unrealistic pace (N=`r removed`, `r removed_pct`%), our final sample was N = `r nrow(d)` (`r nationT1['American']` US and `r nationT1['Indian']` India), with `r sexT1['Male']` males and `r sexT1['Female']` females. The median number of participants per condition was `r median(n_cond)` (min=`r min(n_cond)`, max=`r max(n_cond)`). For the number of participants in each condition, see Table \@ref(tab:conditions). For summary statistics, see Table \@ref(tab:sumtab). For the distributions of participants by age, income, and nationality, see Figure \@ref(fig:socioeconomic).

```{r sumtab}

d$minutes <- d$TotalTime/60

vars <- c(
  Age = 'Age (years)',
  Income = 'Income (USD)',
  years_education = 'Education (years)',
  minutes = 'Time to complete (minutes)',
  T1Belief = 'T1 Belief',
  T1Action = 'T1 Action',
  T1Divide = 'T1 Division',
  T1LowMood = 'T1 Low mood',
  T1Manipulative = 'T1 Manipulative',
  T2Belief = 'T2 Belief',
  T2Action = 'T2 Action',
  T2Divide = 'T2 Division',
  T2LowMood = 'T2 Low mood',
  T2Manipulative = 'T2 Manipulative',
  T3Action = 'T3 Action'
)
kable(custom.summarize(d, vars), caption = "Summary statistics for study variables. Original 0-100 values scaled to 0-1. Indian participants reported income in rupees, which we converted to USD at the current exchange rate (1 rupee = 0.014 USD).") %>% kable_styling()
```

## Distributions of beliefs and actions at Time 1

Across the four vignettes, mean belief of the victim was relatively low at baseline (T1), Mean = `r signif(mean(d$T1Belief, na.rm = T), 2)`, albeit with wide variation, SD = `r signif(sd(d$T1Belief, na.rm=T), 2)`; `r sum(d$T1Belief==0)` participants (`r 100*signif(sum(d$T1Belief==0)/nrow(d), 2)`%) rated their belief = 0, and `r sum(d$T1Belief==1)` participants (`r 100*signif(sum(d$T1Belief==1)/nrow(d), 2)`%) rated it as = 1. The distribution of likelihood of helping the victim (action) was similar, Mean = `r signif(mean(d$T1Action, na.rm = T), 2)`, SD = `r signif(sd(d$T1Action, na.rm=T), 2)`, with `r sum(d$T1Action==0)` participants (`r 100*signif(sum(d$T1Action==0)/nrow(d), 2)`%) rating their action = 0, and `r sum(d$T1Action==1)` participants (`r 100*signif(sum(d$T1Action==1)/nrow(d), 2)`%) rated action as = 1. Although not a pre-registered prediction, T1 Belief and T1 Action were highly correlated across the four vignettes, consistent with help being worth providing if the signaler's claims were true (see Figure \@ref(fig:t1plot)).

```{r t1plot, fig.width=10, fig.height=10, fig.cap="The distributions of Time 1 Belief vs. Time 1 Action, by vignette. Each dot is one participant. Original 0-100 values rescaled to 0-1. Lines fit by linear regression. Dot size indicates the number of overlapping points. Black cross is the mean of each variable."}
plot_T1Belief_T1Action
```

## Confirmatory: Effects of signals on beliefs and actions

```{r}
sm1 <- summary(models$Margins$m1)
suicide_belief_eff <- signif(sm1$AME['signalSuicide attempt']/sd(d$T2Belief), 2)

sm3 <- summary(models$Margins$m3)
suicide_action_eff <- signif(sm3$AME['signalSuicide attempt']/sd(d$T2Belief), 2)

signal_belief_ttest <- t.test(100*d$T2Belief, 100*d$T1Belief, paired = T)
signal_action_ttest <- t.test(100*d$T2Action, 100*d$T1Action, paired = T)

```

As predicted, there was a strong statistically significant positive causal effect of the ordinal signal on T2 Belief and Action (controlling for T1 Belief and Action, respectively), with the effect of each signal increasing in the predicted rank order. See Figure \@ref(fig:beliefeffect) and models m1 and m3 in Tables \@ref(tab:regressiontables) and \@ref(tab:anovatables).

The marginal between-subjects effect of the high cost *Suicide attempt* vs. low cost *Verbal request* control on T2 *Belief* (averaging over all four vignettes and all values of T1 *Belief*) was an increase of `r fmt_margins(models$Margins$m1, 'signalSuicide attempt')` points on the original 0-100 point scale. The equivalent increase for T2 *Action* was `r fmt_margins(models$Margins$m3, 'signalSuicide attempt')` points. Given that the standard deviations of T2 *Belief* and *Action* are `r signif(sd(100*d$T2Belief), 2)` and `r signif(sd(100*d$T2Action), 2)`, respectively, these represent increases of `r suicide_belief_eff` and `r suicide_action_eff` SD's, respectively.

```{r beliefeffect, fig.width=6, fig.height=8, fig.cap=paste0("The effect of the signals on T2 Belief (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively. Effects plotted for T1 values set to their median value (T1Belief=", median(d$T1Belief, na.rm=T), ", T1Action=", median(d$T1Action, na.rm=T), "). Fit using generalized linear regression models with the quasibinomial family. For coefficents, p-values, and other statistics, see models m1 and m3 in Tables \\@ref(tab:regressiontables) and \\@ref(tab:anovatables).")}
(effectsPlots$m1plot + theme(axis.text.x = element_blank()))/effectsPlots$m3plot + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

Within-subjects, the average effect across all signals on *Belief* at T2 compared to T1 was `r inline_ttest(signal_belief_ttest)`, and on *Action* was `r inline_ttest(signal_action_ttest)` (on the original 0-100 point scale). For the change in *Belief* and *Action* for, T1 to T2 for all participants in all conditions, see Figure \@ref(fig:rawbeliefactions2d).

## Exploratory: effects of signals by vignette

```{r}

thwart_sv_bdiff <- visreg_diff(models$Model$m2, 'Thwarted marriage', 'Suicide attempt', 'Verbal request', d)

thwart_sv_adiff <- visreg_diff(models$Model$m4, 'Thwarted marriage', 'Suicide attempt', 'Verbal request', d)

brother_sv_bdiff <- visreg_diff(models$Model$m2, 'Brother-in-law', 'Suicide attempt', 'Verbal request', d)

brother_sv_adiff <- visreg_diff(models$Model$m4, 'Brother-in-law', 'Suicide attempt', 'Verbal request', d)

```

The effects of the signals varied substantially by vignette. Signals had the smallest effect in the "Thwarted marriage" vignette, where the effect of the *Suicide attempt* signal on *Belief* and *Action* relative to the *Verbal request* signal was only `r thwart_sv_bdiff` and `r thwart_sv_adiff` points, respectively (on the original 0-100 point scale). Signals had the largest effect in the "Brother-in-law" vignette, where the effect of the *Suicide attempt* signal on *Belief* and *Action* relative to the *Verbal request* signal was `r brother_sv_bdiff` and and `r brother_sv_adiff` points, respectively. See Figures \@ref(fig:beliefvignette) and \@ref(fig:ecdf).

```{r}
bil_b_suicide <- signif(effect_sizes$est[effect_sizes$id=="Suicide attempt,Brother-in-law" & effect_sizes$Outcome=="Belief"], 2)

rom_b_crying <- signif(effect_sizes$est[effect_sizes$id=="Crying,Romantic partner" & effect_sizes$Outcome=="Belief"], 2)

bil_a_suicide <- signif(effect_sizes2$effsize[effect_sizes2$id=="Suicide attempt,Brother-in-law" & effect_sizes2$Outcome=="Action"], 2)

bil_a_verbal <- signif(effect_sizes2$effsize[effect_sizes2$id=="Verbal request,Brother-in-law" & effect_sizes2$Outcome=="Action"], 2)

```

Effect sizes (Cohen's d), between- and within-subjects, reveals a similar pattern. The largest between-subjects effect was *Suicide attempt* vs. *Verbal request* on *Belief* in the "Brother-in-law" vignette, d = `r bil_b_suicide`, and the smallest was the effect of *Crying* on *Belief* in the "Romantic partner" vignette, d =`r rom_b_crying`. The largest within-subjects effect was *Suicide attempt* vs. *Verbal request* on *Action* in the "Brother-in-law" vignette, d = `r bil_a_suicide`, and the smallest was the *negative* effect of *Verbal request* on *Action* in the "Brother-in-law" vignette, d =`r bil_a_verbal`. See Figures \@ref(fig:effectsizes) and \@ref(fig:effectsizes2).

We note the following patterns, which we will return to in the Discussion section. First, as we intended, the distributions of *Belief* and *Action* in the *Verbal request* control condition were very similar to their distributions at T1 baseline across vignettes. In the "Basketball coach" and "Brother-in-law" vignettes, though, their distributions in the *Verbal request* condition were shifted to somewhat lower values relative to baseline (i.e., the verbal request slightly *reduced* belief and action in those vignettes). Second, in the "Romantic partner" and "Brother-in-law" vignettes, the effect of *Crying* differed little from *Verbal request*, but in the "Basketball coach" vignette, it differed little from the *Depression* and *Suicide attempt* signals. Third, the effect of the *Suicide attempt* signal on *Belief* was similar to that of *Depression* across all vignettes, but had a noticeably greater effect on *Action* in the "Romantic partner" and "Brother-in-law" vignettes. Fourth, there was little difference between the effect of the *Mild depression* vs. *Depression* signals. See Figures \@ref(fig:beliefvignette) and \@ref(fig:ecdf).

```{r beliefvignette, fig.width=8, fig.height=8, fig.cap=paste0("The effect of the signals on T2 Belief (top) and Action (bottom) in each vignette, controlling for T1 values of Belief and Action, respectively. Effects plotted for T1 values set to their median value (T1Belief=", median(d$T1Belief, na.rm=T), ", T1Action=", median(d$T1Action, na.rm=T), "). Fit using generalized linear regression models with the quasibinomial family. For coefficents, p-values, and other statistics, see models m2 and m4 in Tables \\@ref(tab:regressiontables) and \\@ref(tab:anovatables).")}
(effectsPlots$m2plot + theme(axis.text.x = element_blank()))/effectsPlots$m4plot + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```


```{r ecdf, fig.width=11, fig.height=6, fig.cap="Empirical cummulative distribution functions for Time 2 Belief (top) and Action (bottom) compared to their Time 1 baseline values (red), by signal and vignette. Dots indicate median values of Belief and Action, by signal and vignette."}
ecdf_plot
```

## Confirmatory: The effect of signals on action is largely mediated by belief

According to our theoretical model, credible signals of need increase observers' belief that the victim is telling the truth. Increased belief then increases the likelihood of helping the victim (action). See Figure \@ref(fig:medmodel).

The total effect of the *Depression* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med1b$tau.coef, 3)` points from T1 to T2 (on the original 100-point scale). Of this increase, `r 100*signif(med1b$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth. See Figure \@ref(fig:mediation). The total effect of the *Suicide attempt* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med4b$tau.coef, 3)` points from T1 to T2. Of this increase, `r 100*signif(med4b$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth (Check if med4 or med4b is correct; they look identical to me).

```{r mediation, fig.height=4, fig.cap="The effect of Depression (treatment) vs. Verbal request (control) on Action (0-1 scale) is largely mediated by Belief. ACME: Average causal mediation effect. ADE: Average direct effect. The mediation model controlled for T1 Belief, and the outcome model controlled for T1 Belief and T1 Action. Both models were GLMs with the binomial family (the mediation package does not support the quasibinomial family)."}
ggmediation(med1b)
```

## Exploratory: Effects of signals on victim's perceived emotional state

The proportion of participants inferring each emotion in each vignette at T1 is depicted in Figure \@ref(fig:t1emotionheatmap), after each signal at T2 in Figure \@ref(fig:t2emotionheatmap), and the change in proportions from T1 to T2 in Figure \@ref(fig:diffemotions). The increase in inferred depression in the *Depression* and *Suicide attempt* conditions, but not the *Verbal request* and *Crying* conditions, helps validate these manipulations, as does the increase in inferred suicidality in the *Suicide attempt* condition.

To further explore the effect of signaling on participants' inferences of the signaler's emotional state, we created four new variables: *Low mood* (T1 and T2) was the sum of the binary variables *Depressed*, *Distressed* and *Sad*; and *Manipulative* (T1 and T2) was the sum of the binary variables *Devious* and *Jealous.* The mean change in the *Low mood* and *Manipulative* variables from T1 to T2 for each signal in each vignette reveals a large decrease in inferred *Manipulation* for costly signals, and a large increase in inferred *Low mood* (with the exception of the Thwarted marriage vignette, in which changes are small). See Figure \@ref(fig:emotionsplot)).

```{r emotionsplot, fig.width=10, fig.height=8, fig.cap="The within-subjects effect of the signals on participants' inferred *Manipulative* and *LowMood* emotional states of the victim, from T1 (arrow base) to T2 (arrow head), for each signal in each of the four vignettes. Thick arrows: mean change. Thin arrows: means of 500 bootstrap resamples."}
plot_emotions
```

## Exploratory: Demographic associations

The demographic variables were strongly confounded with the Indian sample (younger, lower income, more years of education) vs. the US sample (older, higher income, fewer years of education), which, in turn, were confounded with vignette (responses in the Thwarted marriage vignette differed substantially from those in the other vignettes). We therefore restricted our exploration of the sociodemographic variables to the US sample.

Overall, women participants were more likely to believe and help the victim than men, and younger participants were more likely to believe and help the victim than older participants. The caveats are that the sex differences in *Belief* only emerged for the costly signals, the sex differences in *Action* were quite small (but applied across all signals). We found no significant effects of *Income* or *Education* on *Belief* or *Action* (results not reported). See Figures \@ref(fig:demographicsignal), \@ref(fig:demovignettebelief), and \@ref(fig:demovignetteaction).


```{r demographicsignal, fig.width=11, fig.height=8, fig.cap="Demographic effects on *Belief* and *Action*. **Top**: Association of age with T2 Belief and Action, controlling for T1 Belief and Action, respectively, and sex and vignette. **Bottom, left**: Sex interacting with the effect of the signals on T2 Belief controlling for T1 Belief, age, and vignette. **Bottom, right**: Main effect of sex on T2 Action, controlling for T1 Action and age. Estimated using generalized linear regression models with the quasibinomial family."}
# effectsPlots$m24plotAge / effectsPlots$m24plotSex 

(effectsPlots_US$m24plotAge + ylim(0, 1) + effectsPlots_US$m25plotAge + ylim(0, 1))  / (effectsPlots_US$m24plotSex + effectsPlots_US$m25plotSex + plot_layout(widths = c(5,1)))

```


# Discussion

## The effect of signaling on belief and likelihood of support

Consistent with costly signaling theory, costlier signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler than the less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. However, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family members or a romantic partner). With the thwarted marriage vignette, the fact that support came at the cost of one's younger daughter may explain the relatively small effect of signal strength. However, data from the Indian sample appears to be of lower quality, limiting our confidence in this interpretation (see limitations for more information). It is also possible that our research in designing the vignettes was insufficient for the population or the content of the vignette.

In the "brother-in-law" and "romantic partner" scenarios, *Crying* had little effect on the magnitude of pro-signaler responses, while both *Depression* conditions increased support to similar degrees. One potential reason for this is the increase in symptoms from mild depression to depression was not costly enough to result in a large effect as the symptoms only differed in severity (e.g., grades dropping from As to Bs [mild depression] or Cs [depression]). Such small changes may be less impactful in vignette form than in real-life scenarios in which the effects of signaling may increase in severity and be prolonged over a lengthy period of time.

While the effect of *Suicide attempt* on belief at T2 was similar to the *Depression* conditions, it resulted in greater likelihood of support. This may be due to our decision to measure belief in need as the degree participants believed the claims of the signaler (i.e., the assaults or incompatibility with a potential husband in the context of arranged marriage) rather than a more inclusive measure. For this reason, signal cost may have provided additional information about the costliness of the signaler's adversity in ways that would still be consistent with signaling theory.

Support for this interpretation comes from our mediation analyses, which showed that the likelihood of help was largely, but not entirely, mediated by signal's effect on belief in need in this study but totally mediated by belief in need when measured as belief in need of financial support in the pilot study. However, the incomplete mediation in this study may have arisen if participants viewed the signal as providing information about adveristy outside of the signaler's claims. For example, participants who saw the "brother-in-law" vignette may have thought the daughter was lying about the assault due to jealously over her cousin, with the stronger signals still conveying the severity of the competition between her and her cousin.

## Broader discussion

Our results on the connection between belief in need and likelihood of help are consistent with costly signaling and bargaining approaches to depression and suicide. In addition to the effect of signal cost on support being largely mediated by the signal's effect on belief in need, belief in need and likelihood of support were highly correlated but distinct at T1. This situation should arise in any instance when a target of a signal of need would benefit from helping if the need was genuine and known to reasonable precision. Importantly, such problems are common among modern humans and have likely been so since individuals could increase their inclusive fitness through supporting kin or important cooperative partners in times of genuine need [@SugiyamaSugiyama2003; @tooby1996].

Although this study was not designed to examine whether the costs of signaling would be made up for by the increased support stemming from depression and suicide attempts directly, the results are consistent with this hypothesis. In general, the costs required to honestly convey one's qualities should be expected to increase with the potential payoffs of deception [@Rosenthal1993], with the vignettes in this study emphasizing the existence of private information and sizable conflicts of interest to increase the perceived risk of being taken advantage of. In similar real-world scenarios, this difficulty of signaling honest need would coincide with the substantial fitness-relevant benefits of preventing future assaults or being able to marry the partner of one's choice. These benefits coupled with the fact (too strong?) that depression and suicide attempts become relatively less costly with increased need [@Hagen2003], suggest that the increased support seen in this study may be worth the costs required to signal need more often than is often appreciated when counting the entire costs of depression as the signal rather than only the upregulation of symptoms for a communicative purpose.

## Limitations

The main limitation of this study is its lack of ecological validity, something which had to be sacrificed for any experimental design involving depression and suicide attempts. It is almost certain that participants experienced this study differently than they would if the event actually happened to them. This may have biased results in a pro-signaler direction if the lack of meaningful costs made support feel less costly or if the likely inability of participants to imagine themselves in an anonymous scenario [@hagen2006] led them to take an opportunity to signal generosity cheaply. It may have also biased participants against helping if they could not fully imagine the characters in the story as kin or important cooperative partners, while the survey's short duration may have weakened the strength of the costlier signals as bargaining tools. However, the fact that the relationships among signal cost, belief in need, and likelihood of providing support all largely fit existing theory (narrow and broad) indicates that participants paid attention to the relevant details and responded in ways that suggest some degree of external validity (needs improvement. For this reason, the benefits of addressing causal relationships that came through our experimental vignette approach likely outweighed the costs.

Another limitation is our uncertainty over the data quality from the Indian sample. Compared to the US sample, far more participants were filtered out by our attention checks, something which could be consistent with botting, unfamiliarity with English, or low effort responses [@kennedy2020]. Assuming this corresponded with lower-quality responses among those who passed the attention checks, the differences seen among those in the thwarted marriage vignette may simply be due to added noise rather than differences in the scenarios or the effectiveness of the signals among Indians compared to those in the US. 

# Conclusion

The need to explain suicide's connection to depression is often overlooked by explanations for major depression, particularly those which propose non-signaling functions. One explanation for their connection, is that both suicidality and depression may function to limit negative fitness outcomes in times of substantial adversity [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], with costlier signals being required to elicit support when the payoffs to faking need are high enough so that less costly signals are not reliable indicators of one's need [@Rosenthal1993]. For this hypothesis to be correct, instrumental support would need to be provided frequently enough with large enough impacts so that the costs of signaling are outweighed by the benefits on average. Although possible benefits to depressive states are rarely examined, we found exposure to depression and suicide attempts resulted in greater belief in need and increased likelihood of support in fictional scenarios. This coupled with findings that depression can be simultaneously aversive and elicit support [@HokansonEtAl1986; @stephens1987], suggests that more attention should be paid to the responses of others to depressed individuals, especially from those the depressed are in conflict with and those with the capability to provide meaningful benefits.

# Acknowledgements

This study was funded through a Washington State University, Vancouver research mini grant.

Who to thank:
Lab group?
Arpita?
Committee? 
Anne sounds like she will provide a lot of feedback, so her after that?

# To do:
- add vignettes to SI (Mild depression = crying plus behavior change)
- mention somewhere referring to them all as signals even though we know they provide info as cues too

# Supplementary information

## Pilot study description

Briefly, Participants read a vignette that put them in situations in which an imagined sister was requesting a loan of $50,000 from the participant that was originally saved for the participant's imagined daughter’s college education. The sister claims the money is needed for medical treatment for her own child, but she has incentives to deceive. The manipulations were signals very similar to those used in this study, and the outcomes were belief that the sister really needed the money and intention to loan her the money. See XX for the full write-up. See https://osf.io/3rg8b for preregistration of the pilot study. See http://doi.org/10.5281/zenodo.4637883 for pilot data.

```{r piloteffects, fig.width=10, fig.cap="Pilot study effect of the signals. A. Effect of signal on Belief at T2, controlling for Belief at T1. B: Proportion of participants who inferred the victim was mentally ill. Fit with a generalized linear model with the quasibinomial family."}
pilot_results$effectplot + (pilot_results$mentalplot + theme(axis.text.y = element_blank())) + plot_annotation(tag_levels = 'A')
```


## Power from pilot study

We estimated power using our pilot study data to compare the effect of the *Depression* signal vs. the *Verbal request* signal on *Belief* (that the sister needed the money) at T2 controlling for *Belief* at T1 in a linear regression model. Specifically, we resampled the pilot data with replacement, fit the linear regression model, and extracted the p-value for the signal coefficient, repeating this 2000 times for each sample size ranging from N=20 to N=200. Power was computed as the proportion of p-values < 0.05. Power was about 80% for a sample size of about `r pwr80_90[[2,2]]`, and was about 90% for a sample size of about `r pwr80_90[[2,3]]`. We then repeated this entire procedure to compare *Depression* to a *Control* condition that was meant to not include any signaling (but which we realized, post hoc, did include a verbal request). Power was about 80% for a sample size of about `r pwr80_90[[1,2]]`, and was about 90% for a sample size of about `r pwr80_90[[1,3]]`. See Figure \@ref(fig:power).

```{r power, fig.cap="Power curve for between-subjects effect of Depression vs. Verbal request signal on T2 Belief, controlling for T1 Belief."}
  ggplot(power_curve, aes(sample_size, power, color=Base)) +
    geom_line() +
    labs(x = '\n Sample size', y = 'Power\n') +
    theme_minimal(15)
```

<!-- Second, for participants in the *Depression* condition only, we compared their T2 Belief to their T1 Belief using a paired t-test (in the full pilot data, *Depression* increased belief by about 8 points on a 100 point scale). We again resampled our pilot data 2000 times for each sample size from N=20 to N=200. Power was about 80% for a sample size of 125, and about 90% for a sample size of 175. See Figure \@ref(fig:power2). -->

```{r power2, fig.cap="Power curve for within-subjects effect of Depression signal on T2 Belief relative to T1 Belief."}
  # ggplot(power_curve2, aes(sample_size, power)) +
  #   geom_line() +
  #   labs(x = '\n Sample size', y = 'Power\n') +
  #   theme_minimal(15)
```

## Sample size in each condition in the current study

```{r conditions}
as.data.frame.matrix(xtabs(~signal+vignette, d)) %>% 
  kable(caption = "Sample size in each condition.") %>% 
  kable_styling()
```

<!-- ## Distributions of T2 Belief and Action by signal and vignette -->

```{r distbelief, fig.width=10, fig.height=10, fig.cap="Histograms of T2 Belief by signal and vignette."}

# h <- hist(d$T2Belief, breaks = "FD", plot = FALSE)
# 
# ggplot(d, aes(x = T2Belief, fill=signal)) +
#   geom_histogram(colour='grey', breaks = h$breaks) +
#   scale_x_continuous(breaks = c(0, 0.5, 1)) +
#   facet_grid(signal~vignette, as.table = F) +
#   guides(fill='none') +
#   xlab('\n T2 Belief') +
#   theme_minimal(15) +
#   theme(strip.text.y = element_text(angle=0, hjust=0))
```

```{r distaction, fig.width=10, fig.height=10, fig.cap="Histograms of T2 Action by signal and vignette."}

# h <- hist(d$T2Action, breaks = "FD", plot = FALSE)
# 
# ggplot(d, aes(x = T2Action, fill=signal)) +
#   geom_histogram(colour='grey', breaks=h$breaks) +
#   scale_x_continuous(breaks = c(0, 0.5, 1)) +
#   facet_grid(signal~vignette, as.table = F) +
#   guides(fill='none') +  
#   xlab('\n T2 Action') +
#   theme_minimal(15) + 
#   theme(strip.text.y = element_text(angle=0, hjust=0))
```

## Socioeconomic distribution

```{r socioeconomic, fig.width=8, fig.cap="Distribution of participants by age, log income (USD), and nationality. A small number of participants who did not report income, or reported values < $10 USD, were removed from this plot. A small amount of jitter was added to reveal overlapping points."}

clrs <- viridisLite::magma(11)[c(4, 8)]

d %>% 
  dplyr::filter(!is.na(Age), !is.na(Income), Income >= 10) %>% 
  mutate(
    Nationality = ifelse(sample == 'American', 'US', 'India'),
    logIncome = log10(Income)
  ) %>% 
  ggplot(aes(Age, logIncome, colour = Nationality)) +
  geom_jitter(width = 0.5, alpha = 0.35) +
  geom_hline(yintercept = log10(1670), linetype = 'dotted', colour = clrs[1]) +
  geom_hline(yintercept = log10(35977), linetype = 'dotted', colour = clrs[2]) +
  annotate('text', 15, log10(35977), label = 'Median US income', hjust = 1) +
  annotate('text', 15, log10(1670), label = 'Median India income', hjust = 1) +
  scale_color_binary() +
  guides(colour = guide_legend(reverse = T, override.aes = list(alpha=1))) +
  coord_cartesian(clip = 'off') +
  labs(x = '\nAge (years)', y = 'Log annual\nincome (USD)') +
  theme_minimal(15) +
  theme(axis.title.y = element_text(angle = 0, hjust=0))

```



## Signal effect sizes


```{r effectsizes, fig.width=10, fig.height=7, fig.cap="Between-subjects Cohen's d (95% CI) for the effect of each signal on Belief and Action in each vignette relative to the Verbal request signal, ranked by effect size."}

relbl <- function(lb){
  map_chr(lb, ~str_split(.x, ',')[[1]][2])
}

ggplot(effect_sizes, aes(est, fct_reorder(id, est), xmin=low, xmax=high, colour = signal, shape=Outcome)) +
  geom_pointrange(position = position_dodge(width = 0.7), lwd = 2.5, fatten = 1, alpha=0.5) +
  geom_vline(xintercept = 0, linetype = 'dotted') +
  scale_y_discrete(labels = relbl) +
  scale_color_manual(name=NULL, values = c('darkblue', 'darkred', 'red', 'orange')) +
  # scale_color_viridis_d(option='A') +
  guides(colour = guide_legend(override.aes = list(size = 1)), shape = guide_legend(title=NULL, override.aes = list(size = 1))) +
  # facet_wrap(~Outcome, ncol=1) +
  labs(x="\nCohen's d (95% CI)", y=NULL) +
  theme_minimal(15) +
  theme(legend.position = 'top')
```


```{r effectsizes2, fig.width=10, fig.height=7, fig.cap="Within-subjects Cohen's d (95% CI) for the effect of each signal on Belief and Action in each vignette relative to T1 baseline Belief and  Action, respectively, ranked by effect size."}

ggplot(effect_sizes2, aes(effsize, fct_reorder(id, effsize), xmin=conf.low, xmax=conf.high, colour = signal, shape=Outcome)) +
  geom_pointrange(position = position_dodge(width = 0.7), lwd = 2.5, fatten = 1, alpha=0.5) +
  geom_vline(xintercept = 0, linetype = 'dotted') +
  scale_y_discrete(labels = relbl) +
  scale_color_manual(name=NULL, values = c('black', 'darkblue', 'darkred', 'red', 'orange')) +
  # scale_color_viridis_d(option='A') +
  guides(colour = guide_legend(override.aes = list(size = 1)), shape = guide_legend(title=NULL, override.aes = list(size = 1))) +
  # facet_wrap(~Outcome, ncol=1) +
  labs(x="\nCohen's d (95% CI)", y=NULL) +
  theme_minimal(15) +
  theme(legend.position = 'top')
```


## Demographic effects

```{r demoeffectsaction, fig.width=9, fig.height=8, fig.cap="Demographic effects on Action. **Top**: Association of age with T2 Action, controlling for T1 Action, sex, and vignette. **Bottom**: Association of sex with T2 Action controlling for T1 Action, age, and vignette. Estimated using a generalized linear regression model with the quasibinomial family."}
# effectsPlots$m25plotAge / effectsPlots$m25plotSex
```

```{r demovignettebelief, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with T2 Belief by vignette, controlling for T1 Belief and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
(effectsPlots_US$m18plot / effectsPlots_US$m13plot)
```


```{r demovignetteaction, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with T2 Action by vignette, controlling for T1 Action and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
effectsPlots_US$m19plot / effectsPlots_US$m15plot
```

## Mediation of Suicide attempt signal

The total effect of the *Suicide attempt* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med4$tau.coef, 3)` points from T1 to T2 (on the original 100-point scale). Of this increase, `r 100*signif(med4$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth. See Figure \@ref(fig:mediationsuicide).

```{r mediationsuicide, fig.height=4, fig.cap="The effect of a Suicide attempt (treatment) vs. Verbal request (control) on Action (0-1 scale) is largely mediated by Belief. ACME: Average causal mediation effect. ADE: Average direct effect. The mediation model controlled for T1 Belief, and the outcome model controlled for T1 Belief and T1 Action. Both models were GLMs with the binomial family (the mediation package does not support the quasibinomial family)."}
ggmediation(med4)
```

## Change in beliefs and actions from T1 to T2

```{r rawbeliefs, fig.width=10, fig.height=10, fig.cap="Change in beliefs from T1 to T2 for participants in each condition. Each grey line is one participant. Red lines are the mean change."}
# plot_raw_beliefs
```

```{r rawactions, fig.width=10, fig.height=10, fig.cap="Change in actions from T1 to T2 for participants in each condition. Each grey line is one participant. Red lines are the mean change."}
# plot_raw_actions
```

```{r rawbeliefactions2d, fig.width=10, fig.height=10, fig.cap="Change in beliefs and actions from T1 (arrow base) to T2 arrow head) for participants in each condition. Each grey arrow is one participant. Red arrows are the mean change."}
plot_raw_belief_action2d
```


## T1 emotion vs vignette heatmap

```{r t1emotionheatmap, fig.width=8, fig.cap="Time 1 inference of victim's emotional state in each vignette. Each cell represents the proportion of participants who inferred that emotion. Participants could check multiple emotions from the list. Rows and columns ordered with the PCA angle method in the seriation package."}
hagenheat(dfT1emotion_vignette, rotate_labels = F, seriation_method = 'PCA_angle', viridis_option = 'B') + theme_minimal(15)
```

## T2 emotion vs signal heatmap

```{r t2emotionheatmap, fig.width=8, fig.cap="Time 2 inference of victim's emotional state following the signal. Each cell represents the proportion of participants who inferred that emotion. Participants could check multiple emotions from the list. Rows and columns ordered with the PCA angle method in the seriation package."}

hagenheat(dfT2emotion_signal, rotate_labels = F, seriation_method = 'PCA_angle', viridis_option = 'B') + theme_minimal(15)
```

## Change in emotions T1 to T2 by signal

```{r diffemotions, fig.width=8, fig.cap="Change in inference of victim's emotional state From pre- to post-signal. Each cell represents the difference in proportions of participants who inferred that emotion (T2 - T1). Participants could check multiple emotions from the list. Rows and columns ordered with the PCA angle method in the seriation package."}

d_diff <- as.matrix(dfT2emotion_signal[-1]) - as.matrix(dfT1emotion_signal[-1])
rownames(d_diff) <- dfT2emotion_signal$name

hagenheat(d_diff, rotate_labels = F, seriation_method = 'PCA_angle') + scale_fill_gradient2()

```


## Mentally ill

```{r mentallyillT1, fig.width=8, fig.cap="Proportion of participants in each vignette who thought that the victim was mentally ill at T1. Means estimated with a generalized linear regression model with the quasibinomial family."}
effectsPlots$m27plot + ggtitle('Time 1')
```


```{r mentallyillT2, fig.width=8, fig.cap="Proportion of participants in each condition who thought that the victim was mentally ill at T2, controlling for perceptions of mental illness at T1. Means estimated with a generalized linear regression model with the quasibinomial family."}
effectsPlots$m21plot + ggtitle('Time 2')
```

## The effect of signals on T2 division in thwarted marriage vignette

```{r division, fig.width=9, fig.cap="Effect of signals on division of dowry between the daughters in the thwarted marriage vignette, with an interaction between T1 Division and signal. Means estimated with a generalized linear regression model with the quasibinomial family."}
effectsPlots$m7
```

```{r divisionage, fig.width=9, fig.cap="Interaction of age with signals on T2 Division of dowry between the daughters in the thwarted marriage vignette."}
effectsPlots$m8
```


## T3 Action vs. T2 Action

```{r t2action, fig.width=10, fig.cap="The effect of post-signal (T3) information that the victim was telling the truth on the likelihood of acting to help her (T3 Action), controlling for T2 Action."}
effectsPlots$m26plot + theme(panel.spacing = unit(1.5, "lines"))
```


## Regression tables

```{r regressiontables}
tmp <- models %>% 
  dplyr::filter(Name %in% c('m1', 'm2', 'm3', 'm4', 'm21', 'm24', 'm25'))

regressiontable(tmp$Model, caption = 'Regression coefficients')
```


## ANOVA tables

ANOVA tables for generalized linear regression models (quasibinomial family).

```{r anovatables}

bind_rows(tmp$TidyANOVA, .id = 'Model') %>%
  dplyr::rename(Variable = term, `LR Chisq` = statistic) %>% 
  gt::gt(groupname_col = "Model", caption = "ANOVA tables") %>% 
  gt::cols_label(Variable = "") %>% 
  gt::fmt_number(columns=3, decimals = 1) %>% 
  gt::fmt_scientific(5, decimals = 1) %>% 
  gt::tab_style(style = gt::cell_text(indent = gt::px(40)), 
                locations = gt::cells_body(columns = "Variable"))
```

## Preregistered OLS models

```{r preregbelief, fig.width=8, fig.height=8, fig.cap="The effect of the signals on T2 Belief (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively. Estimated using OLS linear regression models (gaussian)."}
p1 <- effect_plot(fit=models_prereg$Model$m1, xvar='signal',ylbl='T2 Belief', data=d)

p2 <- effect_plot(fit=models_prereg$Model$m3, xvar='signal', ylbl='T2 Action', data=d)

(p1 + theme(axis.text.x = element_blank()))/p2 + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```


```{r preregbeliefB, fig.width=8, fig.height=8, fig.cap="The effect of the signals on T2 Belief (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively, with an interaction between T1 values and the signal. Estimated using OLS linear regression models (gaussian)."}
p1 <- effect_plot(fit=models_prereg$Model$m1b, xvar='signal', by='T1Belief', ylbl='T2 Belief', data=d)

p2 <- effect_plot(fit=models_prereg$Model$m3b, xvar='signal', by='T1Action', ylbl='T2 Action', data=d)

(p1 + theme(axis.text.x = element_blank()))/p2 + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

### OLS regression tables


```{r preregtables}
regressiontable(models_prereg$Model, caption = 'OLS regression coefficients for preregistered linear regression models.')
```

### OLS Anova tables

```{r prepreganova}
bind_rows(models_prereg$TidyANOVA, .id = 'Model') %>%
  dplyr::rename(Variable = term, `LR Chisq` = statistic) %>% 
  gt::gt(groupname_col = "Model", caption = "OLS ANOVA tables") %>% 
  gt::cols_label(Variable = "") %>% 
  gt::fmt_number(columns=c(3,5), decimals = 1) %>% 
  gt::fmt_scientific(6, decimals = 1) %>% 
  gt::fmt_missing(columns=5:6, rows = everything(), missing_text = "---") %>% 
  gt::tab_style(style = gt::cell_text(indent = gt::px(40)), 
                locations = gt::cells_body(columns = "Variable"))

```

## Differences between quasibinomial and bootstrapped confidence intervals

```{r}
m1b_quasi <- models$TidyModel$m1b
m1b_diff <- (m1b_boot[,'boot 97.5%'] - m1b_boot[,'boot 2.5%'])/(m1b_quasi$conf.high-m1b_quasi$conf.low)

m3b_quasi <- models$TidyModel$m3b
m3b_diff <- (m3b_boot[,'boot 97.5%'] - m3b_boot[,'boot 2.5%'])/(m3b_quasi$conf.high-m3b_quasi$conf.low)

diff_sum <- summary(c(m1b_diff, m3b_diff))
```


# References
