---
title: "Depression and suicidality as costly signals of need"
author: "Michael Gaffney, Kai Adams, Kristen L. Syme, and Edward H. Hagen"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
always_allow_html: true
output: 
  bookdown::html_document2:
    number_sections: false
bibliography: MTBib.bib
csl: apa-6th-edition.csl
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, warning = F, error = F, message = F)
# options(warn = 2)
library(equatiomatic)
library(knitr)
library(kableExtra)
library(ggplot2)
library(patchwork)
library(hagenutils)

tar_load(
  c(
    'd',
    'power_curve',
    'models',
    'models_prereg',
    'effect_sizes',
    'effect_sizes2',
    'effectsPlots', 
    'plot_T1Belief_T1Action',
    'plot_emotions',
    'plot_raw_beliefs',
    'plot_raw_actions',
    'med1b',
    'med4',
    'm1b_boot',
    'm3b_boot',
    'dfT1emotion_vignette',
    'dfT1emotion_signal',
    'dfT2emotion_signal'
    )
  )

```

# Introduction
Since Durkheim’s pioneering effort to use quantitative social science to explain suicide, varied attempts by anthropologists, psychologists, and sociologists and substantial investment by funding agencies, rates of depression and suicide attempts have remained largely unchanged over the last 100 years (confirm and cite), and there is still substantial uncertainty about why both are as common as they are (cite). Due to their high costs, the vast majority of these approaches have been built on the assumption that they are the result of internal dysfunction (cite examples). However, a subset of anthropologists and evolutionary theorists have challenged this idea, suggesting that despite the subjective suffering they cause, they may function as a means to limit negative fitness outcomes in times of substantial adversity (Firth + others).

Within anthropology, Raymond Firth was likely the first to do this. Based on observations that suicide attempts often followed social conflict or loss and varied substantially in their likelihood of death, he argued that a sizable subset of the suicide attempts among the Tikopia were not meant to end in death but instead were a means to elicit attention, status, and immediate reintegration into the community following negative events (Firth 1967[not original date]). Compared to suicide, however, anthropological explanations for depression are a much more recent occurrence (cite), despite frequent reports of depression or depression-like behavior in a wide range of societies (cite) and the potential for cross-cultural research to help understand depressive states (cite).

## Connection to Adversity
Likely the most reliable finding when it comes to depression and suicidal behavior is that most cases are closely tied to major negative life events (Hammen, 2005 + others; citations for suicide). Compared to non-depressed individuals, those with depression tend to report at least twice as many instances of negative events in both community and clinical samples (Mazure, 1998) and more negative events than those with schizophrenia and bipolar depression across a wide range of studies (Paykel, 1994). In instances where this is the case, the onset of depression tends to be rapid, often occurring within one month of the event (Kendler et al., 1999 [double check and find others]). 

This strong connection between negative life events and depression along with the need and distress that likely accompanies these events has led to the suggestion that this relationship is causal (Kendler et al., 1999). Support for this comes from the fact that the relationship between negative events and depression holds true when looking solely at events outside of one's control, which shows that the connection is unlikely to be driven solely by depressed individuals creating situations where negative events are likely to be common (Hammen, 2005; Kendler et al., 1999). In addition to this, twin studies have shown that one's history of negative events is still a strong predictor when controlling for genetic similarity and that part of the heritability of depression stems from the heritability of negative events, like divorce and family conflict (Kendler & Baker, 2007; Kendler et al., 1999).

Of the many types of negative life events associated with depression, those which involve social adversity tend to be more strongly associated with depression (Hammen, 2005; Kendler et al., 1999; Mazure, 1998). Like the relationship between negative events as a whole and depression, this relationship is also seen cross-culturally, with the loss of a loved one, threat of divorce, or abusive relationships all increasing one's chances of becoming depressed (Kendler et al., 1999, 1995). Of these events, physical assault has been found to be the strongest predictor of depression in the few studies in which it has been examined (Kendler et al., 1999, 1995) and may have been much more common throughout our evolutionary history than is often thought based on rates of spousal abuse in small-scale societies with little or no policing (Levinson, 1989; Stieglitz et al., 2018). 

Suicide attempts also tend to co-occur with social adversity (Hagen et al., 2008 + cite and describe), with evidence of this existing cross-culturally. Greif over the loss of a loved one (Firth; Devries et al. 2011), bullying (Alavi et al., 2017; Butler et al., 2019; Ford et al., 2017; Wang et al., 2020), and instances of social transgression (Firth; Malinowski; Syme and Hagen 2018; find Western citation to confirm) have all been reported to lead to suicidal behavior in both industrialized (confirm the transgression part) and traditional societies. As with depression, assault is a powerful risk factor for suicidal behavior, something which has been found in multiple large-scale multi-country studies (Devries et al. 2011; Ellsberg et al. 2008). Both sexual and non-sexual assault were also associated with more self-harm and suicidal behaviors among both indigenous Sami and majority Norwegian adolescents, with those who attempted suicides reporting more than twice the number of negative life events than those who did not (Reigstad and Kvernmo 2017). The connection between social adversity and suicide is also seen in the Human Relations Area Files (HRAF), in which 77.4% of attempts were found to be linked to times of social conflict, 88.7% were linked to fitness threats in general, and 84.9% were linked to situations in which attempters were powerless to resolve their situation alone (Syme et al., 2016).

## Depression and suicide attempts as both signals and bargaining tools 
One possible reason for the connection between depression and social adversity is that depression functions to elicit increased support from others in times when an individual cannot solve substantial fitness-relevant challenges on their own. The bargaining model of depression is based on the idea that for much of our evolutionary history the small groups humans lived in created many situations in which certain forms of cooperation were unlikely to be easily replaced by others while also allowing for harmful or unfair social relationships to develop (Hagen, 2003). In these instances, one's monopoly on one or more types of cooperation could be leveraged to change how one is treated if others value the depressed person's cooperation sufficiently.

One way depression and suicidal behavior are thought to lead to beneficial treatment is by reliably demonstrating that one’s need is genuine by making one’s private information about their condition public. This information is necessary due to the ease in which signals can be faked (Zahavi, 1975) and the costs of accepting a deceptive signal as reliable (cite), both of which have favored adaptations to spot fake signals in many species (Hasson, 1994) and demand trustworthy information before believing a given signal (check Hasson and cite + try to find others). 

Although there may be other ways for honest signals evolve (cite), one of the most widely accepted (confirm) means is through costly signaling. The evolution of costly signals is possible whenever the signal in question is comparatively cheaper to produce for those who have the underlying quality compared to those who lack it (Grafen 1990). In these instances, individuals with the quality maximize their fitness at greater levels of investment in the signal compared to those who lack the quality or possess it to a lesser degree (Gangestad and Scheyd 2005; Zahavi 1977). The end result of this is expected to be a signal that either tends to not be worth faking (Gangestad and Scheyd 2005) or is impossible to fake to a large degree due to selection against fakers constraining the physiology behind the signal (Zahavi 1977).

Both depression and suicidal behavior fit this criterion as they are both relatively costly for those not facing adversity than those who are. In the case of depression, the largescale reduction in one's normal routine of fitness enhancing behavior hurts those in need less when one has limited ability to obtain fitness-relevant benefits, with depression likely adding no additional costs in instances of extreme deprivation (Hagen 2003). However, when it comes to suicide, the bulk of the costs associated with suicide attempts are only felt in times that result in death or severe bodily harm. For this reason, the way others view and respond to suicide attempts is likely to be strongly influenced by the probability that the attempt would result in death or severe injury (Rosenthal, 1993). That given, the previously mentioned logic still applies with increased risk of death being relatively costly for those able to enhance their fitness prospects in their day-to-day lives.

Beyond resulting in the cost differential that allows for costly signaling in the first place, the costs of depression and suicidal behavior may be particularly important in times of conflict or when the level of support needed to address one’s situation is substantial. Importantly, there are likely few forms of adversity that do not come with conflict. For example, breakups in both romantic and cooperative relationships are often preceded by conflict (Amato & Hohmann-Marriott, 2007), and surely involve conflict when one side wants reconciliation and the other does not. Likewise, the death of a loved one may also be a source of conflict as the survivor may turn to others for more support who might not want to invest more in the survivor (Hagen, 2003). Regardless of the source of conflict, greater conflict should correspond with greater conflicts of interests, something which would favor greater investment in manipulating others through fake signals. In these instances, less costly signals like verbal requests, facial sadness, and crying are likely to be unreliable as their low costs can easily be made up for with increased social support, thereby favoring the use of costlier signals in those experiencing need they cannot resolve on their own. Importantly, the amount of conflict 

In addition to signaling need to those invested in the well-being of the signaler, depression and suicidal behavior may also help an individual bargain for increased support. This is possible due to the reduced investment in others that frequently accompanies depression or the threat to future cooperation that comes from a suicide. In these instances, depression is suggested to function analogously to a labor strike, sending the message that one’s need is genuine and that if the target of the signal wanted to continue receiving benefits from the depressed, he or she would need to increase support towards the signaler.

## Responses to depression and suicidal behavior 
For depression and suicide attempts to be worth the costs, substantial benefits in terms of social support are expected, either in the form of temporary but impactful support or long-term changes in how one is treated (Hagen, 2003; Hagen et al., 2008; Syme et al., 2016). This can take the form of reconciliations in failing relationships, increased fairness in cooperative relationships, or an increase of resources being directed towards an individual. Of all the benefits, a reduction in physical abuse may be one of the most important due to the threat of death or severe injury to oneself and one’s family (cite).

Most studies examining responses to depression have been observational and focused on responses outside of changes in social support. In general, most report aversive responses and high levels of rejection towards the depressed, and depressed individuals are often linked to impaired social functioning (Evraire & Dozois, 2011; Gotlib & Lee, 1989; Hirschfeld et al., 2000). One explanation for this is that depressed individuals engage in behaviors that drive others away. In line with this reasoning, both excessive reassurance seeking and negative feedback seeking, which often accompany depression, have been shown to elicit negative emotional responses from others and increase the chance of rejection from others (Joiner & Metalsky, 1995; Starr & Davila, 2008). Negative interactions are also seen in families as depressed children tend to report less support and more negative responses from parents than non-depressed children, although it has not been demonstrated that such interactions did not lead to depression in the first place (Messer & Gross, 1995; Schwartz et al., 2012).

Despite being looked for less frequently, beneficial responses to depressed individuals are also reported. This is seen in that depression has been shown to reduce the likelihood of aggression within families (Hops 1987[read]; Dadds et al. 1982 [read]), particularly in families already experiencing high levels of conflict (Sheeber et al., 2001). There is also limited evidence that depression can result in increased levels of support from others outside the family. For example, non-depressed individuals were shown to be more cooperative and exhibit more caretaking behavior towards their college depressed roommates over the course of their time living together (Hokanson et al., 1986). Likewise, a study examining responses to confederates trained to appear ‘depressed’, ‘anxious’, or ‘normal’ found that individuals were more likely to provide advice and support to ‘depressed’ individuals than ‘anxious’ or ‘normal’ individuals despite the confederate in ‘depression’ condition being viewed more negatively than the other conditions and the ‘depression’ condition involving similar levels of distress compared to the ‘anxious’ condition (Stephens et al., 1987 [tweaked based on your paper with Kristen]). 

Unfortunately, there are few studies examining how individuals respond to suicide attempt survivors, with the majority examining responses after death. In those which exist, stigmatization is commonly reported within and outside families (Frey et al., 2017; Scocco et al., 2012), with perceptions of survivors as being weak, selfish, mentally ill, and antisocial all being commonly reported (Batterham et al., 2013; Frey et al., 2016a; Tzeng & Lipson, 2004). Although these types of stigmatization have been reported across relationships types, the likelihood of stigma has been found to differ based on who the survivor is to the person being questioned. This is seen in a study which found that individuals felt the most stigma from members of their preexisting social networks than mental health providers, with family members providing more stigmatization (57.1% of the time) than romantic partners (41.2% of the time) and close friends (28.2% of the time) (Frey et al., 2016b). 

Although negative responses to suicide survivors are more often studied and reported, positive responses have also been found. Increased social support and changes to important relationships are often reported to follow suicide attempts with some indication that these effects may hold long term (Stengel, 1956). For example, a study of 100 women who survived suicide attempts found that in 75 cases individuals gained identifiable benefits through the attempt, with 41 individuals benefiting from reconciliations with others (Lukianowicz, 1971). As for the type of support received, emotional support seems to be more commonly provided than instrumental support based on initial evidence (Frey et al., 2017); however, it is currently unclear how this might change with social context.

Beneficial social changes are also reported in traditional societies. Unlike Western countries, where suicide is often viewed as pathological (Hidaka, 2012), members of traditional societies have been reported to view suicide attempts as cries for help rather than mental illness (Shostak, 1981). For example, attempts have been described as ways of escaping unwanted marriage arrangements, persistent abuse, or a lack of support in obtaining mates by those engaging in the behavior and observers (Gutierrez de Pineda & Muirden, 1948; Hilger, 1957; Karsten, 1935; Tessmann, 1930; Wilson, 1960), suggesting that help is at least occasionally provided or reliably expected. In support of this, 30 out of 84 examples of suicidal behavior included in the HRAF resulted in positive changes for the survivor, even though suicide was rarely if ever the main topic (Syme et al., 2016).

# Materials and methods

## notes: 
add funding info somewhere
update MTurk (Ed email links + mine)
Find a place for the predictions and preregistration 
go back and emohasize conflit and PI in the methods (at least in the pilot studies portion)

## Design

This study involved a vignette study with between subjects pre-test/post-test design with one factor (signal) and 5 levels (verbal request, crying, crying with behavioral changes, depression, and suicide attempt), with the outcome of interest being how believe and likelihood of helping changed in response to a given signal. 

## Pilot Studies

The first pilot study used a vignette study to examine how the presence of conflict, private information, and emotional signaling impacted the degree participants believed the signaler to be in need and provide support using an MTurk sample of 1,447 Americans. Of the signals, depression resulted in the most pro-signaler responses (reword), separating itself from the less costly signals when conflict and private information was present and substantially outperforming the schizophrenic break control. However, the suicide attempt condition was responded to relatively poorly, contrary to predictions.

One potential explanation is that the initial scenario of a mother asking their sibling for help paying for their young daughter's heart surgery was not severe enough for a suicide attempt to be a functional signal, due to a lack of beleivability or due to its costs appearing manipulative. For this reason, we designed this study with more severe scenarios that were also associated with suicide attempts both ethnographically and clinically, including instances of sexual assasult, non-sexual assault, and thwarted marriage.


add that we ask now ask same question twice? (interpretabilty)

The second was posted on the samplesize subreddit with the goals of testing the beleivability of the new vignettes and how pro-signaler participants were at T1, with a power analysis with simulated data showing this would limit our power. In this small sample of 28, we saw no evidence that beleivabilty was a problem and no evidence that participants as a whole would be too pro-sister at T1. 

## Vignettes

Four vignettes were used in this study, with participants in the US sample being randomly assigned to either the "basketball coach," "romantic partner," or "brother-in-law" vignettes. In the "basketball coach" scenario, the star player on a unversity basketball team comes to the participant (who is asked to imagine themselves as the athletic director at a university) and claims she was sexually assaulted by the coach. However, there are indications the player might be lying to get more playing time.

The "brother-in-law" vignette involved participants imagining that they let their sister, brother-in-law, and niece move in with the participant and an imagined 15-year-old daughter after they lost their house in a fire. During this time, the participant's daughter becomes jealous of your niece before accusing her uncle of sexually assaulting her.

The "romantic partner" vignette hard participants imagine that they found the best partner they have ever had after years of being single. However, their 13-year-old daughter, who has a history of trying to get in the way of their relationships is clearly not happy with this. At one point, she accuses the romantic partner of physically assaulting her but cannot produce any evidence.

Those in the Indian sample only saw the "thwarted marriage vignette." In this scenario, the participant and their family is trying to arrange a dowry for their older daughter so she can marry the man she loves. After his family demands more money, the participant's family tries to find another man, who the older daughter claims to find unattractive. However, it is unclear how unattractive he is and any additional investment in the dowry will come at the expense of the younger daughter.

The full vignettes are available in the supplementary information.

## Signals

The sole treatment in this study was the signal participants saw, with these signals including: (1) a verbal request without additional signaling, (2) crying, (3) mild-depression, (4) depression and, (5) a suicide attempt.

As the amount of crying in those with depression is moderately associated with depression severity [@VingerhoetsEtAl2016] and depression is a major risk factor for suicide [@BostwickPankratz2000; @Kessler2012], each signal was designed to build off of the less costly signals so that each signal contained elements of any signal it was costlier than.

## Preregistration

https://osf.io/g3s6n

## Sampling

Participants for this study were recruited from Amazon Mechanical Turk, a crowdsourcing platform that allows for the creation of Human Intelligence Tasks (HITs) that workers can complete for pay. As Amazon provides the infrastructure, it allows for a relatively low-cost way of collecting data for academic research at the cost of the data not being representative of any real population. Despite this limitation, it is more diverse than most convenience samples [@KennedyEtAl2018; @DworkinEtAl2016], with turkers in the U.S. mainly differing in being younger, more educated, lower income, and more female biased than the greater U.S. population [@RossEtAl2010]. 
	
Overall, the quality of data provided by MTurk workers tends to resemble that of university sample pools, with there being some indication that it might outperform university samples in a few important areas. This is seen in that while both MTurk workers and university students engage in problematic behaviors that might lessen data quality to similar amounts [@NeckaEtAl2016], workers on MTurk have been shown to be more attentive than samples of university students [@HauserSchwarz2016]. For this reason, concerns about data quality come primarily from potential bot use and respondents faking their location to take surveys in a language they do not understand well, with the former concern being largely unfounded and the latter able to be corrected for with the tools used in this study [@KennedyEtAl2018]
	
## Sample

All participants in the US sample were over 18, located in the United States, had a HIT approval rate of over 98%, and had successfully completed at least 100 HITs as anything less results in approval rates of 100% [@KennedyEtAl2018]. The Indian sample was the same except for only including participants who resided in India at the time of the study and the inclusion two participants who were included with a stricter HIT approval rate of over 99%.

To ensure participants were paying attention, we used a variety of attention checks, which we clearly labeled in an attempt to avoid demand characteristics. The first was shown immediately after the consent form and provided participants with a random word and asked them to enter the vowels in the order in which they are found in the word. Capitalization did not matter and no word contained repeat vowels. Failure to enter the correct vowels in order resulted in the participants being filtered out of the study. The second followed the vignette and involved asking three questions about the story that should be easy to answer for anyone paying attention. 

We also filtered out those who read the vignette faster than one-third the time it took MG to read it as fast as possible without concern for comprehension immediately after the vignette. 

Informed consent was given by proceeding with the study after reading the consent form, and all participants who passed the attention checks were paid $1 for their time. (mention time here)

## Survey

After accepting the HIT on MTurk, every participant took the survey using Qualtrics. In every instance, the elements of the survey were laid out in the following order: 1) the consent form, 2) the vowel attention check, 3) the main body of the vignette, 3) the pre-test baseline measures (T1), 4) the story content attention check, 5)  the signal portion of the vignette, 7) the post-test outcome measures (T2), 8) an attention check that did not filter out participants, 9) a set of demographic questions, and 10) an open feedback prompt. The full survey can be found in the SI.

### Baseline measures (T1)

In both samples, we recorded self-reports on how much participants believed the signaler to be telling the truth (BeliefT1) and the likelihood of them helping the signaler as requestedc (ActionT1). With the thwarted marriage vignette we also asked how they would split the money they had saved for the dowry between their daughters (Divide T1). In every instance, the order of the questions was randomized to avoid order effects [@KrosnickAlwin1987].

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. In every instance the position started at the left due to findings that people may be less likely to move the slider when it starts in the center (cite and word better). For the exact wording of the question and the ends of each continuum, see table 1.

We also asked which emotions participants felt the signaler was experiencing using a multiple choice question in which they could select multiple options. This included emotions directly related to the signals (e.g., sad, depressed, and suicidal), states which suggest genuine need (traumatized and violated), and states which suggest deception (e.g., deviousness or jealousy). The complete list can be found in the SI.

### Post-treatment measures (T2-T4)

We asked questions identical to those asked at T1 as the main post-treatment variables of interest (T2). For the belief, action, and divide variables, the position of the slider starting where participants placed it at T1. For the T2 emotion variables this involved no changes from the T1 version. 

As both a validity check and a way to understand the degree participants would be willing to help if they believed the participant completely, we presented participants with strong evidence that the claims were true (Action T3). In the US sample, this involved video evidence of the event in question occuring or a similar event after the fact. With the thwarted marriage vignette, this involved the participant seeing the man their daughter wants to marry trash talk their daughter and their family. As a final attention check, we presented participants with the Action T3 question but included instructions to ignore this and drag the slider all the way to the left (T4AttentionCheck). We did not exclude anyone from the study for failing this nor did we withhold payment.

Variable | Vignette | Wording | Coding 
---------|----------|---------|-------
T1Belief | Basketball | How much do you believe the star player is telling the truth about the coach? | 0: Not at all; 100: Completely
T1Belief | Brother-in-law | How much do you believe your daughter is telling the truth about your brother-in-law? | 0: Not at all; 100: Completely
T1Belief | Romantic partner | How much do you believe your daughter is telling the truth about your partner? | 0: Not at all; 100: Completely
T1Belief | Thwarted Marriage | How much do you believe your older(bold) daughter is telling the truth about the second man not being a good match for her?  | 0: Not at all; 100: Completely
T1Action | Basketball | How likely are you to recommend that the president of the university should suspend the coach? | 0: Very unlikely; 100: Very likely
T1Action | Brother-in-law | How likely are you to kick your brother-in-law out? | 0: Very unlikely; 100: Very likely
T1Action | Romantic partner | How likely are you to kick your partner out? | 0: Very unlikely; 100: Very likely
T1Action | Thwarted Marriage | How likely are you to provide the large dowry so your older(bold) daughter can marry the man she is in love with?  | 0: Very unlikely; 100: Very likely
T1Divide | Thwarted Marriage | How would you plan on dividing the money between your daughters?  | 0: Much more for the younger daughter; 50: Even division; 100: Much more for the older daughter

Table: (\#tab:outcomes) Outcome variables in each vignette.

### Demographic Questions

The final part of the survey was a brief demographic questionnaire which examined: (1) the age, (2) sex, (3) number of siblings, (4) number of sons, (5) number of daughters, (6) current relationship status, (7) highest level of education, and (8) the annual household income of each participant.

## Confirmatory Analyses

We used linear regression to analyze our results for the main effect of signal on belief in need and likelihood of help, controlling for baseline levels of belief in need and likelihood of help. (Need more detail)

Belief in need t2 ~ Signal + Belief in need t1
Likelihood of help t2 ~ Signal + Likelihood of help t1

Based on advice from our statistical consultant, we also fit the following interaction models:

Belief in need t2 ~ Signal * Belief in need t1
Likelihood of help t2 ~ Signal * Likelihood of help t1

The rationale for the interaction is that for individuals who indicate high levels of belief and helping at baseline, the signals might have reduced effect.

To examine our hypothesis that beneficial treatment would be largely due to a signal's positive impact on the participant's beleif in need we ran the following mediaton models for each pairwise combination of signals: 

Mediator model: Belief in need t2 ~ Signal + Belief in need t1
Outcome model: Likelihood of help t2  ~ Signal + Belief in need t2 + Likelihood of help t1 + Belief in need t1 

# Results

## Summary statistics

```{r sumtab}

# d$Income <- ifelse(d$CurrencyType == 'USD', d$Income, d$Income*0.014)

d$minutes <- d$TotalTime/60

vars <- c(
  Age = 'Age (years)',
  Income = 'Income (USD)',
  minutes = 'Time to complete (minutes)',
  T1Belief = 'T1 Belief',
  T1Action = 'T1 Action',
  T1Divide = 'T1 Division',
  T1LowMood = 'T1 Low mood',
  T1Manipulative = 'T1 Manipulative',
  T2Belief = 'T2 Belief',
  T2Action = 'T2 Action',
  T2Divide = 'T2 Division',
  T2LowMood = 'T2 Low mood',
  T2Manipulative = 'T2 Manipulative',
  T3Action = 'T3 Action'
)
kable(custom.summarize(d, vars), caption = "Summary statistics for study variables. Original 0-100 values scaled to 0-1. Indian participants reported income in rupees, which we converted to USD at the current exchange rate (1 rupee == 0.014 USD).") %>% kable_styling()
```

## Distributions of beliefs and actions at time 1

```{r t1plot, fig.width=10, fig.height=10, fig.cap="The distributions of Time 1 Belief vs. Time 1 Action, by vignette. Each dot is one participant. Original 0-100 values scaled to 0-1. Contour lines represent equal densities, estimated using a bivariate normal kernal (Venables and Ripley, 2002). Lines fit by linear regression. Dot size indicates the number of overlapping points."}
plot_T1Belief_T1Action
```

## Effects of signals on beliefs and actions

```{r beliefeffect, fig.width=8, fig.height=8, fig.cap="The effect of the signals on T2 Belief (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively, with an interaction between T1 values and the signal. Estimated using generalized linear regression models with the quasibinomial family."}
(effectsPlots$m1bplot + theme(axis.text.x = element_blank()))/effectsPlots$m3bplot + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

```{r beliefvignette, fig.width=8, fig.height=8, fig.cap=paste0("The effect of the signals on T2 Belief (top) and Action (bottom) in each vignette, controlling for T1 values of Belief and Action, respectively. Effects computed with T1 values set to their median value (T1Belief=", median(d$T1Belief, na.rm=T), ", T1Action=", median(d$T1Action, na.rm=T), "). Estimated using generalized linear regression models with the quasibinomial family.")}
(effectsPlots$m2plot + theme(axis.text.x = element_blank()))/effectsPlots$m4plot + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

## The effect of signals on action is largely mediated by belief

According to our theoretical model, credible signals of need increase observers' belief that the victim is telling the truth. Increased belief then increases the likelihood of helping the victim (action). See Figure \@ref(fig:medmodel).

```{r medmodel, fig.cap="Possible causal effects of the signal on helping behavior. A credible signal of need increases belief that the victim is telling the truth and needs help, which increase the likelihood of helping. The direct path from the signal to action represents other causal effects of the signal on observers, such as perceptions of the victim's emotional state (and other factors not measured in this study) that might alter observer behavior."}

knitr::include_graphics('paths.svg')
```

The total effect of the *Depression* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med1b$tau.coef, 3)` points from T1 to T2 (on the original 100-point scale). Of this increase, `r 100*signif(med1b$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth. See Figure \@ref(fig:mediation).

```{r mediation, fig.height=4, fig.cap="The effect of Depression (treatment) vs. Verbal request (control) on Action (0-1 scale) is largely mediated by Belief. ACME: Average causal mediation effect. ADE: Average direct effect. The mediation model controlled for T1 Belief, and the outcome model controlled for T1 Belief and T1 Action. Both models were GLMs with the binomial family (the mediation package does not support the quasibinomial family)."}
ggmediation(med1b)
```

## Effects of signals on victim's perceived emotional state

```{r emotionsplot, fig.width=10, fig.height=8, fig.cap="The effect of the signals on participants' inferred *manipulative* and *low mood* emotional states of the victim, from T1 (arrow base) to T2 (arrow head), in each of the four vignettes. Thick arrows: mean change. Thin arrows: means of 500 bootstrap resamples."}
plot_emotions
```

# Discussion

## The effect of signaling on belief and likelihood of support

Consistent with costly signaling theory, costlier signals resulted in greater perceptions of need and increased the likelihood of participants helping the signaler than the less costly signals. This effect was largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters. However, the effect of signal type was much smaller in both the "basketball coach" and the "thwarted marriage" vignette. In the "basketball coach" vignette, part of the reason for this may have been the lack of relatedness between participants and the alleged victim. However, this was still a surprising finding due to the relatively small size of the request (i.e., recommend the coach be suspended rather than kick out family or a romantic partner). With the thwarted marriage vignette, the fact that support came at the cost of one's younger daughter may explain the relatively small effect of signal strength. However, data from the Indian sample appears to be of lower quality, limiting our confidence in this interpretation (see limitations for more information).

In the "brother-in-law" and "romantic partner" scenarios, there was a noticeable jump in the magnitude of pro-signaler responses when going from the crying to mild depression and then a second, smaller jump from depression to a suicide attempt. One potential reason for this is the increase in symptoms from mild depression to depression was not costly enough to result in a large effect as the symptoms only differed in severity (e.g., grades dropping from As to Bs [mild depression] or Cs [depression]). Such small changes may be less impactful in vignette form than in real-life scenarios in which the effects of signaling may increase in severity and be prolonged over a lenghty period of time.

We also found evidence for the predicted interaction between signal and belief in need and likelihood of help at T1, with the pro-signaler effects being larger at intermediate but low levels of believe at T1. In real-life scenarios, it would be reasonable to expect that costly-signaling would not be engaged in when others already believe one's need. For this reason, too strong of a signal may be seen as manipulative or a cue of dysfunction, and therefore, be responded to aversively. However, belief in need and likelihood of help at T2 still remained high when it was high at T1, suggesting that this may not be a significant risk when one is in severe enough need (want to bring in MA here. Is that a bad idea?).

## Mediation models

Our prediction that the effect of signal cost on likelihood of help would largely be mediated by signal's effect on belief in need was also met, as would be expected if the main function of signaling need is a reduction in the amount of private information one has relative to those motivated to care for the signaler (cite). However, there was still a positive main effect of the signal in many pairwise comparisons between the costlier signals and the less costly signals, including the comparison between the depression and verbal request condition (See SI for other mediation models). One possible reason for this is that someone may still be distressed and face real adversity without being forthcoming about one's problems. For example, participants who saw the "brother-in-law" vignette may have thought the daughter was lying about the assault due to jealously over her cousin, but the stronger signals may have still conveyed the severity of the competition between her and her cousin in ways words alone could not.

## Broader discussion

Our results on the connection between belief in need and likelihood of help are consistent with the bargaining models of depression and suicide. In addition to the effect of signal cost on support being largely mediated by the signal's effect on belief in need, belief in need and likelihood of support were highly correlated but distinct at T1 (pre-signal). This situation should arise in any instance when a target of a signal of need would benefit from helping if the need was genuine and known to reasonable precision. Importantly, such problems are common among humans today (cite) and have likely been so since individuals could increase their inclusive fitness through supporting kin or important cooperative partners in times of need or conflict over acceptable one's contributions joint endeavors (cite).

Although this study was not designed to examine whether the costs of signaling would be made up for by the increased support stemming from depression and suicide attempts directly, the results are encouraging. In general, the costs required to honestly convey one's qualities should be expected to increase with the potential payoffs of deception (cite), with the vignettes in this study emphasizing the existence of private information and sizable conflicts of interest to increase the perceived risk of being taken advantage of. In similar real-world scenarios, this difficulty of signaling honest need would coincide with the substantial fitness-relevant benefits of preventing future assaults or being able to marry the partner of one's choice. These benefits coupled with the fact (too strong?) that depression and suicide attempts become relatively less costly with increased need (Hagen 2003), suggest that the increased support seen in this study may be worth the costs required to signal need more often than is often appreciated when counting the entire costs of depression as the signal rather than only the upregulation of symptoms for a communicative purpose.

## Limitations

The largest source of limitations of this study is its lack of ecological validity, something which had to be sacrificed for any experimental design involving depression and suicide attempts. It is almost certain that participants experienced this study differently than they would if the event actually happened to them. This may have biased results in a pro-signaler direction due to the lack of meaningful costs making support feel less costly or through the likely inability of participants to imagine themselves in an anonymous scenario (cite your paper on this) leading them to take an opportunity to signal generosity cheaply. It may have also biased participants against helping if they could not fully imagine the characters in the story as kin or important cooperative partners, while the survey's short duration may have weakened the strength of the bargaining of the costlier signals. However, both the findings of a main effect of signal and the effect of signal type on likelihood of support was largely mediated by change in belief suggest that participants were paying attention to the theoretically relevant details in ways that influenced their behavior. 

Another limitation is our uncertainty over the data quality from the Indian sample. Compared to the US sample, far more participants were filtered out by our attention checks, something which could be consistent with botting, unfamiliarity with English, or low effort responses. Assuming this corresponded with more low-quality responses among those who passed the attention checks, the differences seen among those in the thwarted marriage vignette may simply be due to added noise rather than differences in the scenarios or the effectiveness of the signals among Indians compared to Americans.

# Conclusion

# Supplementary information

## Pilot study description

Briefly, Participants read a vignette that put them in situations in which an imagined sister was requesting a loan of $50,000 from the participant that was originally saved for the participant's imagined daughter’s college education. The sister claims the money is needed for medical treatment for her own child, but she has incentives to deceive. The manipulations were signals very similar to those used in this study, and the outcomes were belief that the sister really needed the money and intention to load her the money. See XX for the full write-up. See https://osf.io/3rg8b for preregistration of the pilot study.

## Power from pilot study

We estimated power using our pilot study data by comparing the effect of the *Depression* signal vs. the *Verbal request* signal on *Belief* (that the sister needed the money) at T2 controlling for *Belief* at T1 in a linear regression model. Specifically, we resampled the pilot data with replacement with sample sizes ranging from N=20 to N=100, fit the linear regression model, and extracted the p-value for the signal coefficient, repeating this 1000 times for each sample size. Power was computed as the proportion of p-values < 0.05. Power was about 80% for a sample size of about 35, and was about 90% for a sample size of about 45. 

```{r power}
  ggplot(power_curve, aes(sample_size, power)) +
    geom_line() +
    labs(x = '\n Sample size (2 conditions)', y = 'Power\n') +
    theme_minimal(15)
```


## Sample size in each condition in the current study

```{r conditions}
as.data.frame.matrix(xtabs(~signal+vignette, d)) %>% 
  kable(caption = "Sample size in each condition.") %>% 
  kable_styling()
```

## Distributions of T2 Belief and Action by signal and vignette

```{r distbelief, fig.width=10, fig.height=10, fig.cap="Histograms of T2 Belief by signal and vignette."}
ggplot(d, aes(x = T2Belief, fill=signal)) +
  geom_histogram(colour='grey') +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  facet_grid(vignette~signal) +
  guides(fill='none') +
  xlab('\n T2 Belief') +
  theme_minimal(15) + 
  theme(strip.text.y = element_text(angle=0, hjust=0))
```

```{r distaction, fig.width=10, fig.height=10, fig.cap="Histograms of T2 Action by signal and vignette."}
ggplot(d, aes(x = T2Action, fill=signal)) +
  geom_histogram(colour='grey') +
  scale_x_continuous(breaks = c(0, 0.5, 1)) +
  facet_grid(vignette~signal) +
  guides(fill='none') +  
  xlab('\n T2 Action') +
  theme_minimal(15) + 
  theme(strip.text.y = element_text(angle=0, hjust=0))
```

## Signal effect sizes

```{r effectsizes2, fig.width=10, fig.height=7, fig.cap="Within-subjects Cohen's d (95% CI) for the effect of each signal on Belief and Action in each vignette relative to T1 baseline Belief and  Action, respectively, ranked by effect size."}

relbl <- function(lb){
  map_chr(lb, ~str_split(.x, ',')[[1]][2])
}

ggplot(effect_sizes2, aes(effsize, fct_reorder(id, effsize), xmin=conf.low, xmax=conf.high, colour = signal, shape=Outcome)) +
  geom_pointrange(position = position_dodge(width = 0.7), lwd = 2.5, fatten = 1, alpha=0.5) +
  geom_vline(xintercept = 0, linetype = 'dotted') +
  scale_y_discrete(labels = relbl) +
  scale_color_manual(name=NULL, values = c('black', 'darkblue', 'darkred', 'red', 'orange')) +
  # scale_color_viridis_d(option='A') +
  guides(colour = guide_legend(override.aes = list(size = 1)), shape = guide_legend(title=NULL, override.aes = list(size = 1))) +
  # facet_wrap(~Outcome, ncol=1) +
  labs(x="\nCohn's d (95% CI)", y=NULL) +
  theme_minimal(15) +
  theme(legend.position = 'top')
```


```{r effectsizes, fig.width=10, fig.height=7, fig.cap="Between-subjects Cohen's d (95% CI) for the effect of each signal on Belief and Action in each vignette relative to the Verbal request signal, ranked by effect size."}

ggplot(effect_sizes, aes(est, fct_reorder(id, est), xmin=low, xmax=high, colour = signal, shape=Outcome)) +
  geom_pointrange(position = position_dodge(width = 0.7), lwd = 2.5, fatten = 1, alpha=0.5) +
  geom_vline(xintercept = 0, linetype = 'dotted') +
  scale_y_discrete(labels = relbl) +
  scale_color_manual(name=NULL, values = c('darkblue', 'darkred', 'red', 'orange')) +
  # scale_color_viridis_d(option='A') +
  guides(colour = guide_legend(override.aes = list(size = 1)), shape = guide_legend(title=NULL, override.aes = list(size = 1))) +
  # facet_wrap(~Outcome, ncol=1) +
  labs(x="\nCohn's d (95% CI)", y=NULL) +
  theme_minimal(15) +
  theme(legend.position = 'top')
```

## Demographic effects

```{r demoeffectsbeliefs, fig.width=9, fig.height=8, fig.cap="Demographic effects on Belief. **Top**: Association of age with T2 Belief, controlling for T1 Belief, sex, and vignette. **Bottom**: Sex interacting with the effect of the signals on T2 Belief controlling for T1 Belief, age, and vignette. Estimated using a generalized linear regression model with the quasibinomial family."}
effectsPlots$m24plotAge / effectsPlots$m24plotSex
```

```{r demoeffectsaction, fig.width=9, fig.height=8, fig.cap="Demographic effects on Action. **Top**: Association of age with T2 Action, controlling for T1 Action, sex, and vignette. **Bottom**: Association of sex with T2 Action controlling for T1 Action, age, and vignette. Estimated using a generalized linear regression model with the quasibinomial family."}
effectsPlots$m25plotAge / effectsPlots$m25plotSex
```

```{r demovignettebelief, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with T2 Belief by vignette, controlling for T1 Belief and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
effectsPlots$m18plot / effectsPlots$m13plot
```


```{r demovignetteaction, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with T2 Action by vignette, controlling for T1 Action and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
effectsPlots$m19plot / effectsPlots$m15plot
```

## Mediation of Suicide attempt signal

The total effect of the *Suicide attempt* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med4$tau.coef, 3)` points from T1 to T2 (on the original 100-point scale). Of this increase, `r 100*signif(med4$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth. See Figure \@ref(fig:mediationsuicide).

```{r mediationsuicide, fig.height=4, fig.cap="The effect of a Suicide attempt (treatment) vs. Verbal request (control) on Action (0-1 scale) is largely mediated by Belief. ACME: Average causal mediation effect. ADE: Average direct effect. The mediation model controlled for T1 Belief, and the outcome model controlled for T1 Belief and T1 Action. Both models were GLMs with the binomial family (the mediation package does not support the quasibinomial family)."}
ggmediation(med4)
```

## Raw beliefs and actions

```{r rawbeliefs, fig.width=10, fig.height=10, fig.cap="Change in beliefs from T1 to T2 for participants in each condition. Each grey line is one participant. Red lines are the mean change."}
plot_raw_beliefs
```

```{r rawactions, fig.width=10, fig.height=10, fig.cap="Change in actions from T1 to T2 for participants in each condition. Each grey line is one participant. Red lines are the mean change."}
plot_raw_actions
```

## T1 emotion vs vignette heatmap

```{r t1emotionheatmap, fig.width=8, fig.cap="Time 1 inference of victim's emotional state in each vignette. Each cell represents the proportion of participants who inferred that emotion. Participants could check multiple emotions from the list. Rows and columns ordered with the PCA angle method in the seriation package."}
hagenheat(dfT1emotion_vignette, rotate_labels = F, seriation_method = 'PCA_angle', viridis_option = 'B') + theme_minimal(15)
```

## T2 emotion vs signal heatmap

```{r t2emotionheatmap, fig.width=8, fig.cap="Time 2 inference of victim's emotional state following the signal. Each cell represents the proportion of participants who inferred that emotion. Participants could check multiple emotions from the list. Rows and columns ordered with the PCA angle method in the seriation package."}

hagenheat(dfT2emotion_signal, rotate_labels = F, seriation_method = 'PCA_angle', viridis_option = 'B') + theme_minimal(15)
```

## Change in emotions T1 to T2 by signal

```{r diffemotions, fig.width=8, fig.cap="Change in inference of victim's emotional state From pre- to post-signal. Each cell represents the difference in proportions of participants who inferred that emotion (T2 - T1). Participants could check multiple emotions from the list. Rows and columns ordered with the PCA angle method in the seriation package."}

d_diff <- as.matrix(dfT2emotion_signal[-1]) - as.matrix(dfT1emotion_signal[-1])
rownames(d_diff) <- dfT2emotion_signal$name

hagenheat(d_diff, rotate_labels = F, seriation_method = 'PCA_angle') + scale_fill_gradient2()

```


## Mentally ill

```{r mentallyillT1, fig.width=8, fig.cap="Proportion of participants in each vignette who thought that the victim was mentally ill at T1. Means estimated with a generalized linear regression model with the quasibinomial family."}
effectsPlots$m27plot + ggtitle('Time 1')
```


```{r mentallyillT2, fig.width=8, fig.cap="Proportion of participants in each condition who thought that the victim was mentally ill at T2, controlling for perceptions of mental illness at T1. Means estimated with a generalized linear regression model with the quasibinomial family."}
effectsPlots$m21plot + ggtitle('Time 2')
```

## The effect of signals on T2 division in thwarted marriage vignette

```{r division, fig.width=9, fig.cap="Effect of signals on division of dowry between the daughters in the thwarted marriage vignette, with an interaction between T1 Division and signal. Means estimated with a generalized linear regression model with the quasibinomial family."}
effectsPlots$m7
```

```{r divisionage, fig.width=9, fig.cap="Interaction of age with signals on T2 Division of dowry between the daughters in the thwarted marriage vignette."}
effectsPlots$m8
```


## T3 Action vs. T2 Action

```{r t2action, fig.width=10, fig.cap="The effect of post-signal (T3) information that the victim was telling the truth on the likelihood of acting to help her (T3 Action), controlling for T2 Action."}
effectsPlots$m26plot + theme(panel.spacing = unit(1.5, "lines"))
```


## Regression tables

```{r regressiontables}
tmp <- models %>% 
  dplyr::filter(Name %in% c('m1', 'm2', 'm3', 'm4', 'm21', 'm24', 'm25'))

regressiontable(tmp$Model, caption = 'Regression coefficients')
```


## ANOVA tables

ANOVA tables for generalized linear regression models (quasibinomial family).

```{r anovatables}

bind_rows(tmp$TidyANOVA, .id = 'Model') %>%
  dplyr::rename(Variable = term, `LR Chisq` = statistic) %>% 
  gt::gt(groupname_col = "Model", caption = "ANOVA tables") %>% 
  gt::cols_label(Variable = "") %>% 
  gt::fmt_number(columns=3, decimals = 1) %>% 
  gt::fmt_scientific(5, decimals = 1) %>% 
  gt::tab_style(style = gt::cell_text(indent = gt::px(40)), 
                locations = gt::cells_body(columns = "Variable"))
```

## Preregistered OLS models

```{r preregbelief, fig.width=8, fig.height=8, fig.cap="The effect of the signals on T2 Belief (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively. Estimated using OLS linear regression models (gaussian)."}
p1 <- effect_plot(fit=models_prereg$Model$m1, xvar='signal',ylbl='T2 Belief', data=d)

p2 <- effect_plot(fit=models_prereg$Model$m3, xvar='signal', ylbl='T2 Action', data=d)

(p1 + theme(axis.text.x = element_blank()))/p2 + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```


```{r preregbeliefB, fig.width=8, fig.height=8, fig.cap="The effect of the signals on T2 Belief (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively, with an interaction between T1 values and the signal. Estimated using OLS linear regression models (gaussian)."}
p1 <- effect_plot(fit=models_prereg$Model$m1b, xvar='signal', by='T1Belief', ylbl='T2 Belief', data=d)

p2 <- effect_plot(fit=models_prereg$Model$m3b, xvar='signal', by='T1Action', ylbl='T2 Action', data=d)

(p1 + theme(axis.text.x = element_blank()))/p2 + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

### OLS regression tables


```{r preregtables}
regressiontable(models_prereg$Model, caption = 'OLS regression coefficients for preregistered linear regression models.')
```

### OLS Anova tables

```{r prepreganova}
bind_rows(models_prereg$TidyANOVA, .id = 'Model') %>%
  dplyr::rename(Variable = term, `LR Chisq` = statistic) %>% 
  gt::gt(groupname_col = "Model", caption = "OLS ANOVA tables") %>% 
  gt::cols_label(Variable = "") %>% 
  gt::fmt_number(columns=c(3,5), decimals = 1) %>% 
  gt::fmt_scientific(6, decimals = 1) %>% 
  gt::fmt_missing(columns=5:6, rows = everything(), missing_text = "---") %>% 
  gt::tab_style(style = gt::cell_text(indent = gt::px(40)), 
                locations = gt::cells_body(columns = "Variable"))

```

## Differences between quasibinomial and bootstrapped confidence intervals

```{r}
m1b_quasi <- models$TidyModel$m1b
m1b_diff <- (m1b_boot[,'boot 97.5%'] - m1b_boot[,'boot 2.5%'])/(m1b_quasi$conf.high-m1b_quasi$conf.low)

m3b_quasi <- models$TidyModel$m3b
m3b_diff <- (m3b_boot[,'boot 97.5%'] - m3b_boot[,'boot 2.5%'])/(m3b_quasi$conf.high-m3b_quasi$conf.low)

diff_sum <- summary(c(m1b_diff, m3b_diff))
```


# References
