---
title: "Depression and suicidality as credible signals of need"
author:
- name: Michael R. Gaffney
  affiliation: Department of Anthropology, Washington State University
- name: Kai Adams
  affiliation: UC Berkeley Haas School of Business
- name: Kristen L. Syme
  affiliation: Department of Experimental and Applied Psychology, Vrije Universiteit Amsterdam
- name: Edward H. Hagen
  affiliation: Department of Anthropology, Washington State University
  email: edhagen@wsu.edu
date: '`r format(Sys.Date(), "%B %d, %Y")`'
always_allow_html: false
output:
  bookdown::html_document2:
    toc: true
    toc_float: true
    number_sections: false
    css: style.css
bibliography: MTBib.bib
csl: apa-6th-edition.csl
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, warning = F, error = F, message = F)
# options(warn = 2)
library(equatiomatic)
library(knitr)
library(kableExtra)
library(glue)
library(ggplot2)
library(patchwork)
library(hagenutils)

tar_load(  
    c(
    'd',
    'd_US',
    'd_India',
    'signalingdata2018b',
    'signaling2020',
    'pilot_results',
    'power_curve',
    'power_curve2',
    'models',
    'models_prereg',
    'belief_compare_plot',
    'action_compare_plot',
    'effect_sizes',
    'effect_sizes2',
    'effectsPlots', 
    'effectsPlots_US',
    'effectsPlots_India',
    'ecdf_plot',
    'plot_T1Belief_T1Action',
    'plot_emotions',
    'plot_raw_beliefs',
    'plot_raw_actions',
    'plot_raw_belief_action2d',
    'plot_feedback',
    'med_prereg',
    'med1',
    'med2',
    'all_mediations',
    'm1_boot',
    'm3_boot',
    'm1b_boot',
    'm3b_boot',
    'dfT1emotion_vignette',
    'dfT1emotion_signal',
    'dfT2emotion_signal'
    )
)

if (!dir.exists('Figures')) dir.create('Figures')

figname <- function(){
  if (!exists('fignum')) fignum <<- 0
  fignum <<- fignum + 1
  paste0('Figures/Figure', fignum, '.pdf')
  }
```

# Abstract

Major depression and suicide are generally viewed as pathological responses to stress that elicit aversive responses from others. Alternatively, depression and suicide attempts might involve honest signals of need in response to stress that can increase support from reluctant others. To test this hypothesis, we examined responses to emotional signals in an experimental vignette study involving claims of substantial need in the presence of conflicts of interest and private information about whether the signaler was actually in need. In a sample of 1,240 participants recruited from Amazon Mechanical Turk, costlier signals like depression and suicide attempts resulted in greater perceptions of need, reduced perceptions of manipulativeness, and increased likelihood of support compared to simple verbal requests and crying without further symptoms. Additionally, as predicted, the effect of signaling on likelihood of support was largely mediated by the effect of signaling on belief in need. Our results support the hypothesis that depression and suicide attempts, apparent human universals, are credible signals of need.

# Introduction

Almost all individuals will suffer adversity, such as injury or loss of material or social resources, at some point in their lives. Psychological pain, such as sadness and low mood, probably evolved to motivate victims of adversity to shift their attention to the causes of adversity so as to mitigate its negative fitness consequences and to learn to avoid future such adverse events [@thornhill1989; @andrews2010; @Nesse1990]. In many cases, social partners could help victims, and therefore signals of psychological pain, such as sad expressions and crying, probably evolved to indicate need [e.g., @darwin1872; @bowlby1980; @balsters2013].

<!-- Almost all individuals will suffer adversity, such as injury or loss of material or social resources, at some point in their lives. Psychological pain, such as sadness and low mood, probably evolved to motivate victims of adversity to shift their attention to the causes of the problem so as to mitigate its negative fitness consequences and to learn to avoid future such adverse events [@thornhill1989; @andrews2010; @Nesse1990]. In many types of adversity, social partners could help minimize the costs. Therefore signals of psychological pain, such as sad expressions and crying, probably evolved to indicate need [e.g., @balsters2013]. CurrentDraft-->

There is strong evidence that most cases of depression are caused by major negative life events [@Hammen2005; @Mazure1998; @devries2011; @ellsberg2008]. Compared to non-depressed individuals, those with depression report about twice as many instances of negative events than non-depressed individuals [@Mazure1998] and more negative events than those with schizophrenia and bipolar depression across multiple studies [@Paykel1994]. In longitudinal studies, depression onsets soon after a negative event [@KendlerEtAl1999] or coincides with periods where adversity is likely to increase prior to it [e.g., depression starting before a divorce rather than after, @rosenstrom2017]. Additional support for the causal role of negative life events comes from the fact that the relationship between negative events and depression holds true when looking at events outside of one's control, showing that the connection is unlikely to be driven solely by individuals who are already depressed creating situations where negative events are likely to be common [@Hammen2005; @KendlerEtAl1999]. Furthermore, twin studies have shown that one's history of negative events is still a strong predictor of major depression when controlling for genetic similarity, and that part of the heritability of depression stems from the heritability of negative events like divorce and family conflict [@KendlerBaker2007; @KendlerEtAl1999]. For these and other reasons, most cases of depression are probably not due to mental illness but are instead functional instances of psychological pain [@Hagen:1999; @Hagen2003; @horwitz_loss_2007; @andrews2010; @SymeHagen2021].

There is equally strong evidence that depression is closely associated with anger and conflict [@cassiello2016]. Of the types of negative life events associated with depression, those which involve conflict tend to be more strongly associated with depression [@Hammen2005; @Mazure1998]. Marital problems, bullying, and abusive relationships are all common risk factors for depression [@KendlerEtAl1995; @KendlerEtAl1999; @klomek2007; @klomek2019], with assault (both sexual and non-sexual), in particular, greatly increasing one's risk of depression [@KendlerEtAl1995; @KendlerEtAl1999]. This holds true even in a small-scale, non-Western society: among the Tsimane, Amazonian horticulturalists, depression is also associated with conflict, particularly conflict involving non-kin [@StieglitzEtAl2015a]. See @SymeHagen2021 for a review of the association of depression with anger and conflict.

<!-- Other notable risk factors, like loss of a loved one or severe or prolonged illness, might seem less related to conflict. In these situations, however, individuals will often need increased investment from social partners, who might not be able, or want, to provide more investment than they already are. The ability for others to provide meaningful support to limit the harm in times of need [@SugiyamaSugiyama2003] combined with the ubiquity of conflicts of interests mean that many types of problems will eventually involve social conflict even if they did not start that way [@Hagen2003]. There is indeed increasing evidence that loss of a loved one is often followed by increased family conflict [see @SymeHagen2021 for a brief review]. -->

Other notable risk factors, like loss of a loved one or severe or prolonged illness, might seem less related to conflict. In these situations, however, the fitness costs that stem from reduced access to resources could be mitigated with help from others [@SugiyamaSugiyama2003] who might not be able, or want, to provide more investment than they already are. Indeed, there is increasing evidence that loss of a loved one is often followed by increased family conflict [see @SymeHagen2021 for a brief review]. Problems whose solutions require substantial investment or other changes on the part of social partners will therefore often involve social conflict even if they did not start that way [@Hagen2003]. When there are severe conflicts of interest, individuals have incentives to deceive, so "cheap" signals of need, such as sad expressions and crying, might not be believed.

## Bargaining: Credibly signaling need under conflict

<!-- Individuals who have suffered a negative life event will often need help to mitigate the fitness cost. To obtain this help, they must signal need, e.g., by verbal requests, sad facial expressions, and crying. But if they have conflicts of interest with social partners who could provide that help, and if their need is private information, their signals might not be believed if the signals, like crying, could easily be faked [@Zahavi1975; @mokkonen2016] and if the cost of believing a deceptive signal was high [@dawkins1978; @johnstone1993; @Hasson1994]. -->

<!-- Conflicts of interest are ubiquitous, even among closely related individuals. In species that rely heavily on cooperation, like our own, conflicts can arise over the levels of investment in a cooperative endeavor and the division of the resulting benefits. -->

With a cooperative species like our own, ubiquitous conflicts of interest means that there will always be disagreement over the levels of investment in a cooperative endeavor and the division of the resulting benefits. In many such cases, individuals who were dissatisfied with the terms of cooperation could seek other partners [@hammerstein2016]. In many other cases, however, it would be difficult or impossible to switch partners. Spouses who were dissatisfied with their partner's investment in their new infant, for example, would not easily find a different partner to invest in that infant [@Hagen:1999]. Similarly, an adolescent who was dissatisfied with her parent's investment in her would not easily find other parents who were willing to invest more, nor could parents easily produce another adolescent. In these examples, and many cooperative endeavors central to human biological fitness, all parties have monopoly power over the benefits they bring to the endeavor -- no one is easily replaced [@Hagen2002; @Hagen2003]. Such interdependence is increasingly recognized to be important to the evolution of cooperation in humans and other animals [@roberts2005; @tomasello2012; @aktipis2018].

@Hagen2003 proposed that physical aggression and core depression symptoms like loss of interest in virtually all activities were complementary strategies to resolve conflicts in interdependent relationships. @SellEtAl2009 found that physically formidable individuals were more prone to anger, prevailed more in conflicts of interest, and considered themselves entitled to better treatment. Physically or socially weaker individuals, though, are not without options to resolve conflicts in their favor. An individual with monopoly power over the benefits she contributes to a critical cooperative endeavor can withhold those benefits, or put them at risk, until her partners change their behaviors in ways that benefit her. Depression is characterized, in part by a profound loss of interest in virtually all activities [@dsm5], a symptom that is probably responsible for a sharp reduction in productivity [for review, see @SymeHagen2021]. Depression might therefore be a bargaining strategy for relatively powerless individuals in the wake of adversity and social conflict [@Hagen2003; @SymeHagen2021; see also @Watson:2002].

Bargaining models assume that delaying cooperation is costly so that there is an incentive to quickly agree on a division of benefits, and more so for those who highly value the fruits of the cooperative endeavor. In a classic non-cooperative game theory model of bargaining, @rubinstein1982 showed that two parties can come to an immediate agreement over division of benefits despite conflicts of interest if the parties' valuations of cooperation are not private information: in this case, each party knows exactly what division of benefits the other will accept, and can therefore make that offer immediately, avoiding the cost of delay. 

If valuations are private information, however, then costly delays might be unavoidable because each party has an incentive to request more than their actual valuation, and to reject the likely inflated requests from partners, leading to multiple rounds of bargaining. Models of bargaining with private information have a close relationship to models of credible signaling. A willingness to delay (i.e., refuse offers) credibly reveals one's low valuation of the endeavor, and therefore genuine need -- the benefit of waiting for a better offer outweighs the low cost of delay. Eagerness to reach a deal, on the other hand, credibly reveals a high valuation -- the benefit of waiting for a better offer does not outweigh the higher cost of delay. Once valuations are known, the game reduces to the one in @rubinstein1982, and the parties can reach an agreement on the division of benefits [@KennanWilson1993]. Delays also typically require that additional factors come into play [@feinberg2005 and references therein].

@Hagen2003 proposed that the substantial reduction in productivity that characterizes many cases of depression corresponds to a willingness to delay, and is therefore a credible signal of low valuation and need. In other words, reduced productivity is relatively less costly for those whose efforts are currently not yielding many fitness benefits than for those whose efforts are yielding substantial benefits.

<!-- Formal models of bargaining highlight the role of private information regarding the value each party places the cooperative endeavor. Withholding or risking one's contributions to a cooperative endeavor can be conceptualized as credibly signaling private information about one's need. Cooperation generates benefits over time. Thus, withholding or risking contributions incurs a *cost of delay* -- the more one values the endeavor, the greater the cost of delay.  -->

## Suicidality

Suicidal ideation is one of nine diagnostic criteria for a major depressive episode (MDE) [@dsm5], it is associated with depression across cultures [@haroz2017], and depression is a major risk factor for suicidal behavior [@hawton2013].

Anthropology has long viewed suicidality as largely the result of social problems. Early in the field's history, anthropologists tackled the problem of suicide as manifested in the small, traditional societies in which they conducted their fieldwork. Suicide, they found, was commonly a form of protest, revenge, and/or appeal [@firth1936; @Firth:1961; @Malinowski:1932; @niehaus2012]. Some ethnographers emphasized suicide as a form of anger or social pressure [@Hezel:1987; @Giddens:1964], whereas others emphasized the powerlessness of suicide victims [@Counts:1980].

Common to almost all theoretical and empirical investigations of suicide is a focus on *completed* suicides, i.e., suicide deaths. The vast majority of suicidal behavior, however, does not result in death. In young adult women in the US, for example, there are hundreds of attempts for every death (see Figure \@ref(fig:cdc)). @SymeEtAl2016 therefore argued that the theoretical focus should be on suicide ideation and suicide attempts.

```{r cdc, fig.width=8, fig.cap="US Suicidality non-fatal injury and death rates by age and sex (2001-2019). Data from @CDC2021."}
library(hrafsuicidedata)

plot_cdc <-
  cdc2 %>% 
  na.omit() %>% 
  ggplot(aes(Age, `Crude Rate`, colour = Sex, linetype = Type)) + 
  geom_line() +
  hagenutils::scale_color_binary() +
  guides(
    colour = guide_legend(title=NULL),
    linetype = guide_legend(title=NULL, reverse = T)
  ) +
  labs(x = '\nAge (years)', y = 'Annual rate\n(per 100,000)') +
  theme_minimal(15) +
  theme(axis.title.y = element_text(angle = 0, hjust=0))

plot_cdc
ggsave(figname(), plot_cdc)
```

Raymond Firth was one of the first to view suicidality as a gamble to improve one's circumstances in the here and now. Based on observations that suicide attempts often followed loss or conflict and varied substantially in their likelihood of death, he argued that a sizable subset of the suicide attempts among the Tikopia were not meant to end in death but instead were a means to elicit attention, status, or immediate reintegration into the community following negative events [@firth1936; @Firth:1961].

<!-- @Rosenthal1993 proposed a credible signaling model of what he, like Firth, viewed as 'gambling-type' suicide attempts. In the Rosenthal model, as long as that risk is greater than or equal to the net benefit a non-needy person would realize by sending the signal (discounted by the risk of death), social partners can trust that signal senders are genuinely in need despite conflicts of interest [@Rosenthal1993].  -->

In the bargaining framework, suicidality is conceptualized as putting all future contributions to cooperative endeavors with social partners at risk with some low but non-zero probability, credibly signaling low valuation of current circumstances [see @Rosenthal1993 for a game theory model of suicide attempts as costly signals]. On this view, most suicides deaths, especially in young, physically healthy individuals, would therefore be the inevitable consequence that some individuals lose this gamble. These and similar ideas [@andrews2006; @HagenEtAl2008; @nock2008; @Stengel1956] suggest that depression symptoms that put one's value to others at risk, such as loss of interest and suicidality, are credible signals of need.

Some evolutionary theorists have questioned the bargaining model based on evidence that depression is aversive to social partners [@nettle2004; @allen2006; @nesse2006; @myers2016]. These critics failed to recognize the central role of *conflict* in the model. Anger is aversive to targeted social partners, yet it is probably an adaptation that exploits advantages in physical or social formidability to force concessions from them [@SellEtAl2009]. Aversive responses to depressive bargaining are likewise expected throughout the process, with potential helpers signaling the costliness of providing support in turn. Importantly, as with anger, aversive responses to depression encourage concessions by interdependent social partners [@SymeHagen2021].

<!-- Compared to suicide, however, there are fewer anthropological investigations of depression [@kleinman1985] despite frequent reports of depression or depression-like behavior in a wide range of societies [@haroz2017; @liu2020] and the potential for cross-cultural research to help understand depressive states [@haroz2017]. -->

<!-- As they are prohibitively costly for those in minor need, the costs of depression and suicidal behavior may be particularly important in times of substantial conflicts of interest or when the level of support needed to address one's situation is substantial, both of which should be expected to increase the payoff of manipulating others through fake signals. When either condition applies, less costly signals like verbal requests, facial sadness, and crying are likely to be less reliable as their low costs can easily be made up for with increased social support, thereby favoring the use of costlier signals so long as the costs of signaling are outweighed by the benefits [@SymeHagen2021]. -->

<!-- Alternative start to line 92: Although verbal requests, sad facial expressions, and crying can be used to signal need, they are likely to be less effective in times of severe conflicts of interests or times when the level of support needed to address one's situation is substantial. This is because both conditions increase the payoff of manipulating others through fake signal.  -->

<!-- Both depression and suicide attempts are more common outcomes of these sorts of conflict than is often appreciated by the biomedical approach (Syme and Hagen 2021). With notable exceptions (Kendler; Mazure; Hammen), the biomedical model tends to frame both suicide and depression as a problem with the individual, e.g., as one involving psychiatric disturbance, drug addiction, or stress sensitivity [@wray2011 (changed wording to include depression]. This theoretical approach has made little progress: rates of suicide, which vary widely by country, have remained relatively flat in countries for which there are long-term data, such as the US, UK, Japan, and Sweden [@owidsuicide] (tweak for depression). -->

<!-- ## Depression and suicide attempts as both signals and bargaining tools -->

<!-- Evolutionary theories of depression tend to either ignore suicidality, or take it as evidence that depression is dysfunctional. Wakefield, for instance, who has devoted much of his career to arguing that most cases of depression are simply severe but normal sadness and grief, nevertheless takes suicidality to indicate pathology for some cases of depression [e.g., @wakefield2017]. @nettle2004 argues that the costs of depression, including suicidality, outweigh any benefits, and therefore adaptationist models of depression are implausible. @gilbert2006 sees suicidality as an example of a dysregulated escape defense [for reviews of evolutionary theories of depression, see @Hagen2011]. -->

<!-- Another possible explanation for the connection between depression, suicide, and social adversity is that some instances of depression and suicide attempts serve as means to elicit increased support from others by honestly signaling need to those invested in the well-being of the signaler, [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], something which may be particularly useful in times when an individual cannot solve substantial fitness-relevant challenges on their own [@Hagen2003; @SymeEtAl2016]. -->

<!-- Due to the ease with which signals can be faked [@Zahavi1975; @mokkonen2016] and the costs of accepting a deceptive signal as reliable [@dawkins1978; @johnstone1993; @Hasson1994], demonstrating that one's need is genuine has likely been a recurrent adaptive problem since humans evolved the capacity to provide instrumental support to one another. One potential solution to this is to produce signals that are unprofitable to fake [@Zahavi1975]. This is possible whenever the benefits of signaling outweigh the costs for those who have the relevant underlying quality but the costs outweigh the benefits for those who lack the quality or possess it to a lesser degree [@zahavi1977; @grafen1990]. In particular, if the benefits of sending the signal are the same for signalers regardless of condition, but the signal cost outweighs this benefit for signalers without the underlying quality (a "costly signal"), then individuals with the underlying quality can evolve to send the signal and receivers can evolve to respond in mutually beneficial ways to the signal. -->

<!-- Depression and suicidal behavior fit this criterion as both are relatively costly for those not facing adversity than those who are. In the case of depression, the large-scale reduction in one's normal routine of fitness enhancing behavior is relatively less costly for those whose situations result in a limited ability to obtain fitness-relevant benefits, with depression likely adding no additional costs in instances of extreme deprivation [@Hagen2003]. When it comes to suicide, the bulk of the costs associated with suicide attempts are only felt in times that result in death or severe bodily harm. For this reason, the way others view and respond to suicide attempts is likely to be strongly influenced by the probability that the attempt would result in death or severe injury [@Rosenthal1993]. Despite this difference, the previously mentioned logic still applies with increased risk of death being relatively more costly for those able to advance their fitness prospects in their day-to-day lives. -->

<!-- The costs of depression and suicidal behavior may be particularly important in times of conflict or when the level of support needed to address one's situation is substantial, both of which should be expected to increase the payoff of manipulating others through fake signals. When either condition applies, less costly signals like verbal requests, facial sadness, and crying are likely to be less reliable as their low costs can easily be made up for with increased social support, thereby favoring the use of costlier signals so long as the costs of signaling are outweighed by the benefits [@SymeHagen2021]. -->

<!-- Durkheim's pioneering study of suicide [@durkheim1897] helped launch sociology as a quantitative discipline. Sociological investigations of suicide flourished through much of the 20th century, but were supplanted by medical and public health approaches as the biomedical model of mental health took hold in the late 1970's and early 1980's [@wray2011; @syme2020]. The biomedical model tends to frame suicide as a problem with the individual, e.g., as one involving psychiatric disturbance, drug addiction, or stress sensitivity [@wray2011]. This theoretical approach has made little progress: rates of suicide, which vary widely by country, have remained relatively flat in countries for which there are long-term data, such as the US, UK, Japan, and Sweden [@owidsuicide]. -->

<!-- ## Conflict plays a central role in suicidality and depression -->

<!-- Early on, anthropologists also tackled the problem of suicide as manifested in the small traditional societies in which they conducted their fieldwork. Suicide, they found, was commonly a form of protest, revenge, and/or appeal [e.g., @Malinowski:1932; @niehaus2012]. Some ethnographers, such as @Hezel:1987, conceptualized suicide as a form of anger, whereas others, such as @Giddens:1964, conceptualized suicide as a form of social pressure. @Counts:1980 emphasized the powerlessness of suicide victims. Common to almost all theoretical and empirical investigations of suicide is a focus on *completed* suicides, i.e., suicide deaths. @SymeEtAl2016, in contrast, argued that because the vast majority of suicidal behavior does not result in death, the proper focus should be on suicide ideation and suicide attempts (see Figure @ref(fig:cdc)). Suicide attempts, as we explain next, are a gamble to improve one's life by taking a low but non-zero risk of death. Most suicides deaths, especially in young, physically healthy individuals, are therefore the inevitable consequence that some individuals lose this gamble. -->

<!-- Within anthropology, Raymond Firth was likely the first to conceptualize suicidality as a gamble to improve one's circumstances in the here and now. Based on observations that suicide attempts often followed social conflict or loss, and varied substantially in their likelihood of death, he argued that a sizable subset of the suicide attempts among the Tikopia were not meant to end in death but instead were a means to elicit attention, status, or immediate reintegration into the community following negative events [@firth1936; @Firth:1961]. Human communites are highly interdependent [\@] -->

<!-- Suicidality is closely linked to depression. Suicidal ideation is one of nine diagnostic criteria for a major depressive episode [MDE, @dsm5], it is associated with depression across cultures [@haroz2017], and depression is a major risk factor for suicidal behavior [@hawton2013]. Compared to suicide, however, there are fewer anthropological investigations of depression [@kleinman1985] despite frequent reports of depression or depression-like behavior in a wide range of societies [@haroz2017; @liu2020] and the potential for cross-cultural research to help understand depressive states [@haroz2017]. -->

<!-- Evolutionary theories of depression tend to either ignore suicidality, or take it as evidence that depression is dysfunctional. Wakefield, for instance, who has devoted much of his career to arguing that most cases of depression are simply severe but normal sadness and grief, nevertheless takes suicidality to indicate pathology for some cases of depression [e.g., @wakefield2017]. @nettle2004 argues that the costs of depression, including suicidality, outweigh any benefits, and therefore adaptationist models of depression are implausible. @gilbert2006 sees suicidality as an example of a dysregulated escape defense [for reviews of evolutionary theories of depression, see @Hagen2011]. -->

<!-- and anthropological explanations for low mood have become more common. In general, these approaches see sadness and minor depression as ways to motivate adaptive behavior changes in response to negative outcomes in one's environment (cite Nesse and others), mainly differing in how they view major depression. While some approaches present most forms of major depression as dysfunctional (inflammation and some of Nesse), others have extended the logic on the benefits of low mood and reduced behavior to more severe instances of sadness. Of the possible benefits, the ability of depression to help solve fitness-relevant in times of social conflict is the most emphasized, with depression being suggested to help by reallocating resources with the body for problem solving (cite ARH), reducing the costs of interacting with higher status individuals (cite), or serving as a means to both signal one's need and bargain for increased support (cite).  -->

<!-- As with suicide, research on depression since the 1980's has been dominated by biomedical approaches of limited success [@patten2016; @ormel2019].  -->

<!-- and there is still substantial uncertainty about why both are as common as they are (cite). Likely due to  the suffering they cause, the vast majority of these approaches have been built on the assumption that they are the result of internal dysfunction (cite examples). However, a subset of anthropologists and evolutionary theorists have challenged this idea, suggesting that despite the subjective suffering they cause, they may function as a means to limit negative fitness outcomes in times of substantial adversity (Firth + others). -->

<!-- Due to depression's relationship to suicide [@who2001; @hawton2013], it should not be surprising that suicidality also often coincides with social adversity and conflict. Grief over the loss of a loved one [@firth1936: @Firth:1961; @devries2011], bullying [@alavi2017; @butler2019; @ford2017; @wang2020], and instances of social transgression [@Firth:1961; @Malinowski:1932; @SymeHagen2018; @lester1997] have all been reported to lead to suicidal behavior in both industrialized and traditional societies. Of the recorded instances of suicide attempts in the Human Relations Area Files (HRAF), 77.4% of attempts were found to be linked to times of social conflict, 88.7% were linked to fitness threats in general, and 84.9% were linked to situations in which attempters were powerless to resolve their situation alone [@SymeEtAl2016]. As with depression, assault is an particularly powerful risk factor for suicidal behavior based on multiple large-scale multi-country studies [@devries2011; @ellsberg2008] and more localized examinations [@reigstad2017]. -->

<!-- ## Depression and suicide attempts as both signals and bargaining tools -->

<!-- One possible reason for the connection between depression, suicide, and social adversity is that some instances of depression and suicide attempts serve as means to elicit increased support from others [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], particularly in times when an individual cannot solve substantial fitness-relevant challenges on their own [@Hagen2003; @SymeEtAl2016]. According to the bargaining models of depression and suicide, the small groups humans lived in for much of our evolutionary history created many situations in which certain forms of cooperation were unlikely to be easily replaced by others while also allowing for harmful or unfair social relationships to develop [@Hagen2003; @SymeEtAl2016]. In these instances, one's monopoly on one or more types of cooperation could be leveraged to change how one is treated if others valued an individual's cooperation sufficiently and believed them to be in genuine need. -->

<!-- Due to the ease with which signals can be faked [@Zahavi1975; @mokkonen2016] and the costs of accepting a deceptive signal as reliable [@dawkins1978; @johnstone1993; @Hasson1994], demonstrating that one's need is genuine has likely been a recurrent adaptive problem since humans evolved the capacity to provide instrumental support to one another. One potential solution to this is to produce signals that are unprofitable to fake [@Zahavi1975]. This is possible whenever the benefits of signaling outweigh the costs for those who have the relevant underlying quality but the costs outweigh the benefits for those who lack the quality or possess it to a lesser degree [@zahavi1977; @grafen1990]. In particular, if the benefits of sending the signal are the same for signalers regardless of condition, but the signal cost outweighs this benefit for signalers without the underlying quality (a "costly signal"), then individuals with the underlying quality can evolve to send the signal and receivers can evolve to respond in mutually beneficial ways to the signal. -->

<!-- Depression and suicidal behavior fit this criterion as both are relatively costly for those not facing adversity than those who are. In the case of depression, the large-scale reduction in one's normal routine of fitness enhancing behavior is relatively less costly for those whose situations result in a limited ability to obtain fitness-relevant benefits, with depression likely adding no additional costs in instances of extreme deprivation [@Hagen2003]. When it comes to suicide, the bulk of the costs associated with suicide attempts are only felt in times that result in death or severe bodily harm. For this reason, the way others view and respond to suicide attempts is likely to be strongly influenced by the probability that the attempt would result in death or severe injury [@Rosenthal1993]. Despite this difference, the previously mentioned logic still applies with increased risk of death being relatively more costly for those able to advance their fitness prospects in their day-to-day lives. -->

<!-- The costs of depression and suicidal behavior may be particularly important in times of conflict or when the level of support needed to address one's situation is substantial, both of which should be expected to increase the payoff of manipulating others through fake signals. When either condition applies, less costly signals like verbal requests, facial sadness, and crying are likely to be less reliable as their low costs can easily be made up for with increased social support, thereby favoring the use of costlier signals so long as the costs of signaling are outweighed by the benefits [@SymeHagen2021]. -->

<!-- Beyond signaling need to those invested in the well-being of the signaler, depression and suicidal behavior may also help an individual bargain for increased support [@Hagen2003]. This is possible due to the reduced investment in others that frequently accompanies depression or the threat to future cooperation that comes from suicide. In these instances, depression and suicide attempts may function analogously to a labor strike, sending the message that one’s need is genuine and that if the target of the signal wanted to continue receiving benefits from the depressed, they would need to increase support towards the signaler. Due to this bargaining component, aversive responses from the targets of the signal to the signaler are expected throughout the bargaining process, with potential helpers likely signaling the costliness of providing support. Importantly, such responses do not prevent the delivery of meaningful benefits, they are simply expected to reduce the risk of helpers providing too much support for too little need. -->

<!-- ### Determining the impact of cues versus signals -->

<!-- Distinguishing between the information provided by signals versus cues is a major challenge for signaling approaches to behavior. A signal differs from a cue in that it is intentionally sent by an organism to alter the perceptions and behaviors of others, while a cue is any piece of information that influences an individual's behavior or physiological state that was not created for communicative effect [@MaynardSmithHarper1995]. Regarding the signals of interest, suicide attempts are more likely to take the form of pure, external signals than other symptoms of depression. This is because low mood's impact on behavior for non-signaling benefits likely conveys information as cues in addition to what might exist for signaling purposes. In other words, the signaling component of depression may come from the upregulation of symptoms beyond what would be optimal for non-signaling purposes rather than the creation of entirely new symptoms. Suicidality is likely an exception to this, due to the likelihood most suicide attempts do not have non-signaling functions [see @deCatanzaro1991 for a possible exception]. -->

<!-- ## Responses to depression and suicidal behavior  -->

<!-- For depression and suicide attempts to be worth the costs involved with signal creation, meaningful social support would be required [@Hagen2003; @SymeEtAl2016]. Most studies examining responses to depression have been observational and focused on responses outside of changes in social support. In general, most report aversive responses and high levels of rejection towards the depressed, and depressed individuals are often linked to impaired social functioning [@EvraireDozois2011; @GotlibLee1989; @HirschfeldEtAl2000]. One explanation for this is that depressed individuals engage in behaviors that drive others away. In line with this reasoning, both excessive reassurance seeking and negative feedback seeking, which often accompany depression, have been shown to elicit negative emotional responses from others and increase the chance of rejection from others [@JoinerMetalsky1995; @StarrDavila2008]. Negative interactions are also seen in families as depressed children tend to report less support and more negative responses from parents than non-depressed children, although it has not been demonstrated that such interactions did not lead to depression in the first place [@MesserGross1995; @SchwartzEtAl2012]. -->

<!-- Despite being looked for less frequently, beneficial responses to depressed individuals are also reported. Depression has been associated with reduced likelihood of aggression within families [@hops1987; @dadds1992], including those already experiencing high levels of conflict [@SheeberEtAl2001]. There is also limited evidence that depression can result in increased levels of support from others outside the family. For example, non-depressed individuals were shown to be more cooperative and exhibit more caretaking behavior towards their college depressed roommates over the course of their time living together [@HokansonEtAl1986]. Likewise, a study examining responses to confederates trained to appear ‘depressed’, ‘anxious’, or ‘normal’ found that individuals were more likely to provide advice and support to ‘depressed’ individuals than ‘anxious’ or ‘normal’ individuals despite the confederate in the ‘depression’ condition being viewed more negatively than the other conditions and the ‘depression’ condition involving similar levels of distress compared to the ‘anxious’ condition [@stephens1987].  -->

<!-- Most studies examining responses to those who have survived suicide attempts have focused on stigmatization. In these studies, perceptions of survivors as being weak, selfish, mentally ill, and antisocial are all being commonly reported [@BatterhamEtAl2013; @TzengLipson2004], with stigmatization often being found within and outside one's social network [@FreyEtAl2016; @ScoccoEtAl2012]. Despite this potential for stigmatization, increased social support and changes to important relationships have been reported to follow suicide attempts with some indication that these effects may hold long term [@Stengel1956]. For example, a study of 100 women who survived suicide attempts found that individuals gained identifiable benefits through the attempt in 75 cases, with 41 individuals benefiting from reconciliations with others [@Lukianowicz1971]. Unlike many Western countries, where suicide is often viewed as pathological [@Hidaka2012], members of traditional societies have been reported to view suicide attempts as cries for help rather than mental illness [@Shostak1981], with attempts having been described as ways of escaping unwanted marriage arrangements, persistent abuse, or a lack of support in obtaining mates by those engaging in the behavior and observers [@GutierrezdePinedaMuirden1948; @Hilger1957; @Karsten1935; @Tessmann1930; @Wilson1960]. Although findings that individuals view suicide attempts as cries for help is not necessarily evidence for their ability to lead to beneficial responses, 30 out of 84 examples of suicidal behavior included in the HRAF resulted in positive changes for the survivor, despite suicide rarely if ever being the main topic [@SymeEtAl2016]. -->

## Study aims and predictions

Most of the limited literature on social responses to depression and suicidality comprises observational studies of depressed individuals interacting with family, friends, or roommates [@hops1987; @dadds1992; @SheeberEtAl2001; @JoinerMetalsky1995; @StarrDavila2008]. These have ecological validity, but cannot easily determine causal relationships. Some studies, though, have employed an experimental design in which participants were randomized into conditions in which they listened to, watched, or interacted with either a depressed or non-depressed person, where in some cases the depressed person was a non-depressed confederate enacting a depressed role [@marcus1992]. These designs can demonstrate causation but the transient, inconsequential relationships and laboratory settings lack ecological validity.

Experimental vignette studies, which employ a short, carefully constructed description of a person, object, or situation, aim to approach the ecological validity of observational studies by presenting participants with rich, real-world scenarios, while at the same time allowing researchers to randomize participants into conditions in which theoretically relevant dimensions of the vignettes are systematically manipulated, thus enabling robust causal inferences [@atzmuller2010]. Experimental vignette studies are conducted in a broad range of disciplines, including psychology, economics, sociology, management studies, political science, and education [@atzmuller2010; @aguinis2014].

In the bargaining framework, the willingness to delay credibly signals the valuation of current cooperative arrangements, with greater delays indicating lower valuations. The primary aim of this study was to examine responses to emotional signals that varied in the extent to which they reduced productivity, which we refer to as costs, in an experimental vignette study in which a possible victim of adversity asks for help from the participant, but has incentives to exaggerate her need. As signal cost increased, we predicted that participants would report (1) increased belief in the signaler's claims and (2) increased likelihood of providing help, with (3) the increased likelihood of providing help mediated by the increased belief in the signaler's need. 

<!-- A secondary aim was to investigate a scenario that is common in traditional societies but uncommon in Western societies -- arranged marriages -- where conflicts between parents and offspring often result in depression and suicidality [@SymeEtAl2016] -->

# Materials and methods

## Design

This study utilized a between-subjects pretest-posttest design to examine how four different emotional signals (treatments) would influence (1) the degree participants believed a fictional character to be in need (*Belief*) and (2) the likelihood they would provide help (*Action*) relative to a simple *Verbal request* without additional signaling (the control condition), in four different vignettes, for a total of 20 conditions. In this design, the outcomes are measured at pre-treatment (T1). Participants are then randomized into either a control group or a treatment group, i.e., one of the emotional signals, and the outcomes are measured again (T2). Regression models (described later) are used to determine the effect of the treatment conditions on the posttest outcome variables, relative to the control condition, controlling for pretest levels of the outcome variables (we also explored within-subjects effects of the signal on outcomes at T2 compared to T1).

In principle, pretest-posttest designs, by controlling for pretest variation in the outcome, increase the precision of the estimate of the treatment effect on the outcome [@dimitrov2003]. In survey experiments, however, researchers often favor posttest-only designs over pretest-posttest designs. The common concern is that the pre-treatment measurement of the outcome will influence the treatment effect on the outcome (i.e., the effects of asking the same question twice) due to, e.g., demand effects, in which participants try to conform to experimenter expectations, or to consistency pressures, in which participants try to provide consistent responses regardless of treatments [@clifford2021]. In a study with six experiments that randomly assigned respondents to alternative designs (e.g., pretest-posttest, posttest only) @clifford2021 found these concerns to be overblown. In all cases, the pretest-posttest design had substantially greater precision than the posttest-only design, with little evidence that pretest measurement altered the treatment effect.

## Lessons learned from two pilot studies

The current study is a refinement of a large MTurk experimental vignette pilot study (N=`r nrow(signalingdata2018b)`) that used a different vignette but very similar signals and outcomes (see below for more details on MTurk samples), and a much smaller pilot study posted to [reddit.com/r/SampleSize/](https://www.reddit.com/r/SampleSize/) (N=28) that used draft versions of three vignettes used in the current study, along with the same signals and outcomes. One major goal of the MTurk pilot study was to determine if believability and willingness to help were simply artifacts of the fictional victim's psychiatric distress. We therefore included a "signaling" condition in which the victim exhibited schizophrenic symptoms. As predicted, believability and willingness to help in this condition were dramatically lower than in any other condition (see Figure \@ref(fig:piloteffects)), ruling out this alternative explanation. We therefore did not include the schizophrenic condition in the current study. See the SI for more details on the pilot studies.

## Power analysis

```{r}
pwr80_90 <- 
  power_curve %>% 
  summarise(
    pwr80 = min(sample_size[power>=.8]), 
    pwr90 = min(sample_size[power>=.9])
  )
```

We used the MTurk pilot data to estimate the sample sizes needed to detect an effect of the *Mild depression* signal vs. *Verbal request* control on *Belief* in the victim's need. Power was about 80% for a sample size of about `r pwr80_90[[1, 1]]`, and was about 90% for a sample size of about `r pwr80_90[[1, 2]]`. See Figure \@ref(fig:power). Given our \$1500 USD budget, we aimed for a sample size of 120-130 for treatment plus control conditions, and 1200-1300 for all conditions in the study. For more details, see the SI.

## Sampling

Participants for this study were recruited from Amazon Mechanical Turk, a crowdsourcing platform that allows for the creation of Human Intelligence Tasks (HITs) that workers can complete for pay. As Amazon provides the infrastructure, it allows for a relatively low-cost way of collecting data for academic research at the cost of the data not being representative of any real population [@thomas2017]. Despite this limitation, MTurk samples have a wider range of ages and incomes than most university samples [@kennedy2020; @DworkinEtAl2016; @thomas2017]. US MTurk samples also differ from the general US population, mainly in being younger, more educated, and lower income [@RossEtAl2010; @boas2020].

To test emotional signals in an arranged marriage vignette, we recruited an Indian MTurk sample. Indian samples are also likely to be younger, more educated, and have higher income than the general Indian population, and are more likely to come from regions with good internet access [@boas2020].

Overall, the quality of data provided by MTurk workers tends to resemble that of university sample pools [@NeckaEtAl2016; @robinson2019; @thomas2017], with some studies reporting that MTurk samples are more attentive than samples of university students [@HauserSchwarz2016]. In vignette studies, MTurk data quality also compares favorably to that from much more expensive population-based samples [@weinberg2014]. For these reasons, concerns about data quality come primarily from the threat of bot use or respondents faking their location to take surveys in a language they do not understand well, with there being little evidence of the former [@kennedy2020] and the risk of the latter able to be minimized through well designed attention checks, timed responses, and good study design [@aguinis2020; @huang2015; @kennedy2020; @thomas2017].

## Participants

All participants were over 18, located in the United States or India, and had high quality MTurk metrics [completed at least 100 HITs with a HIT approval rate of over 98%, @kennedy2020]. Participants were excluded from the study if (1) they read the vignette too quickly (one-third of the time it took MG to read it), and (2) they failed clearly labeled attentions checks. The first attention check was shown immediately after the consent form and provided participants with a random word and asked them to enter the vowels in the order in which they are found in the word. The second attention check followed the vignette and involved asking three questions about the story that were easy to answer for anyone paying attention.

## Ethics

All participants provided informed consent, and the consent form warned that some content might involve sexual assault. We estimated the study would take 4-8 minutes to complete for participants who did not take breaks [MTurkers commonly multitask, or leave the survey page and return later, @NeckaEtAl2016]. All participants who passed the attention checks were paid \$1 for their time, for an estimated rate of \$7.50/hr to \$15/hr (75% of US participants completed in `r signif(quantile(d_US$TotalTime, 0.75)/60, 2)` minutes; 75% of Indian participants completed in `r signif(quantile(d_India$TotalTime, 0.75)/60, 2)` minutes). This study was certified exempt by the Washington State University Human Research Protection Program.

## Survey

Four vignettes were used in this study that involved (1) a female's claim of severe adversity that was private information, (2) conflicts of interest between the victim and the participant that would undermine the believability of her claims, and (3) her emotional signals. The vignette scenarios involved potentially severe types of adversity, such as sexual and non-sexual assault and thwarted marriage, that often precede cases of depression and suicidality in the ethnographic and clinical record. See Table \@ref(tab:design).

### Time 1: Claim of need in a conflictual relationship

At Time 1 (T1) participants in the US sample were randomly assigned to either the "basketball coach," "romantic partner," or "brother-in-law" vignettes, and the Indian sample was assigned to the Thwarted marriage vignette.

**Basketball coach vignette**: Each participant was asked to imagine that they are a university athletic director. The star player on the women's basketball team comes to the participant and claims she was sexually assaulted by her head coach, a physically powerful man. However, there is a history of conflict between the star player and the coach over playing time, and police are unable to find evidence to corroborate her claims.

**Brother-in-law vignette**: Each participant was asked to imagine that they let their sister, brother-in-law, and niece move in with them after their sister's family lost their house in a fire. During this time, the participant's 15 year old daughter becomes jealous of her cousin, who appears to be a social competitor. A few weeks after claiming her cousin was trying to steal her boyfriend, the participant's daughter accuses her uncle of sexually assaulting her.

**Romantic partner vignette**: Each participant was asked to imagine that they found a highly desirable romantic partner after years of being single. However, the participant's 13-year-old daughter, who has a history of interfering with the participant's past relationships, is clearly unhappy with the new partner. After a period of sustained conflict with both the participant and the romantic partner, she accuses the romantic partner of physically assaulting her, but cannot produce any evidence.

**Thwarted marriage vignette**: Indian sample only. Each participant was asked to imagine that their family was trying to arrange a dowry for their older daughter (the signaler in this vignette) so she can marry a man she already loves, while still saving enough money for their younger daughter's dowry. After the man's family demands more money, the participant's family tries to find a second man, who the older daughter claims to find unattractive. Any increase in the dowry will come at the younger daughter's expense. The participant therefore proceeds to arrange a marriage to the second man as the first man's family makes arrangements with a different woman.

The full vignettes are available in the SI.

+-------------------+--------------------------------------+---------------------------------------------------------------+--------------------------------------------+
| Vignette          | Cooperative Endeavor                 | Conflict of interest                                          | Private information                        |
+===================+======================================+===============================================================+============================================+
| Thwarted marriage | Inclusive fitness (parent-offspring) | Parental investment in sibling                                | Value of arranged marriage with second man |
+-------------------+--------------------------------------+---------------------------------------------------------------+--------------------------------------------+
| Basketball coach  | Winning the championship             | Coach's investment in other players; keeping the coach        | Did sexual assault happen?                 |
+-------------------+--------------------------------------+---------------------------------------------------------------+--------------------------------------------+
| Romantic partner  | Inclusive fitness (parent-offspring) | Investment in offspring vs. romantic partner                  | Did physical assault happen?               |
+-------------------+--------------------------------------+---------------------------------------------------------------+--------------------------------------------+
| Brother-in-law    | Inclusive fitness (parent-offspring) | Investment in child vs. investment in adult sibling and niece | Did sexual assault happen?                 |
+-------------------+--------------------------------------+---------------------------------------------------------------+--------------------------------------------+

: (\#tab:design) The cooperative endeavor, conflict of interest, and private information in each vignette

### Baseline measures (T1)

After reading the vignettes, participants rated their belief the signaler to be telling the truth (*T1 Belief*: 0-100) and the likelihood of them helping the signaler as requested (*T1 Action*: 0-100). With the thwarted marriage vignette we also asked how they would split the money they had saved for the dowry between their daughters (*T1 Divide*: 0-100; 50 is equal split). In every instance, the order of the questions was randomized to avoid order effects [@KrosnickAlwin1987].

Responses were recorded with sliders due to the fact they allow for finer grained changes than categorical scales [@KlimekEtAl2017]. Based on findings that a slider's starting position may bias results [@liu2019], we had concerns that participants would be less likely to move away from intermediate starting values than they would be in reality. For this reason, we set each T1 slider to start fully to the left (0). For the exact wording of the question and the labels on the sliders, see Table \@ref(tab:outcomes).

We also asked which emotions participants felt the signaler was experiencing using a multiple choice question in which they could select as many options as they would like. This included emotions directly related to the signals (e.g., sad, depressed, and suicidal), states which suggest genuine need (traumatized and violated), states which suggest deception (e.g., deviousness or jealousy), and was the victim mentally ill. The complete list can be found in the SI. To further explore the effect of signaling on participants' inferences of the signaler's emotional state, we created two new variables: *Low mood* was the sum of the binary variables *Depressed*, *Distressed* and *Sad*; and *Manipulative* was the sum of the binary variables *Devious* and *Jealous.*

### Time 2: Signals

After rating their beliefs and actions, and which emotions they thought the potential victim was experiencing, participants were randomized into either the control condition or one of four emotional signals by the victim (in order of increasing signal cost): (1) control condition: a verbal request without additional signaling; (2) crying; (3) mild-depression; (4) depression; and (5) a suicide attempt. The signals involved the participant encountering the victim some time after the adverse event and observing, e.g., crying; sad expressions; reduced effort, fatigue, and poor personal hygiene; and suicidal self-injury. These descriptions did not use the terms depression, depressed, suicidal, or mental health.

The signals were cumulative: crying can be an important feature of depression [for discussion on the relationship between crying and depression see @bylsma2020], and depression is a major risk factor for suicide [@BostwickPankratz2000; @Kessler2012]. Accordingly, components of less-costly signals were included in more-costly signals. Although we use the term 'signals' throughout the paper for brevity, we expect that the hypothesized signals of need, like many signals, would also provide information to others in the form of cues [for discussion on the evolution of signals from cues see: @biernaskie2018; @Tiokhin2016; @Steinkopf2015]. The complete texts of the signals are available in the SI.

### Post-treatment measures (T2-T3)

After reading the signaling text, participants answered questions identical to those asked at T1 as the main post-treatment variables of interest (T2). For the *Belief*, *Action*, and *Divide* variables, the position of the slider starting where participants placed it at T1. T2 emotion multiple-choice questions were identical to those used in T1, as were the composite variables *Low mood* and *Manipulative*.

As both a validity check and a way to understand the degree participants would be willing to help if they believed the participant completely, we presented participants with strong evidence that the claims were true. In the US sample, this involved telling participants there was video evidence of the event in question occurring or a similar event after the fact. With the thwarted marriage vignette, this involved the participant seeing the man their daughter wants to marry trash talk their daughter and their family. Participants were then asked to rate their likelihood of acting (*T3 Action*).

### Demographic Questions

The final part of the survey was a brief demographic questionnaire which asked for the (1) age, (2) sex, (3) number of siblings, (4) number of sons, (5) number of daughters, (6) current relationship status, (7) highest level of education, and (8) the annual household income of each participant.

## Statistical analyses: Preregistered and modified

Our intervention was *signal*, an ordinal variable with the following preregistered rank order: verbal request (control), crying, mild depression, depression, suicide attempt. We coded this 5-level ordinal factor variable using default 4th-order orthogonal polynomial contrasts. We preregistered a test of our hypothesis that used ordinary least squares (OLS) regression models with the following form:

$$
\begin{aligned}
\operatorname{Belief}_{T2} &= \beta_{0} + \beta_{1}(\operatorname{Belief}_{T1}) + \beta_{2}(\operatorname{signal}_{\operatorname{.L}}) + \beta_{3}(\operatorname{signal}_{\operatorname{.Q}})\ + \\
&\quad \beta_{4}(\operatorname{signal}_{\operatorname{.C}}) + \beta_{5}(\operatorname{signal}_{\operatorname{\text{^}4}})
\end{aligned}
$$ 

$$
\begin{aligned}
\operatorname{Action}_{T2} &= \beta_{0} + \beta_{1}(\operatorname{Action}_{T1}) + \beta_{2}(\operatorname{signal}_{\operatorname{.L}}) + \beta_{3}(\operatorname{signal}_{\operatorname{.Q}})\ + \\
&\quad \beta_{4}(\operatorname{signal}_{\operatorname{.C}}) + \beta_{5}(\operatorname{signal}_{\operatorname{\text{^}4}})
\end{aligned}
$$ 

We predicted that there would be a statistically significant monotonically increasing effect of the signal on *Belief* and *Action*.

We decided to use generalized linear regression models instead of OLS, however, for the following reasons. Our pre-test and post-test measures, *T1 & T2 Belief* and *T1 & T2 Action*, were all measured on a 0-100 point scale. A substantial number of participants rated their beliefs and actions as exactly 0 or exactly 100 at either T1 or T2. OLS linear regression models are not suitable for a closed and bounded distribution with so many values on the boundary because the residuals would not be normally distributed or have constant variance. Beta regression, a possible alternative, is only defined on the open interval, (0, 1), whereas our data had many values on the boundaries (a closed interval). A variant of beta regression, zero-one inflated beta, is inappropriate because it assumes different processes for the generation of data on the boundaries vs. the interior, and we had no reason to believe that our zeros and ones were generated by a different process than the rest of our observations.

We therefore fit models of our two outcome variables (which we rescaled to the continuous interval [0, 1]) in five ways. First, we fit the preregistered OLS models (reported in the SI). Second, we fit fractional regression models, introduced in a seminal paper by @papke1996 for continuous outcomes on the interval [0, 1], which, in brief, involve fitting a quasi-likelihood version of logistic regression with robust standard errors. This method is implemented in the `frm` package [@ramalho2019]. Third, we fit GLM logistic regression models, computing 95% CI's via bootstrapping, using the `glmmboot` package created to model continuous outcomes on [0, 1] [@humphrey2020]. Fourth, we fit GLM models with the quasibinomial family and logit link using the `glm` function in R stats library, which can fit continuous outcomes on [0, 1]. Finally, we computed robust standard errors and 95% confidence intervals from the quasibinomial models using the `sandwich` package [@zeileis2006]. The estimated coefficients, standard errors, and confidence intervals produced by the latter four methods were *very* similar, and often identical (see Figure \@ref(fig:modelcompareplots). We therefore proceeded to fit and report models using `glm` with the quasibinomial family and logit link because model objects produced by this function are compatible with numerous other R packages for summarizing and plotting effects.

Two of our four preregistered OLS regression models included interaction terms between signal and T1 *Belief* and *Action*, with the rationale that the signals might have reduced effect for individuals who indicate high levels of belief and helping at baseline. This rationale was less convincing for quasibinomial models using a logit link because the expected interaction -- reduced effect of signals in individuals with high baseline belief -- is inherent to logistic regression: the effect of a predictor on the outcome necessarily diminishes as the outcome approaches 1. Our theory does not predict an interaction beyond that already present in a logistic regression model, so we did not include an interaction term in the GLM models (but we still report OLS models with the interaction term).

To test our preregistered hypothesis that the likelihood of acting to help the victim would be largely mediated by a signal's positive impact on the participant's belief in the victim's need, we used the `mediation` package [@tingley2014] to fit a mediation model for *Depression* treatment vs. *Verbal request* control. See Figure \@ref(fig:medmodel).

```{r medmodel, fig.cap="Possible causal effects of the signal on helping behavior. A credible signal of need increases belief that the victim is telling the truth and needs help, which increase the likelihood of helping. The direct path from the signal to action represents other causal effects of the signal on observers, such as perceptions of the victim's emotional state (and other factors not measured in this study) that might alter observer behavior."}

x <- figname()
knitr::include_graphics('paths.svg')
```

For specifications of all regression models, see the SI. Our preregistration is here: <https://osf.io/g3s6n>

## Data availibility

The data are available at <http://doi.org/10.5281/zenodo.4637904>

# Results

```{r}
nationT0 <- table(signaling2020$sample)
removed <- sum(!signaling2020$CompleteSurvey)
removed_pct <- 100*signif(removed/nrow(signaling2020), 2)
nationT1 <- table(d$sample)
sexT1 <- table(d$Sex)
n_cond <- xtabs(~signal+vignette, d)

relstatus <- ifelse(str_detect(d$RelStatus, 'Previously'), 'Divorced', d$RelStatus)
relstatus <- ifelse(str_detect(relstatus, 'Unmarried'), 'Divorced', relstatus)
relstats <- table(relstatus)

```

The study was started by N=`r nrow(signaling2020)` participants who clicked the link to Qualtrics (`r nationT0['American']` US and `r nationT0['Indian']` India). After removing participants who did not finish the survey, failed attention checks, or moved through the study at an unrealistic pace (N=`r removed`, `r removed_pct`%), our final sample was N = `r nrow(d)` (`r nationT1['American']` US and `r nationT1['Indian']` India), with `r sexT1['Male']` males and `r sexT1['Female']` females, `r relstats['Married']` of whom were married or in a long-term relationship, `r relstats['Divorced']` who were divorced, `r relstats['Single']` who were single, and `r relstats['Widowed']` who were widowed. The median number of participants per condition was `r median(n_cond)` (min=`r min(n_cond)`, max=`r max(n_cond)`). For the number of participants in each condition, see Table \@ref(tab:conditions). For summary statistics, see Table \@ref(tab:sumtab). For the distributions of participants by age, income, and nationality, see Figure \@ref(fig:socioeconomic).

```{r sumtab}

d$minutes <- d$TotalTime/60
d$Children <- d$Sons + d$Daughters

vars <- c(
  Age = 'Age (years)',
  Income = 'Income (USD)',
  years_education = 'Education (years)',
  Children = 'Number of children',
  minutes = 'Time to complete (minutes)',
  T1Belief = 'T1 Belief',
  T1Action = 'T1 Action',
  T1Divide = 'T1 Division',
  T1LowMood = 'T1 Low mood',
  T1Manipulative = 'T1 Manipulative',
  T2Belief = 'T2 Belief',
  T2Action = 'T2 Action',
  T2Divide = 'T2 Division',
  T2LowMood = 'T2 Low mood',
  T2Manipulative = 'T2 Manipulative',
  T3Action = 'T3 Action'
)
kable(custom.summarize(d, vars), caption = "Summary statistics for study variables. Values that were on a 0-100 scale were rescaled to 0-1. Indian participants reported income in rupees, which we converted to USD at the current exchange rate (1 rupee = 0.014 USD).") %>% kable_styling()
```

## Distributions of beliefs and actions at Time 1

Across the four vignettes, mean belief of the victim was relatively low at baseline (T1), Mean = `r signif(mean(d$T1Belief, na.rm = T), 2)`, albeit with wide variation, SD = `r signif(sd(d$T1Belief, na.rm=T), 2)`; `r sum(d$T1Belief==0)` participants (`r 100*signif(sum(d$T1Belief==0)/nrow(d), 2)`%) rated their belief = 0, and `r sum(d$T1Belief==1)` participants (`r 100*signif(sum(d$T1Belief==1)/nrow(d), 2)`%) rated it as = 1. The distribution of likelihood of helping the victim (action) was similar, Mean = `r signif(mean(d$T1Action, na.rm = T), 2)`, SD = `r signif(sd(d$T1Action, na.rm=T), 2)`, with `r sum(d$T1Action==0)` participants (`r 100*signif(sum(d$T1Action==0)/nrow(d), 2)`%) rating their action = 0, and `r sum(d$T1Action==1)` participants (`r 100*signif(sum(d$T1Action==1)/nrow(d), 2)`%) rated action as = 1. Although not a pre-registered prediction, *T1 Belief* and *T1 Action* were highly correlated across the four vignettes, consistent with help being worth providing if the signaler's claims were true (see Figure \@ref(fig:t1plot)).

```{r t1plot, fig.width=10, fig.height=10, fig.cap="The distributions of *T1 Belief* vs.*T1 Action*, by vignette. Each dot is one participant. Original 0-100 values rescaled to 0-1. Lines fit by linear regression. Dot size indicates the number of overlapping points. Black dot is the mean of each variable."}
plot_T1Belief_T1Action 
ggsave(figname(), plot_T1Belief_T1Action)
```

## Confirmatory: Signals increase beliefs and actions in the predicted rank order

```{r}
sm1 <- summary(models$Margins$m1)
suicide_belief_eff <- signif(sm1$AME['signalSuicide attempt']/sd(d$T2Belief), 2)

sm3 <- summary(models$Margins$m3)
suicide_action_eff <- signif(sm3$AME['signalSuicide attempt']/sd(d$T2Belief), 2)

signal_belief_ttest <- t.test(100*d$T2Belief, 100*d$T1Belief, paired = T)
signal_action_ttest <- t.test(100*d$T2Action, 100*d$T1Action, paired = T)

```

As predicted, there was a strong statistically significant positive causal effect of the ordinal signal on *T2 Belief* and *T2 Action* (controlling for *T1 Belief* and *T1 Action*, respectively), with the effect of each signal increasing in the predicted rank order (verbal request, crying, mild depression, depression, suicide attempt). See Figure \@ref(fig:beliefeffect) and models m1 and m3 in Tables \@ref(tab:regressiontables) and \@ref(tab:anovatables). Results for the preregistered OLS models were very similar; see Figures \@ref(fig:preregbelief) and \@ref(fig:preregbeliefB). The within-subject effects of each signal are depicted in Figure \@ref(fig:rawbeliefactions2d).

```{r beliefeffect, fig.width=8, fig.height=4, fig.cap=paste0("The mean effects of the signals on *T2 Belief* and *T2 Action*, controlling for *T1 Belief* and *T1 Action*, respectively (original 0-100 scale rescaled to 0-1). Effects plotted for T1 values set to their median value (*T1 Belief*=", median(d$T1Belief, na.rm=T), ", *T1 Action*=", median(d$T1Action, na.rm=T), "). Fit using generalized linear regression models with the quasibinomial family. Bars are 95% CIs. Dots are *T2 Belief* and *T2 Action* values; dot size = number of overlapping data points. For coefficents, p-values, and other statistics, see models m1 and m3 in Tables \\@ref(tab:regressiontables) and \\@ref(tab:anovatables).")}
# plot_beliefeffect <-
#   (effectsPlots$m1plot + theme(axis.text.x = element_blank()))/effectsPlots$m3plot + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))

d_tmp <-
  d %>% 
  dplyr::select(signal, T2Action, T2Belief) %>% 
  pivot_longer(-signal, names_to = 'Outcome') %>% 
  mutate(Outcome = str_remove(Outcome, 'T2'))

d_m1m3 <- bind_rows(Belief = effectsPlots$m1plot$fit, Action = effectsPlots$m3plot$fit, .id = 'Outcome')

plot_actionbeliefeffect <-
  ggplot(d_m1m3, aes(visregFit, signal, xmin = visregLwr, xmax = visregUpr, colour = Outcome, group = Outcome)) +
  geom_errorbarh(lwd = 4.5, alpha = 0.75, position = position_dodge(width = 0.6), height = 0) +
  geom_point(position = position_dodge(width = 0.6), colour = 'black') +
  geom_count(data = d_tmp, aes(value, signal, colour = Outcome), inherit.aes = F, alpha = 0.25, position = position_dodge(width = 0.6)) +
  scale_color_viridis_d(option = 'B', begin = 0.2, end = 0.8) +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  scale_size_area() +
  guides(color=guide_legend(reverse = T, override.aes = list(size=3))) +
  labs(x = '\nT2 Belief/Action', y = '') +
  theme_minimal(15)

plot_actionbeliefeffect
ggsave(figname(), plot_actionbeliefeffect)
rm(d_tmp, d_m1m3)
```

The estimated marginal mean between-subjects effect of the high cost *Suicide attempt* vs. the *Verbal request* control on T2 *Belief* (averaging over all four vignettes and all values of T1 *Belief*) was an increase of `r fmt_margins(models$Margins$m1, 'signalSuicide attempt')` points on the original 0-100 point scale. The equivalent increase for T2 *Action* was `r fmt_margins(models$Margins$m3, 'signalSuicide attempt')` points. Given that the standard deviations of T2 *Belief* and *Action* are `r signif(sd(100*d$T2Belief), 2)` and `r signif(sd(100*d$T2Action), 2)`, respectively, these represent increases of `r suicide_belief_eff` and `r suicide_action_eff` SD's, respectively.

<!-- Within-subjects, the average effect across all signals on *Belief* at T2 compared to T1 was `r inline_ttest(signal_belief_ttest)`, and on *Action* was `r inline_ttest(signal_action_ttest)` (on the original 0-100 point scale). For the change in *Belief* and *Action* for, T1 to T2 for all participants in all conditions, see Figure @ref(fig:rawbeliefactions2d). -->

## Confirmatory: The effect of signals on action is largely mediated by belief

```{r medresults}
rslts <- all_mediations$results

suicidetotal <- 100*signif(rslts$point[rslts$signal == 'Suicide attempt' & rslts$Stat == 'Total'], 2)

suicidemed <- 100*signif(rslts$n.avg[rslts$signal == 'Suicide attempt' & rslts$Stat == 'Total'], 2)

depressiontotal <- 100*signif(rslts$point[rslts$signal == 'Depression' & rslts$Stat == 'Total'], 2)

depressionmed <- 100*signif(rslts$n.avg[rslts$signal == 'Depression' & rslts$Stat == 'Total'], 2)

```

According to our theoretical model, emotional signals that put cooperative benefits at risk increase observers' belief that the victim is genuinely in need. Increased belief then increases the likelihood of helping the victim (action). See Figure \@ref(fig:medmodel).

We fit the mediation and outcome models using GLM's with the binomial family as the mediation package does not support the quasibinomial family. The total effect of the *Suicide attempt* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r suicidetotal` points from T1 to T2 (on the original 100-point scale). Of this increase, `r suicidemed`% was mediated by the increased *Belief* that the victim was telling the truth. The total effect of the *Depression* signal (treatment) vs. *Verbal request* (control) on the likelihood of *Action* was to increase it by `r depressiontotal` points from T1 to T2. Of this increase, `r depressionmed`% was mediated by the increased *Belief* that the victim was telling the truth. See Figure \@ref(fig:mediation).

<!-- (Check if med4 or med4b is correct; they look identical to me). -->

```{r mediation, fig.width=8, fig.height=4, fig.cap="The effect of each signal (treatment) vs. Verbal request (control) on Action. The x-axis is the change in Action on the [0, 1] scale. The effects of the signals on Action are largely mediated by Belief (percent mediated on the right). The mediation model controlled for *T1 Belief*, and the outcome model controlled for *T1 Belief* and *T1 Action*. Both models were GLMs with the binomial family (the mediation package does not support the quasibinomial family). ACME: Average causal mediation effect. ADE: Average direct effect. Total: Total effect (ADE + ACME). Bars are 95% CIs."}
all_mediations$plot
ggsave(figname(), all_mediations$plot)
```

## Exploratory: Costlier signals decrease perceived manipulation and increase perceived low mood

The mean change in the *Low mood* and *Manipulative* variables from T1 to T2 for each signal in each vignette reveals a large decrease in inferred *Manipulation* for costly signals, and a large increase in inferred *Low mood* (with the exception of the Thwarted marriage vignette, in which changes are small). See Figure \@ref(fig:emotionsplot).

```{r emotionsplot, fig.width=10, fig.height=8, fig.cap="The within-subjects effect of the signals on participants' inferred *Manipulative* and *LowMood* emotional states of the victim, from T1 (arrow base) to T2 (arrow head), for each signal in each of the four vignettes. Low mood: higher values indicate lower perceived mood. Manipulative: higher values indicate higher perceived manipulativeness. Thick arrows: mean change. Thin arrows: means of 500 bootstrap resamples."}
plot_emotions
ggsave(figname(), plot_emotions)
```

```{r}
mentallyIll <- signif(100*mean(d$T1MentallyIll), 2)

miT2 <- 
  as_tibble(xtabs(~T2MentallyIll + signal + vignette, d)) %>%
  mutate(T2MentallyIll = as.numeric(T2MentallyIll)) %>% 
  group_by(signal) %>% 
  summarise(percent = signif(100*sum(T2MentallyIll*n)/sum(n), 2))

```


The proportion of participants inferring each emotion in each vignette at T1 and after each signal at T2 is depicted in Figure \@ref(fig:emotionheatmaps), and the change in proportions from T1 to T2 in Figure \@ref(fig:diffemotions). The increase in inferred depression in the *Depression* and *Suicide attempt* conditions, but not the *Verbal request* and *Crying* conditions, helps validate these signals, as does the increase in inferred suicidality in the *Suicide attempt* condition.

At baseline (T1), `r mentallyIll`% of participants thought the victim was mentally ill, which decreased slightly at T2 in the *Verbal request* and *Crying* conditions (`r miT2$percent[miT2$signal=='Crying']`%), and then increased with signal cost to `r miT2$percent[miT2$signal=='Mild depression']`% in *Mild depression*, `r miT2$percent[miT2$signal=='Depression']`% in *Depression*, and `r miT2$percent[miT2$signal=='Suicide attempt']`% in *Suicide attempt*, with some variation by vignette. See Figures \@ref(fig:mentallyillT1) and \@ref(fig:mentallyillT2).

## Exploratory: Signal effects differ by vignette

```{r}

thwart_sv_bdiff <- visreg_diff(models$Model$m2, 'Thwarted marriage', 'Suicide attempt', 'Verbal request', d)

thwart_sv_adiff <- visreg_diff(models$Model$m4, 'Thwarted marriage', 'Suicide attempt', 'Verbal request', d)

brother_sv_bdiff <- visreg_diff(models$Model$m2, 'Brother-in-law', 'Suicide attempt', 'Verbal request', d)

brother_sv_adiff <- visreg_diff(models$Model$m4, 'Brother-in-law', 'Suicide attempt', 'Verbal request', d)

```

```{r}
bil_b_suicide <- signif(effect_sizes$est[effect_sizes$id=="Suicide attempt,Brother-in-law" & effect_sizes$Outcome=="Belief"], 2)

rom_b_crying <- signif(effect_sizes$est[effect_sizes$id=="Crying,Romantic partner" & effect_sizes$Outcome=="Belief"], 2)

bil_a_suicide <- signif(effect_sizes2$effsize[effect_sizes2$id=="Suicide attempt,Brother-in-law" & effect_sizes2$Outcome=="Action"], 2)

bil_a_verbal <- signif(effect_sizes2$effsize[effect_sizes2$id=="Verbal request,Brother-in-law" & effect_sizes2$Outcome=="Action"], 2)

```

The effects of the signals on rated beliefs and actions differed substantially by vignette, which we display in three ways for each signal in each vignette: estimated cumulative distributions plots, which show the entire distributions of rated beliefs and actions, including which conditions have high fractions of 0's and 1's (Figures \@ref(fig:ecdf)); regression estimates of the mean effects of the signals on beliefs and actions by vignette (Figure \@ref(fig:beliefvignette)); and between- and within-subjects signal effect sizes by vignette (Figures \@ref(fig:effectsizes) and \@ref(fig:effectsizes2)).

```{r ecdf, fig.width=11, fig.height=6, fig.cap="Empirical cummulative distribution functions for Time 2 Belief (top) and Action (bottom) compared to their Time 1 baseline values (red), by signal and vignette. Y-values indicate the fraction of all ratings equal to or less than a given x-value. Ratings rescaled to [0, 1]. Dots indicate median values of Belief and Action, by signal and vignette."}
ecdf_plot
ggsave(figname(), ecdf_plot)
```

```{r beliefvignette, fig.width=8, fig.height=6, fig.cap=paste0("The effect of the signals on *T2 Belief* and *T2 Action* in each vignette, controlling for T1 values of Belief and Action, respectively. Ratings rescaled to [0, 1]. Effects plotted for T1 values set to their median value (*T1 Belief*=", median(d$T1Belief, na.rm=T), ", *T1 Action*=", median(d$T1Action, na.rm=T), "). Fit using generalized linear regression models with the quasibinomial family. Bars are 95% CIs. For coefficents, p-values, and other statistics, see models m2 and m4 in Tables \\@ref(tab:regressiontables) and \\@ref(tab:anovatables).")}
# plot_beliefvignette <-
#   (effectsPlots$m2plot + theme(axis.text.x = element_blank()))/effectsPlots$m4plot + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
  
# d_tmp <-
#   d %>% 
#   dplyr::select(signal, vignette, T2Action, T2Belief) %>% 
#   pivot_longer(-c(signal, vignette), names_to = 'Outcome', values_to = 'visregFit') %>% 
#   mutate(
#     Outcome = str_remove(Outcome, 'T2'),
#     visregLwr = 0,
#     visregUpr = 0
#     )

d_m2m4 <- 
  bind_rows(Belief = effectsPlots$m2plot$fit, Action = effectsPlots$m4plot$fit, .id = 'Outcome') %>% 
  dplyr::select(Outcome, signal, vignette, visregFit, visregLwr, visregUpr)

plot_beliefvignette <-
  ggplot(d_m2m4, aes(signal, visregFit, ymin = visregLwr, ymax = visregUpr, colour = Outcome, group = Outcome)) +
  geom_errorbar(lwd = 3, alpha = 0.75, position = position_dodge(width = 0.6), width = 0) +
  geom_point(position = position_dodge(width = 0.6), colour = 'black') +
  scale_color_viridis_d(option = 'B', begin = 0.2, end = 0.8) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  guides(color=guide_legend(reverse = F, override.aes = list(size=3))) +
  labs(x = '', y = '') +
  facet_wrap(~vignette, ncol = 4, strip.position = 'top') +
  theme_bw(15) + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
    legend.position = 'top'
    )

plot_beliefvignette
ggsave(figname(), plot_beliefvignette)
```

We note the following patterns in Figure \@ref(fig:ecdf), which we will return to in the Discussion section. First, as we intended, the distributions of *Belief* and *Action* in the *Verbal request* control condition were very similar to their distributions at T1 baseline across vignettes. In the "Basketball coach" and "Brother-in-law" vignettes, though, their distributions in the *Verbal request* condition were shifted to somewhat lower values relative to baseline (i.e., the verbal request slightly *reduced* belief and action in those vignettes). Second, in the "Romantic partner" and "Brother-in-law" vignettes, the effect of *Crying* differed little from *Verbal request*, but in the "Basketball coach" vignette, it differed little from the *Depression* and *Suicide attempt* signals. Third, the effect of the *Suicide attempt* signal on *Belief* was similar to that of *Depression* across all vignettes, but had a noticeably greater effect on *Action* in the "Romantic partner" and "Brother-in-law" vignettes. Fourth, there was little difference between the effect of the *Mild depression* vs. *Depression* signals.

The largest between-subjects effect was *Suicide attempt* vs. *Verbal request* control on *Belief* in the "Brother-in-law" vignette, Cohen's d = `r bil_b_suicide`, and the smallest was the effect of *Crying* on *Belief* in the "Romantic partner" vignette, Cohen's d =`r rom_b_crying`. The largest within-subjects effect was *Suicide attempt* on *Action* at T2 vs. T1 in the "Brother-in-law" vignette, Cohen's d = `r bil_a_suicide`, and the smallest was the *negative* effect of *Verbal request* on *Action* at T2 vs. T1 in the "Brother-in-law" vignette, Cohen's d =`r bil_a_verbal`.

## Exploratory: Sociodemographic associations

The sociodemographic variables were strongly confounded with nationality (Indian participants were younger, lower income, with more years of education than US; see Figure \@ref(fig:socioeconomic)), which, in turn, were confounded with vignette (responses in the Indian Thwarted marriage vignette differed substantially from those in the US vignettes). We therefore conducted our exploration of the sociodemographic variables separately by nationality.

In the US sample, female participants were more likely to believe and help the victim than male participants, and younger participants were more likely to believe and help the victim than older participants. We found no significant effects of *Income*, *Education*, relationship status (e.g., married, single), or number of sons or daughters on *Belief* or *Action*. See Figure \@ref(fig:demographicsignal).

```{r demographicsignal, fig.width=11, fig.height=8, fig.cap="Sociodemographic associations with *T2 Belief* (top) and *T2 Action* (bottom) in US participants. Each model controlled for *T1 Belief* or *T1 Action*, vignette, Education, and Income. Significant associations displayed here. Estimated using generalized linear regression models with the quasibinomial family. Grey bars are 95% CIs."}

plot_demographicsignal <-
  effectsPlots_US$m12plot + 
    theme(axis.text.x = element_blank()) +
effectsPlots_US$m12bplot + 
    theme(
      axis.text = element_blank(), 
      axis.title.y = element_blank(), 
      axis.title.x = element_blank()
      ) + 
effectsPlots_US$m13plot + 
effectsPlots_US$m13bplot +
    theme(axis.text.y = element_blank(), axis.title.y = element_blank())

plot_demographicsignal
ggsave(figname(), plot_demographicsignal)
```

In the Indian sample and vignette, in contrast, males were more likely to help the victim than females, those with more years of education were less likely to believe the victim, and there was a marked increase in likelihood of helping among older individuals in the suicide signal condition. We found no significant associations with age or income. See Figure \@ref(fig:indiademographics).

```{r indiademographics, fig.width=10, fig.height=8, fig.cap="Sociodemographic associations with *T2 Belief* and *T2 Action* in Indian participants (Thwarted marriage vignette). Each model included *T1 Belief* or *T1 Action*, Income, Age, Sex, Years of education, and signal. Significant associations displayed here. Estimated using generalized linear regression models with the quasibinomial family. Grey bars are 95% CIs."}
plot_indiademo<-
  (effectsPlots_India$m12plot + xlab('Education (years)') + effectsPlots_India$m13plot) / effectsPlots_India$m17plot

plot_indiademo
ggsave(figname(), plot_indiademo)
```

## Validity check: T3 increase in Action with proof that victim was telling the truth

```{r}
t2t3actionttest <- t.test(d$T3Action, d$T2Action, paired = T)
```

As a partial check on the validity of our methods and results, at T3 we presented participants with strong evidence that the victim was telling the truth, and asked them to again rate their likelihood of acting to help the victim (*T3 Action*). Compared to the mean *T2 Action* (M=`r signif(mean(d$T2Action, na.rm=T), 2)`), mean *T3 Action* increased substantially, `r inline_ttest(t2t3actionttest)`. Across the US vignettes, likelihood of acting at T3 increased to near-ceiling (M=`r signif(mean(d$T3Action[d$vignette != 'Thwarted marriage']), 2)`), regardless of participants' *T2 Action* values, indicating that propensity to act was indeed contingent on believing the victim. In the Indian Thwarted marriage vignette, however, there was little change in *T3 Action* compared to *T2 Action* (Figure \@ref(fig:t3action)), perhaps because Indian participants' beliefs and actions were relatively insensitive to the signals to begin with (Figures \@ref(fig:ecdf) and \@ref(fig:beliefvignette)).

# Discussion

As predicted, in vignettes involving conflicts of interest and private information about the need for help, costly signals of need increased participants' belief in the victim and their likelihood of helping her, with the increase in perceived need and likelihood of helping increasing monotonically with signal cost. As predicted, the increase in likelihood of helping was largely mediated by the increase in belief in the victim. In an exploratory analysis, costlier signals also decreased perceptions that the victim was manipulative.

These effects were largest in the "brother-in-law" and "romantic partner" vignettes, both of which involved claims of assault against participants' imagined daughters, and smaller in the "basketball coach" and the "thwarted marriage" vignettes. The small effect in the "basketball coach" vignette might have been due to the lack of relatedness between participants and the victim (the star player), or how participants weighted the costs of suspending the coach vs. punishing a potentially innocent person.

The small signal effect in the "thwarted marriage" vignette, which involved the Indian sample, could indicate that our results do not generalize across cultures, undermining our adaptationist hypothesis. It could also reflect our poor understanding of contemporary Indian culture regarding dowry (the effect was larger in older participants). Baseline belief in the older daughter, and likelihood of helping her, was relatively high at baseline (`r signif(100*mean(d$T1Belief[d$vignette=='Thwarted marriage']), 2)`%) compared to victims in the other vignettes. Private information and conflict therefore probably played a smaller role and thus costly signals were less necessary. We observed a similar pattern in our pilot study, in which baseline belief in the victim's need was high, and costly signals had smaller effects than they did in the current study. Additionally, supporting the older daughter came at the cost of one's younger daughter, which may also help explain the relatively small signal effects. A final consideration is that data from the Indian sample appeared to be of lower quality, limiting our confidence in any of these interpretations (see limitations for more information).

In the "brother-in-law" and "romantic partner" scenarios, *Crying* had little effect on the magnitude of pro-victim responses relative to *Verbal request*, suggesting it was not costly enough to serve as a reliable signal in times of substantial conflicts of interests. In contrast, both *Depression* conditions increased support, albeit to similar degrees. One potential reason for this is the increase in symptoms from mild depression to depression was not costly enough to result in a large effect, as the symptoms differed little in severity (e.g., grades dropping from As to Bs in *Mild depression* vs. Cs in *Depression*). Such small changes may be less impactful in vignettes than in real-life, where the effects of signaling may increase in severity as they persist over a lengthy period of time.

The effect of *Suicide attempt* on belief at T2 was similar to the *Depression* conditions across vignettes, but it resulted in greater likelihood of support. One interpretation is that although some participants did not believe the victim's story, her signal nevertheless convinced them that she needed help. For example, maybe the brother-in-law did not assault her, but the presence of his family in her home was causing genuine distress. Support for this interpretation comes from our mediation analyses, which showed that the likelihood of help was largely, but not entirely, mediated by signal's effect on belief in need.

There were minor associations of age and sex with belief and action in the US participants, with both being higher among females and younger individuals. The US vignettes all involved assaults against young women, which might have been more salient to female and younger participants. In the Indian sample, high baseline belief and action were somewhat lower among those with more education and among females, respectively. Costlier signals, suicidality in particular, had a larger effect among older individuals, perhaps because older individuals were more likely to have children of marriageable age, like the victim in the vignette.

Finally, after the T3 evidence that the victim was telling the truth, likelihood of helping by the US participants increased to near ceiling, an effect that helped validate our vignettes. Among Indian participants, in contrast, participants only slightly increased their likelihood of helping from their T2 level. One interpretation of the latter is that Indian participants tended to believe the older daughter anyway, so their decision to help was not changed by additional information.

<!-- Overall, our results on the connection between belief in need and likelihood of help are consistent with costly signaling and bargaining approaches to depression and suicide. At T1, *Belief* and *Action* were highly correlated yet distinct; while at T2, the effect of signal on both *Belief* and *Action* was to increase the magnitude of pro-signaler responses, with the effect of signal on *T2 Action* being largely mediated by signal's effect on belief in need. This relationship between belief in need and likelihood of support and the importance of signal cost should be expected in scenarios when a target of a signal of need would benefit from helping if the signaler's need was genuine but face substantial net costs if it was not. Importantly, such scenarios appear to be common among modern humans and have likely been since individuals could increase their inclusive fitness through supporting kin or important cooperative partners who could return to providing benefits to supporters once their situation returned to normal [@SugiyamaSugiyama2003; @tooby1996]. -->

<!-- Although this study was not designed to examine whether the costs of signaling would be made up for by the increased support stemming from depression and suicide attempts directly, the results are consistent with this hypothesis. In general, the costs required to honestly convey one's qualities should be expected to increase with the potential payoffs of deception and the number of defectors within the broader population [@Rosenthal1993], with the vignettes in this study emphasizing the existence of private information and sizable conflicts of interest to increase the perceived risk of being taken advantage of. In similar real-world scenarios, this difficulty of signaling honest need would coincide with the substantial fitness-relevant benefits of preventing future assaults or being able to marry the partner of one's choice. These benefits coupled with the fact that depression and suicide attempts become relatively less costly to one's fitness prospects with increased need [@Hagen2003], suggest that the increased support seen in this study may be worth the costs required to signal need more often than is often appreciated when counting the entire costs of depression as the signal rather than only the upregulation of symptoms for a communicative purpose. -->

## Limitations

This study has less ecological validity than real-world observations of depressed individuals interacting with their social partners, which might have biased results in a pro-signaler direction if the lack of real costs of helping made support feel less costly or if there was a social desirability bias toward helping [@grimm2010]. It may have also biased participants against helping if they could not fully imagine the characters in the story as kin or interdependent partners, and the survey's short duration may have weakened the strength of the costlier signals as bargaining tools.

Our design did not include vignettes with male signalers. For this reason, we have no data on the possibility of sex differences in the effectiveness of the signaling strategies examined. Although not predicted theoretically, such differences are possible if the costs of signaling vary between the sexes due to differential access to alternative bargaining strategies [@HagenRosenstrom2016] or if one sex tends to suffer greater negative reputational effects when displaying the emotions and behaviors in the vignettes. It is also possible the costliness of the situations presented in the vignettes differ by sex.

Compared to the US sample, far more Indian participants failed our attention checks, which is consistent with botting, unfamiliarity with English, or low-effort responses [@kennedy2020]. If this high failure rate indicates lower-quality responses among those who passed the attention checks, the weak signal effect in the thwarted marriage vignette may simply be due to greater noise rather than differences in the scenario or the effectiveness of the signals compared to those in the US. Another concern relevant to all vignettes is that our decision to anchor the T1 sliders at 0 may have resulted in participants being more likely to report extreme values.

Finally, we adopted game theory models of bargaining with incomplete information as our theoretical framework, but there are many other models of credible signaling [e.g., @szamado2011], including for need [@szamado2019] and suicidality [@Rosenthal1993]. If depression and suicidality involve signaling, it might be better explained by a different model.

<!-- Slider anchor  -->
<!-- Regex: stopped in SI -->
<!-- T\d \w+ -->

# Conclusion

Depression is costly, sometimes resulting in death by suicide. Our results suggest that these costs, which mainstream theories take as evidence of brain dysfunction [@SymeHagen2021], function to help victims of adversity elicit support when their true level of need is private information and they have conflicts with social partners. These findings align with real-world evidence that depression simultaneously elicits positive and negative responses from social partners [@SymeHagen2021]. If the bargaining model is correct, depression is an important social problem, not a medical one [@syme2020].

<!-- This coupled with , suggests that more attention should be paid to the potential for positive responses towards depressed individuals, especially from social partners who receive substantial benefits from the depressed or suicidal individual. -->
 
 <!-- The need to explain suicide's connection to depression is often overlooked by explanations for major depression, particularly those which propose non-signaling functions. One explanation for their connection, is that both suicidality and depression may function to limit negative fitness outcomes in times of substantial adversity [@andrews2006; @Hagen2002; @Hagen2003; @HagenEtAl2008; @nock2008; @Rosenthal1993; @Stengel1956; @SymeEtAl2016], with costlier signals being required to elicit support when the payoffs to faking need are high enough so that less costly signals are not reliable indicators of one's need [@Rosenthal1993]. For this hypothesis to be correct, instrumental support would need to be provided frequently enough with large enough impacts so that the costs of signaling are outweighed by the benefits on average. In support of this, we found exposure to depression and suicide attempts resulted in greater belief in need and increased likelihood of support in fictional scenarios. -->

# Acknowledgements

This study was funded through a Washington State University, Vancouver Research Mini Grant. We thank Anne Pisor, Darcy Bird, and members of the Hagen lab group for numerous helpful comments. We also thank XX and YY [pending permission to acknowledge] for feedback on the Thwarted marriage vignette.

<!-- # To do: -->

<!-- -   victim or signaler? -->
<!-- -   add anger -->
<!-- -   Intro "State the objectives of the work and provide an adequate background, avoiding a detailed literature survey or a summary of the results." -->
<!-- -   Discussion " This should explore the significance of the results of the work, not repeat them. A combined Results and Discussion section is often appropriate. Avoid extensive citations and discussion of published literature." -->
<!-- -   mention asking for prefered sex -->

<!-- # Notes from my meeting with Anne -->

<!-- -   Table detailing the cooperative endeavor (maybe word as boundedness of fate), conflict of interest, and what the private information is about would be useful. -->
<!-- -   Calling verbal request control when it's a treatment? -->
<!-- -   Can shift much of the pilot studies info to SI with some detail post line 279 in draft she saw. -->
<!-- -   Cumulative distribution plot: needs more info in captions. could benefit from fig6a and fig6b (confusion over x and y axes) -->

# Supplementary information

## Pilot studies

We conducted two pilot studies, one on MTurk with a large sample size (N=`r nrow(signalingdata2018b)`), and one on [reddit.com/r/SampleSize](https://reddit.com/r/SampleSize) with a small sample size (N=28). Both studies used the experimental pre-test, post-test design used in the current study.

In the MTurk study, participants read a vignette that put them in situations in which an imagined sister was requesting a loan of \$50,000 from the participant that was originally saved for the participant's imagined daughter's college education. The sister claims the money is needed for medical treatment for her own child, but she has incentives to deceive. The manipulations were signals very similar to those used in this study, and the outcomes were belief that the sister really needed the money and intention to loan her the money. See XX for the full write-up. See <https://osf.io/3rg8b> for preregistration of the pilot study. See <http://doi.org/10.5281/zenodo.4637883> for pilot data.

```{r piloteffects, fig.width=10, fig.cap="Pilot study effect of the signals. A. Effect of signal on Belief at T2, controlling for Belief at T1. B: Proportion of participants who inferred the victim was mentally ill. Fit with a generalized linear model with the quasibinomial family."}
pilot_results$effectplot + (pilot_results$mentalplot + theme(axis.text.y = element_blank())) + plot_annotation(tag_levels = 'A')
```

In the small reddit.com study, participants viewed and responded to two vignettes used here, each with a different signal.

We learned six major lessons from the MTurk pilot study that influenced the design of the current study. First, as predicted, belief in the sister's need was highest in the *Depression* condition and lowest in the *Verbal request* condition, giving us confidence to continue this research (see Figure \@ref(fig:piloteffects)). Second, to rule out that believability and willingness to help were simply artifacts of the sister's psychiatric distress, we included a "signaling" condition in which the sister exhibited schizophrenic symptoms. As predicted, believability and willingness to help in this condition were dramatically lower than in any other condition, ruling out this alternative explanation. We therefore did not include the schizophrenic condition in the current study. Third, baseline (T1) believability was very high (mean = `r signif(pilot_results$meanT1belief, 2)` and median = `r signif(pilot_results$medianT1belief, 2)` on a 0-100 point scale), leaving little scope for signals to increase believability. In the current study we therefore endeavored to provide details in the vignettes that would undermine the victim's believability at baseline. Fourth, contrary to our hypothesis, the *Suicide threat* and *Suicide attempt* signals reduced believability relative to *Depression*. We surmised that the sister's adversity -- a sick child -- was not sufficient to justify a suicide attempt, thus undermining her believability. In the current study the signaler herself is therefore the victim, and the forms of adversity -- physical and sexual assault and thwarted and forced marriage -- are those that are frequently associated with suicidality. Fifth, the content of the vignette matters, so instead of a single vignette we decided to use four different vignettes to help determine the extent to which signaling effects were sensitive to seemingly irrelevant details of the vignettes. Sixth, in a pre-test post-test design, the control condition should be identical to the treatment condition in every respect except for the factor being tested, which in this study was signal cost. Post hoc, we realized that the control condition in the pilot study differed substantially from the treatment condition beyond signal cost: in the treatment conditions, participants encountered the sister twice: once when she requests help, and a second time when she signals. In our faulty control, participants only encountered the sister once, when she requests help, with no second encounter. In this study we therefore ensured that the control condition (Verbal request) was identical to the treatment conditions except for signal cost.

In the reddit.com pilot study, participants viewed two vignettes (with different signals in each one), and responded with lower belief to the second vignette compared to the first regardless of the vignette or signal. To avoid such order effects, which are a common problem in vignette studies that use multiple vignettes [@su2020], we decided to present each participant in the current study with a single vignette and signal.

## Power from MTurk pilot study

We estimated power using our MTurk pilot study data to compare the effect of the *Mild depression* signal vs. the *Verbal request* control on *Belief* (that the sister needed the money) at T2 controlling for *Belief* at T1 in a linear regression model. Specifically, we resampled the pilot data with replacement, fit the linear regression model, and extracted the p-value for the signal coefficient, repeating this 2000 times for each sample size ranging from N=20 to N=200. Power was computed as the proportion of p-values \< 0.05. Power was about 80% for a sample size of about `r pwr80_90[[1,1]]`, and was about 90% for a sample size of about `r pwr80_90[[1,2]]`. See Figure \@ref(fig:power).

```{r power, fig.cap="Power curve for between-subjects effect of Mild depression vs. Verbal request signal on T2 Belief, controlling for T1 Belief."}

smplesize <- nrow(d)/10

ggplot(power_curve, aes(sample_size, power)) +
  geom_line() +
  geom_vline(xintercept = smplesize, linetype = 'dotted') +
  annotate(geom = 'text', x = smplesize + 3, y = 0.75, label = 'Current study sample size', hjust = 0) +
  labs(x = '\n Sample size', y = 'Power\n') +
  theme_minimal(15) + theme(axis.title.y = element_text(angle = 0))
```

<!-- Second, for participants in the *Depression* condition only, we compared their T2 Belief to their T1 Belief using a paired t-test (in the full pilot data, *Depression* increased belief by about 8 points on a 100 point scale). We again resampled our pilot data 2000 times for each sample size from N=20 to N=200. Power was about 80% for a sample size of 125, and about 90% for a sample size of 175. -->

```{r power2, fig.cap="Power curve for within-subjects effect of Depression signal on *T2 Belief* relative to *T1 Belief*."}
  # ggplot(power_curve2, aes(sample_size, power)) +
  #   geom_line() +
  #   labs(x = '\n Sample size', y = 'Power\n') +
  #   theme_minimal(15)
```

## Vignettes of the current study (main bodies and signals)

### Basketball coach

Imagine yourself in the following scenario:

You are a college athletic director at a university with a long history of success in women's basketball. Unfortunately, despite quality players, the team struggled for the first three years of your leadership. You came close to losing your job but were given two more years to turn things around. You immediately hired a new head coach who you had spotted coaching a low-ranked college team that punched above its weight. Two years later, your team is one of the best in the country and has attracted a top recruit who has become the star player on your team and is probably the best player in the country. You have been told that if the team makes it to the final rounds of the playoff tournament, the university will offer you a high paying contract extension.

Your star player is hard to deal with, however. She is constantly jealous of other players, and often comes to your office to complain that the coach unfairly limits her playing time, claiming this will ruin her chances of going professional. The star player has also posted self-serving accounts of your discussions with her on social media.

A few days ago, after the coach started a new player in one game, the star player went to the police and accused the coach, a physically powerful man, of cornering her in the locker room one night and forcibly kissing her and grabbing her body. The coach strongly denies the allegation, and the police said they do not have any evidence to corroborate the player's accusation.

You reached out to other players on the team, all of whom say that the relationship between the coach and star player has always seemed professional, and they are shocked by the allegation. None of the other players have any complaints about the coach. On the contrary, the other players praise the coach for leading everyone into the playoffs. They mention that the star player is a prima donna, but they are worried that if either the coach or the star player miss the playoffs, it will hurt everyone's chances of making it to the championships.

You find out that the star player then told the university president that she feared the coach might attack her again, and asked her to replace the coach with one of his assistants, who you know wants the star player to have more playing time. The president refused, saying there was no evidence of wrongdoing.

#### Signals

##### Verbal request

A week later the coach informs you that the star player's performance in games is excellent, helping the team's performance and her reputation with pro scouts. Later that day, there is a formal luncheon at the athlete dorm to introduce the coaches and members of each team to the alumni. The alumni are especially interested in meeting the basketball coach and members of the basketball team because they are in the playoffs. You notice your star player is not there, but you see her roommate, a member of the swim team. She tells you that you should talk to your star player in person, and she leads you to their dorm room.

When you arrive, the two of you find her sitting on her bed. She appears fit, well-rested, and her room is clean. She looks calm with her hands folded as she asks you to replace the coach, saying she can't attend the luncheon if he is there.

##### Crying

A week later the coach informs you that the star player's performance in games is excellent, helping the team's performance and her reputation with pro scouts. Later that day, there is a formal luncheon at the athlete dorm to introduce the coaches and members of each team to the alumni. The alumni are especially interested in meeting the basketball coach and members of the basketball team because they are in the playoffs. You notice your star player is not there, but you see her roommate, a member of the swim team. She tells you that you should talk to your star player in person, and she leads you to their dorm room.

When you arrive, the two of you find her sitting on her bed. She appears fit and well-rested, and her room is clean. She looks sad and begins sobbing as she asks you to replace the coach, saying she can't attend the luncheon if he is there.

##### Mild-depression (crying with behavioral changes)

A week later the coach informs you that the star player's performance in games is OK and not harming the team's performance and her reputation with pro scouts. Later that day, there is a formal luncheon at the athlete dorm to introduce the coaches and members of each team to the alumni. The alumni are especially interested in meeting the basketball coach and members of the basketball team because they are in the playoffs. You notice your star player is not there, but you see her roommate, a member of the swim team. She tells you that you should talk to your star player in person, and she leads you to their dorm room.

When you arrive, the two of you find her sitting on her bed. Her room is a little messy and she appears a little less fit than you remember, while also seeming slightly tired. She looks sad and begins sobbing as she asks you to replace the coach, saying she can't attend the luncheon if he is there.

##### Depression

A week later the coach informs you that the star player's performance in games is poor, harming the team's performance and her reputation with pro scouts. Later that day, there is a formal luncheon at the athlete dorm to introduce the coaches and members of each team to the alumni. The alumni are especially interested in meeting the basketball coach and members of the basketball team because they are in the playoffs. You notice your star player is not there, but you see her roommate, a member of the swim team. She tells you that you should talk to your star player in person, and she leads you to their dorm room.

When you arrive, the two of you find her slouched on her bed. Her room is a mess, and you notice she has lost weight and looks fatigued. She looks sad and begins sobbing and wringing her hands as she asks you to replace the coach, saying she can't attend the luncheon if he is there.

##### Suicide attempt

A week later the coach informs you that the star player's performance in games is poor, harming the team's performance and her reputation with pro scouts. Later that day, there is a formal luncheon at the athlete dorm to introduce the coaches and members of each team to the alumni. The alumni are especially interested in meeting the basketball coach and members of the basketball team because they are in the playoffs. You notice your star player is not there, but you see her roommate, a member of the swim team. She tells you that you should talk to your star player in person, and she leads you to their dorm room.

When you arrive, the two of you find her passed out on her bed. The room is a mess, and you notice she has lost weight. Her roommate tries to wake her up, but she doesn't respond. You see an empty bottle of OxyContin painkillers on the table and immediately call 911. That evening you visit her in the hospital. Her doctor tells you she almost died, and then leaves the two of you alone. The star player begins sobbing as she asks you to replace the coach.

### Brother-in-law

Imagine yourself in the following scenario:

You are the single parent of a 15-year-old girl. Your relationship with her has always been full of conflict. She is very self-centered, and the two of you fight frequently.

Recently, your sister lost her house in a fire and was having difficulty finding a place for her family to live because many other families also lost their homes. You are very close with your sister and brother-in-law, who helped you escape your abusive ex-partner. You have a large house with several extra bedrooms, so you decide to let your sister, your brother-in-law, and your teenage niece move in while they look for more permanent housing.

(page break)

Unfortunately, your daughter and your niece are not getting along. They have to share a bathroom, and your daughter complains constantly that she is in there too long. Your daughter is also jealous that her cousin has expensive clothes, a designer handbag, and a new iPhone, all of which are nicer than your daughter's.

One day your daughter had a party at your house with many of her friends. You noticed that your niece was very popular and got a lot of attention, especially from your daughter's friends and boyfriend. You could tell that your daughter was upset. After the party, your daughter tells you that her cousin is a bitch who is trying to steal her boyfriend. However, you saw that your niece was trying to avoid interacting with the boyfriend.

(page break)

Later, your daughter complained that your brother-in-law was looking at her body. A few days later, while grocery shopping, you got a text from your daughter saying that your brother-in-law tried to get into the bathroom while she was showering and came into her room while she was changing. She says that she does not feel safe and wants your sister and her family to leave. However, you know your daughter will exaggerate to get her way.

The next week, your daughter comes to you saying your brother-in-law groped her as they passed in the hallway and that she wants your sister's family to leave immediately. You confront your brother-in-law due to the severity of your daughter's allegation. He strongly denies it.

That night, you call your mother who says your daughter has long been jealous of her cousin. She reminds you that you've taken your daughter's side before and been burned when you found out your daughter wasn't telling the whole truth. She reminds you of the difficult situation your sister's family is in.

#### Signals

##### Verbal request

Your daughter has appeared normal since her claims about your brother-in-law. She has still looked fit and put the same amount of effort into her appearance, always wearing clean clothes and spending time on her hair. She has also slept about the same amount of time as normal and remained active throughout the day.

Her demeanor has also remained normal. She has generally appeared calm and has been able to maintain her concentration. She has also done her homework on time and maintained an A in all her classes as she always has in the past.

When the two of you talk about your brother-in-law, she remains calm and asks you to kick him out.

##### Crying

Your daughter has appeared normal since her claims about your brother-in-law. She has still looked fit and put the same amount of effort into her appearance, always wearing clean clothes and spending time on her hair. She has also slept about the same amount of time as normal and remained active throughout the day.

Her demeanor has also remained normal. She has generally appeared calm and has been able to maintain her concentration. She has also done her homework on time and maintained an A in all her classes as she always has in the past.

When the two of you talk about your brother-in-law, she starts sobbing and asks you to kick him out.

##### Mild-depression (crying with behavioral changes)

Your daughter has appeared sad since her claims about your brother-in-law. She has looked a little less fit and has put less effort into her appearance, occasionally wearing dirty clothes and sometimes not doing her hair. She has also sometimes woken up late and has seemed a little less active than normal.

Her demeanor has also changed slightly. She has occasionally gotten distressed and has seemed distracted. She has also missed a few assignments and her grades dropped from straight A's to a B average for the first time in her life.

When the two of you talk about your brother-in-law, she starts sobbing and asks you to kick him out.

##### Depression

Your daughter started to appear noticeably different since her claims about your brother-in-law. She has lost weight and has been putting no effort into her appearance, often wearing dirty clothes and rarely doing her hair. She has also slept more than normal but still seemed tired and inactive.

Her demeanor also changed drastically. She has often gotten distressed and has had difficulty maintaining her concentration. She has also missed multiple assignments, and her grades dropped from straight A's to a C average for the first time in her life.

When the two of you talk about your brother-in-law, she starts sobbing and asks you to kick him out.

##### Suicide attempt

Your daughter started to appear noticeably different since her claims about your brother-in-law. She has lost weight and has been putting no effort into her appearance, often wearing dirty clothes and rarely doing her hair. She has also slept more than normal but still seemed tired and inactive.

Her demeanor also changed drastically. She has often gotten distressed and has had difficulty maintaining her concentration. She has also missed multiple assignments, and her grades dropped from straight A's to a C average for the first time in her life.

When the two of you talk about your brother-in-law, she starts sobbing and asks you to kick him out. Then, this past week, you and your daughter went through a traumatic experience when your daughter attempted suicide. You rushed her to the hospital, where she was placed in critical but stable condition.

### Romantic Partner (pronouns match sex participants reported primary attraction to)

Imagine yourself in the following scenario:

You are a single parent of a 13-year-old girl. The two of you have lived alone together for the past five years since your breakup with her [father/mother]. During this time, finding potential partners has been difficult.

Recently, you have gotten closer with a new [man/woman]. [He/She] is very attractive, has a good job, and has treated you better than anyone before. Confident that you will marry soon, the two of you decide to move in together.

Your daughter, who wants you to get back together with her [dad/mom], is not happy about this. She frequently complains whenever your boyfriend comes over and is extremely rude and avoidant. In previous relationships, such behavior has driven others away. However, your partner says [he/she] is committed to this relationship and to establishing a bond with your daughter.

After a month of living together with your partner, things have only gotten worse. Your daughter is almost always upset and has treated you partner worse than others in the past.

One night, she accused your partner of hitting her multiple times. You ask if she has any bruises or injuries, but she says she doesn't. Despite this, she maintains that she was hit and says she doesn't know why there aren't any marks on her.

You bring up the accusation with your partner, and [he/she] denies it, saying that [he/she] would never hurt her and that this is just a way for her to split the two of you up.

### Measureing Beliefs and Actions

Table: (\#tab:outcomes) Outcome variables in each vignette.

+-------------+-------------------+-----------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+
| Variable    | Vignette          | Wording                                                                                                                     | Coding                                                                                          |
+=============+===================+=============================================================================================================================+=================================================================================================+
| *T1 Belief* | Basketball        | How much do you believe the star player is telling the truth about the coach?                                               | 0: Not at all; 100: Completely                                                                  |
+-------------+-------------------+-----------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+
| *T1 Belief* | Brother-in-law    | How much do you believe your daughter is telling the truth about your brother-in-law?                                       | 0: Not at all; 100: Completely                                                                  |
+-------------+-------------------+-----------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+
| *T1 Belief* | Romantic partner  | How much do you believe your daughter is telling the truth about your partner?                                              | 0: Not at all; 100: Completely                                                                  |
+-------------+-------------------+-----------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+
| *T1 Belief* | Thwarted Marriage | How much do you believe your older(bold) daughter is telling the truth about the second man not being a good match for her? | 0: Not at all; 100: Completely                                                                  |
+-------------+-------------------+-----------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+
| *T1 Action* | Basketball        | How likely are you to recommend that the president of the university should suspend the coach?                              | 0: Very unlikely; 100: Very likely                                                              |
+-------------+-------------------+-----------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+
| *T1 Action* | Brother-in-law    | How likely are you to kick your brother-in-law out?                                                                         | 0: Very unlikely; 100: Very likely                                                              |
+-------------+-------------------+-----------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+
| *T1 Action* | Romantic partner  | How likely are you to kick your partner out?                                                                                | 0: Very unlikely; 100: Very likely                                                              |
+-------------+-------------------+-----------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+
| *T1 Action* | Thwarted Marriage | How likely are you to provide the large dowry so your older(bold) daughter can marry the man she is in love with?           | 0: Very unlikely; 100: Very likely                                                              |
+-------------+-------------------+-----------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+
| *T1 Divide* | Thwarted Marriage | How would you plan on dividing the money between your daughters?                                                            | 0: Much more for the younger daughter; 50: Even division; 100: Much more for the older daughter |
+-------------+-------------------+-----------------------------------------------------------------------------------------------------------------------------+-------------------------------------------------------------------------------------------------+

### Signals

##### Verbal request

Your daughter has appeared normal since her claims about your partner. She has still looked fit and put the same amount of effort into her appearance, always wearing clean clothes and spending time on her hair. She has also slept about the same amount of time as normal and remained active throughout the day.

Her demeanor has also remained normal. She has generally appeared calm and has been able to maintain her concentration. She has also done her homework on time and maintained an A in all her classes as she always has in the past.

When the two of you talk about your partner, she remains calm and asks you to kick your partner out.

##### Crying

Your daughter has appeared normal since her claims about your partner. She has still looked fit and put the same amount of effort into her appearance, always wearing clean clothes and spending time on her hair. She has also slept about the same amount of time as normal and remained active throughout the day.

Her demeanor has also remained normal. She has generally appeared calm and has been able to maintain her concentration. She has also done her homework on time and maintained an A in all her classes as she always has in the past.

When the two of you talk about your partner, she starts sobbing and asks you to kick your partner out.

##### Mild-depression (crying with behavioral changes)

Your daughter has appeared sad since her claims about your partner. She has looked a little less fit and has put less effort into her appearance, occasionally wearing dirty clothes and sometimes not doing her hair. She has also sometimes woken up late and has seemed a little less active than normal.

Her demeanor has also changed slightly. She has occasionally gotten distressed and has seemed distracted. She has also missed a few assignments and her grades dropped from straight A's to a B average for the first time in her life.

When the two of you talk about your partner, she starts sobbing and asks you to kick your partner out.

##### Depression

Your daughter started to appear noticeably different since her claims about your partner. She has lost weight and has been putting no effort into her appearance, often wearing dirty clothes and rarely doing her hair. She has also slept more than normal but still seemed tired and inactive.

Her demeanor also changed drastically. She has often gotten distressed and has had difficulty maintaining her concentration. She has also missed multiple assignments, and her grades dropped from straight A's to a C average for the first time in her life.

When the two of you talk about your partner, she starts sobbing and asks you to kick your partner out.

##### Suicide attempt

Your daughter started to appear noticeably different since her claims about your partner. She has lost weight and has been putting no effort into her appearance, often wearing dirty clothes and rarely doing her hair. She has also slept more than normal but still seemed tired and inactive.

Her demeanor also changed drastically. She has often gotten distressed and has had difficulty maintaining her concentration. She has also missed multiple assignments, and her grades dropped from straight A's to a C average for the first time in her life.

When the two of you talk about your partner, she starts sobbing and asks you to kick your partner out. Then, this past week, you and your daughter went through a traumatic experience when your daughter attempted suicide. You rushed her to the hospital, where she was placed in critical but stable condition.

#### Thwarted Marriage

Imagine you are the parent of two daughters, ages 19 and 21, and you have no sons.

Your younger daughter is very sweet. Your older daughter is hard to deal with, however. She is very jealous of her younger sister, and constantly accuses you of favoring her, even when your older daughter receives more.

You live in a community where marriages are arranged and dowries are paid. Recently, you and your partner have been attempting to arrange a marriage for your older daughter with a man she is already in love with. The man, who comes from a good family and is in the same caste as yours, appears to be equally interested in your daughter, and his family seems receptive.

(page break)

As your younger daughter is also not far from marriage, you and your partner decide to divide the money you have saved for dowries in half, with equal amounts for both daughters. At first, an equal split of your savings seemed possible. However, the family of the man your older daughter wants to marry is getting offers of marriage from other families, and it now appears that his family is expecting a much larger dowry.

Since providing a large enough dowry to compete with the other offers would leave little savings for your younger daughter's dowry, you decide to look for another husband for your older daughter.

(page break)

After a few months of searching, you are confident you found a good man for your older daughter. He is very devout, kind, and comes from a good family in the same caste as yours. He is a bit younger than your daughter, and has less money than typical for your caste, but his family appears to be willing to accept a smaller dowry that would leave a fair amount for your younger daughter.

You ask your older daughter what she thinks of the man, and she is immediately disappointed. She complains that she already loves the man you originally intended her to marry. She says the second man is younger than she is and not attractive. You explain that she might not be able to marry the man she loves because it would require a much larger dowry, leaving too little for her sister.

(page break)

You encourage your daughter to spend some time with the second man, and she begrudgingly agrees. After each date, she becomes more insistent that there is no chemistry between them, he is immature, has no sense of humor, and just isn't right for her. She also complains that the family of the man she loves is seeking other marriage partners and that she is afraid he will become interested in his other marriage options.

Your partner is not sure if your daughter really thinks the other man is a poor choice, or if she just wants to get her way as usual.

(page break)

Your family talks to the first man's family again but the amount they want is almost all you have saved. Your partner tells your older daughter that it's just not fair to spend so much on her because saving enough for your younger daughter's dowry would be almost impossible in the next two or three years. For this reason, your family begins to hold more serious discussions with the family of the second man despite your older daughter's objections.

Two days later, your family receives word that the first man's family is finalizing an agreement for him to marry another woman.

#### Signals

##### Verbal request

Your older daughter has appeared normal since your attempts at finding another potential husband for her. She has still looked fit and put the same amount of effort into her appearance, always wearing clean clothes and spending time on her hair. She has also slept about the same amount of time as normal and remained active throughout the day.

Her demeanor has also remained normal. She has generally appeared calm and has been able to maintain her concentration. She has also done her homework on time and maintained an A in all her classes as she always has in the past.

When the two of you talk about the arranged marriage, she remains calm and asks you to pay the amount the family of the man she loves wants.

##### Crying

Your older daughter has appeared normal since your attempts at finding another potential husband for her. She has still looked fit and put the same amount of effort into her appearance, always wearing clean clothes and spending time on her hair. She has also slept about the same amount of time as normal and remained active throughout the day.

Her demeanor has also remained normal. She has generally appeared calm and has been able to maintain her concentration. She has also done her homework on time and maintained an A in all her classes as she always has in the past.

When the two of you talk about the arranged marriage, she starts sobbing and asks you to pay the amount the family of the man she loves wants.

##### Mild-depression (crying with behavioral changes)

Your older daughter has appeared sad since your attempts at finding another potential husband for her. She has looked a little less fit and has put less effort into her appearance, occasionally wearing dirty clothes and sometimes not doing her hair. She has also sometimes woken up late and has seemed a little less active than normal.

Her demeanor has also changed slightly. She has occasionally gotten distressed and has seemed distracted. She has also missed a few assignments and her grades dropped from straight A's to a B average for the first time in her life.

When the two of you talk about the arranged marriage, she starts sobbing and asks you to pay the amount the family of the man she loves wants.

##### Depression {#depression-3}

Your older daughter started to appear noticeably different since your attempts at finding another potential husband for her. She has lost weight and has been putting no effort into her appearance, often wearing dirty clothes and rarely doing her hair. She has also slept more than normal but still seemed tired and inactive.

Her demeanor also changed drastically. She has often gotten distressed and has had difficulty maintaining her concentration. She has also missed multiple assignments, and her grades dropped from straight A's to a C average for the first time in her life.

When the two of you talk about the arranged marriage, she starts sobbing and asks you to pay the amount the family of the man she loves wants.

##### Suicide attempt

Your older daughter started to appear noticeably different since your attempts at finding another potential husband for her. She has lost weight and has been putting no effort into her appearance, often wearing dirty clothes and rarely doing her hair. She has also slept more than normal but still seemed tired and inactive.

Her demeanor also changed drastically. She has often gotten distressed and has had difficulty maintaining her concentration. She has also missed multiple assignments, and her grades dropped from straight A's to a C average for the first time in her life.

When the two of you talk about the arranged marriage, she starts sobbing and asks you to pay the amount the family of the man she loves wants. Then, this past week, you and your daughter went through a traumatic experience when your daughter attempted suicide. You rushed her to the hospital, where she was placed in critical but stable condition.

## Emotion Checklist

How do you think [the star player / your daughter] feels? Click **all** that apply.

* Angry
* Sad 
* Suicidal 
* Mentally ill 
* Depressed 
* Guilty
* Calm
* Neutral
* Scared
* Tired
* Distressed
* Devious
* Jealous
* Confident
* Traumatized
* Violated
* None of the above

## T3 Evidence

### Basketball Coach

Since your last response, it was discovered that a security camera captured the assault happening, just as the star player had described.

### Brother-in-law

Since your last response, your daughter has come to you with a video she secretly took which showed your brother-in-law trying to grope her as she had described previously.

### Romantic partner

Since your last response, your daughter has come to you with a video she secretly took which showed your partner hitting her as she had described previously.

### Thwarted marriage

Just now in the market, you saw the man whose family is willing to accept a smaller dowry (the man your daughter does not want to marry). You heard him insult your daughter and your family.

## Preregistered regression models

We preregistered four OLS linear regression models and one mediation model (<https://osf.io/g3s6n>).

The first two OLS regression models tested the effects of our ordinal signal variable on Time 2 *Belief* and *Action*, controlling for T1 *Belief* and *Action*, respectively (a pre-test/post-test design):

`r extract_eq(models_prereg$Model$m1, intercept='beta', wrap=T)` 
`r extract_eq(models_prereg$Model$m3, intercept='beta', wrap=T)`

We also preregisterd models with an interaction term, for reasons explained in the main text:

`r extract_eq(models_prereg$Model$m1b, intercept='beta', wrap=T)` 
`r extract_eq(models_prereg$Model$m3b, intercept='beta', wrap=T)`

The `glm` models we report in the main text:

`r extract_eq(models$Model$m1, intercept='beta', wrap=T)` 
`r extract_eq(models$Model$m3, intercept='beta', wrap=T)`

The preregistered mediation and outcome models, respectively, were:

`r extract_eq(med_prereg$model.m, intercept='beta', wrap=T)` 
`r extract_eq(med_prereg$model.y, intercept='beta', wrap=T)`

The mediation and outcome models, respectively, that we report in the main text were:

`r extract_eq(med1$model.m, intercept='beta', wrap=T)` 
`r extract_eq(med1$model.y, intercept='beta', wrap=T)`

## Sample size in each condition in the current study

```{r conditions}
as.data.frame.matrix(xtabs(~signal+vignette, d)) %>% 
  kable(caption = "Sample size in each condition.") %>% 
  kable_styling()
```

<!-- ## Distributions of T2 Belief and Action by signal and vignette -->

```{r distbelief, fig.width=10, fig.height=10, fig.cap="Histograms of *T2 Belief* by signal and vignette."}

# h <- hist(d$T2Belief, breaks = "FD", plot = FALSE)
# 
# ggplot(d, aes(x = T2Belief, fill=signal)) +
#   geom_histogram(colour='grey', breaks = h$breaks) +
#   scale_x_continuous(breaks = c(0, 0.5, 1)) +
#   facet_grid(signal~vignette, as.table = F) +
#   guides(fill='none') +
#   xlab('\n T2 Belief') +
#   theme_minimal(15) +
#   theme(strip.text.y = element_text(angle=0, hjust=0))
```

```{r distaction, fig.width=10, fig.height=10, fig.cap="Histograms of *T2 Action* by signal and vignette."}

# h <- hist(d$T2Action, breaks = "FD", plot = FALSE)
# 
# ggplot(d, aes(x = T2Action, fill=signal)) +
#   geom_histogram(colour='grey', breaks=h$breaks) +
#   scale_x_continuous(breaks = c(0, 0.5, 1)) +
#   facet_grid(signal~vignette, as.table = F) +
#   guides(fill='none') +  
#   xlab('\n T2 Action') +
#   theme_minimal(15) + 
#   theme(strip.text.y = element_text(angle=0, hjust=0))
```

## Socioeconomic distribution

```{r socioeconomic, fig.width=8, fig.cap="Distribution of participants by age, log annual income (USD), and nationality. Each dot is one participant. A small number of participants who did not report annual income, or reported values < $10 USD, were removed from this plot. A small amount of jitter was added to reveal overlapping points."}

us_median_hh <- 68703 # census.gov

# IHDS2
# Household Income Inequality and Income Mobility: Implications Towards Equalizing Longer-Term Incomes in India. By: Mishra, Aswini Kumar, International Economic Journal, 10168737, Jun2018, Vol. 32, Issue 2
#
# Table 1: per capita household income
# 	Mean	Median	Gini coefficient
# 2011–2012			
# Urban	36,101.4	22,595.0	49.0
# Rural	18,158.0	11,085.3	51.5
# Total	23,831.4	13,860.0	53.1

# 6.4 persons per household

# 13860 * 6.4 = 88704

# According to some website, in 2021 this would be
# 154,115.8 rupees (73.74% inflation), which is $2,087
india_median_hh <- 2087

clrs <- viridisLite::magma(11)[c(4, 8)]

d %>% 
  dplyr::filter(!is.na(Age), !is.na(Income), Income >= 10) %>% 
  mutate(
    Nationality = ifelse(sample == 'American', 'US', 'India'),
    logIncome = log10(Income)
  ) %>% 
  ggplot(aes(Age, logIncome, colour = Nationality)) +
  geom_jitter(width = 0.5, alpha = 0.35) +
  geom_hline(yintercept = log10(india_median_hh), linetype = 'dotted', colour = clrs[1]) +
  geom_hline(yintercept = log10(us_median_hh), linetype = 'dotted', colour = clrs[2]) +
  annotate('text', 85, log10(us_median_hh), label = 'Median US\nincome', hjust = 0) +
  annotate('text', 85, log10(india_median_hh), label = 'Median India\nincome', hjust = 0) +
  scale_color_binary() +
  guides(colour = guide_legend(reverse = T, override.aes = list(alpha=1, size = 3))) +
  coord_cartesian(clip = 'off') +
  labs(x = '\nAge (years)', y = 'Log annual\nincome (USD)') +
  theme_minimal(15) +
  theme(
    plot.margin = unit(c(1,5,1,1), "lines"),
    axis.title.y = element_text(angle = 0, hjust=0), 
    legend.position = 'top'
    )

```

## Signal effect sizes

```{r effectsizes, fig.width=10, fig.height=7, fig.cap="Between-subjects Cohen's d (95% CI) for the effect of each signal on Belief and Action in each vignette relative to the Verbal request control condtion, ranked by effect size."}

relbl <- function(lb){
  map_chr(lb, ~str_split(.x, ',')[[1]][2])
}

ggplot(effect_sizes, aes(est, fct_reorder(id, est), xmin=low, xmax=high, colour = signal, shape=Outcome)) +
  geom_pointrange(position = position_dodge(width = 0.7), lwd = 2.5, fatten = 1, alpha=0.5) +
  geom_vline(xintercept = 0, linetype = 'dotted') +
  scale_y_discrete(labels = relbl) +
  scale_color_manual(name=NULL, values = c('darkblue', 'darkred', 'red', 'orange')) +
  # scale_color_viridis_d(option='A') +
  guides(colour = guide_legend(override.aes = list(size = 1)), shape = guide_legend(title=NULL, override.aes = list(size = 1))) +
  # facet_wrap(~Outcome, ncol=1) +
  labs(x="\nCohen's d (95% CI)", y=NULL) +
  theme_minimal(15) +
  theme(legend.position = 'top')
```

```{r effectsizes2, fig.width=10, fig.height=7, fig.cap="Within-subjects Cohen's d (95% CI) for the effect of each signal on *T2 Belief* and *T2 Action* in each vignette relative to baseline *T1 Belief* and *T1 Action*, respectively, ranked by effect size."}

ggplot(effect_sizes2, aes(effsize, fct_reorder(id, effsize), xmin=conf.low, xmax=conf.high, colour = signal, shape=Outcome)) +
  geom_pointrange(position = position_dodge(width = 0.7), lwd = 2.5, fatten = 1, alpha=0.5) +
  geom_vline(xintercept = 0, linetype = 'dotted') +
  scale_y_discrete(labels = relbl) +
  scale_color_manual(name=NULL, values = c('black', 'darkblue', 'darkred', 'red', 'orange')) +
  # scale_color_viridis_d(option='A') +
  guides(colour = guide_legend(override.aes = list(size = 1)), shape = guide_legend(title=NULL, override.aes = list(size = 1))) +
  # facet_wrap(~Outcome, ncol=1) +
  labs(x="\nCohen's d (95% CI)", y=NULL) +
  theme_minimal(15) +
  theme(legend.position = 'top')
```

<!-- ## Sociodemographic associations -->

```{r demoeffectsaction, fig.width=9, fig.height=8, fig.cap="Demographic effects on Action. **Top**: Association of age with *T2 Action*, controlling for *T1 Action*, sex, and vignette. **Bottom**: Association of sex with *T2 Action* controlling for *T1 Action*, age, and vignette. Estimated using a generalized linear regression model with the quasibinomial family."}
# effectsPlots$m25plotAge / effectsPlots$m25plotSex
```

```{r demovignettebelief, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with *T2 Belief* by vignette, controlling for *T1 Belief* and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
# (effectsPlots_US$m18plot / effectsPlots_US$m13plot)
```

```{r demovignetteaction, fig.width=8, fig.height=8, fig.cap="Associations of age (top) and sex (bottom) with *T2 Action* by vignette, controlling for *T1 Action* and signal. Estimated using a generalized linear regression model with the quasibinomial family."}
# effectsPlots_US$m19plot / effectsPlots_US$m15plot
```

## Mediation of Suicide attempt relative to Depression signals

The total effect of the *Suicide attempt* signal (treatment) vs. *Depression* (control) on the likelihood of *Action* was to increase it by `r 100*signif(med2$tau.coef, 3)` points from T1 to T2 (on the original 100-point scale). Of this increase, `r 100*signif(med2$n.avg, 2)`% was mediated by the increased *Belief* that the victim was telling the truth. See Figure \@ref(fig:mediationsuicide).

```{r mediationsuicide, fig.height=4, fig.cap="The effect of a Suicide attempt (treatment) vs. Depression (control) on Action, [0-1 scale], is only partially mediated by Belief. ACME: Average causal mediation effect. ADE: Average direct effect. The mediation model controlled for *T1 Belief*, and the outcome model controlled for *T1 Belief* and *T1 Action*. Both models were GLMs with the binomial family (the mediation package does not support the quasibinomial family)."}
ggmediation(med2)
```

## Change in beliefs and actions from T1 to T2

```{r rawbeliefs, fig.width=10, fig.height=10, fig.cap="Change in beliefs from T1 to T2 for participants in each condition. Each grey line is one participant. Red lines are the mean change."}
# plot_raw_beliefs
```

```{r rawactions, fig.width=10, fig.height=10, fig.cap="Change in actions from T1 to T2 for participants in each condition. Each grey line is one participant. Red lines are the mean change."}
# plot_raw_actions
```

```{r rawbeliefactions2d, fig.width=10, fig.height=10, fig.cap="Change in beliefs and actions from T1 (arrow base) to T2 arrow head) for participants in each condition. Each grey arrow is one participant. Red arrows are the mean change."}
plot_raw_belief_action2d
```

## Emotions in vignette (T1) and signal (T2) conditions

```{r emotionheatmaps, fig.width=10, fig.cap="Each cell represents the proportion of participants who inferred that emotion. **Time 1 emotions**: inference of victim's emotional state in each vignette at time 1. **Time 2 emotions**: inference of victim's emotional state following the signal at time 2. Participants could check multiple emotions from the list. Rows ordered by mean proportion across vignettes."}

row_order = rowSums(dfT1emotion_vignette[-1])

dfT1emotion_vignette$name <- factor(dfT1emotion_vignette$name, levels = dfT1emotion_vignette$name[order(row_order)])

values <- seq(0, 1, 0.25)

p1 <- 
  dfT1emotion_vignette %>% 
  gather(key = key, value = value, -1) %>%
  mutate(key = factor(key, levels = c('Thwarted marriage', 'Basketball coach', 'Romantic partner', 'Brother-in-law'))) %>%
  ggplot(aes_string("key", "name", fill = "value")) +
  geom_tile() + 
  scale_fill_viridis_c('Proportion', option = 'B', limits = c(0, 1), breaks = values) +
  scale_x_discrete(labels = scales::label_wrap(10)) +
  labs(title = 'T1 emotions', x = "", y = "") + 
  theme_minimal()

dfT2emotion_signal$name <- factor(dfT2emotion_signal$name, levels = dfT1emotion_vignette$name[order(row_order)])

p2 <- 
  dfT2emotion_signal %>% 
  gather(key = key, value = value, -1) %>%
  mutate(key = factor(key, levels = c('Verbal request', 'Crying', 'Mild depression', 'Depression', 'Suicide attempt'))) %>%
  ggplot(aes_string("key", "name", fill = "value")) +
  geom_tile() + 
  scale_fill_viridis_c('Proportion', option = 'B', limits = c(0, 1), breaks = values) +
  scale_x_discrete(labels = scales::label_wrap(10)) +
  labs(title = 'T2 emotions', x = "", y = "") + 
  theme_minimal() +
  theme(axis.text.y = element_blank())

p1 + p2 + plot_layout(guides='collect')
```


```{r t1emotionheatmap, fig.width=8, fig.cap="Time 1 inference of victim's emotional state in each vignette. Each cell represents the proportion of participants who inferred that emotion. Participants could check multiple emotions from the list. Rows and columns ordered with the PCA angle method in the seriation package."}
# hagenheat(dfT1emotion_vignette, rotate_labels = F, seriation_method = 'PCA_angle', viridis_option = 'B') + theme_minimal(15)
```

<!-- ## T2 emotion vs signal -->

```{r t2emotionheatmap, fig.width=8, fig.cap="Time 2 inference of victim's emotional state following the signal. Each cell represents the proportion of participants who inferred that emotion. Participants could check multiple emotions from the list. Rows and columns ordered with the PCA angle method in the seriation package."}

# hagenheat(dfT2emotion_signal, rotate_labels = F, seriation_method = 'PCA_angle', viridis_option = 'B') + theme_minimal(15)
```

## Change in emotions T1 to T2 by signal

```{r diffemotions, fig.width=8, fig.cap="Change in inference of victim's emotional state From pre- to post-signal. Each cell represents the difference in proportions of participants who inferred that emotion (T2 - T1). Participants could check multiple emotions from the list. Rows and columns ordered with the PCA angle method in the seriation package."}

d_diff <- as.matrix(dfT2emotion_signal[-1]) - as.matrix(dfT1emotion_signal[-1])
rownames(d_diff) <- dfT2emotion_signal$name

hagenheat(d_diff, rotate_labels = F, seriation_method = 'PCA_angle', rev_col = T) + scale_fill_gradient2(low='blue', high='red')

```

## Mentally ill

```{r mentallyillT1, fig.width=8, fig.cap="Proportion of participants in each vignette who thought that the victim was mentally ill at T1. Means estimated with a generalized linear regression model with the quasibinomial family."}
effectsPlots$m27plot + ggtitle('Time 1')
```

```{r mentallyillT2, fig.width=8, fig.cap="Proportion of participants in each condition who thought that the victim was mentally ill at T2, controlling for perceptions of mental illness at T1. Means estimated with a generalized linear regression model with the quasibinomial family."}
effectsPlots$m21plot + ggtitle('Time 2')
```

## The effect of signals on T2 division in thwarted marriage vignette

```{r division, fig.width=9, fig.cap="Effect of signals on division of dowry between the daughters in the thwarted marriage vignette, with an interaction between T1 Division and signal. Means estimated with a generalized linear regression model with the quasibinomial family."}
effectsPlots_India$m7plot
```

```{r divisionage, fig.width=9, fig.cap="Interaction of age with signals on *T2 Division* of dowry between the daughters in the thwarted marriage vignette."}
effectsPlots_India$m8plot
```

## Validity check: T3 Action vs. T2 Action

```{r t3action, fig.width=10, fig.cap="The effect of post-signal (T3) information that the victim was telling the truth on the likelihood of acting to help her (*T3 Action*), controlling for *T2 Action*, with a term for the interaction of *T2 Action* and vignette."}
effectsPlots$m26plot + theme(panel.spacing = unit(1.5, "lines"))
```

## Feedback distribution by vignette

At the end of the study, participants could optionally provide written feedback about study. We classified all feedback into two categories: *Generic* for a few short words, such as 'nice study', or 'thanks'; and *Informative* for one or two sentences commenting on specific aspects of the survey. See Figure \@ref(fig:feedbackplot). 

```{r feedbackplot, fig.width=9, fig.cap="The distribution of optional feedback provided at the end of the study. None: no feedback provided (includes, e.g., 'none', 'no', 'N/A'). Generic: A few short words, such as 'nice study', or 'thanks'. Informative: one or two sentences commenting on specific aspects of the survey. Numbers are the number of participants in that category."}

hagenutils::ggxtabs(xtabs(~Feedback3 + vignette, d), cell_counts = T) + xlab('')
```

We then further classified the *Informative* feedback into six categories: the vignette corresponds to real-life experiences; comments on the vignette (e.g., other ways to respond); what the participant thought about the T3 information that the victim was telling the truth; criticisms of the study (e.g., the attention checks); comments on the study; and an explanation of why the participant acted or failed to act. See Figure \@ref(fig:informativeplot). The full texts of all comments are in Table \@ref(tab:informativetab).

```{r informativeplot, fig.width=9, fig.cap="The distribution of informative feedback provided at the end of the study. See Figure \\@ref(fig:feedbackplot) for other types of feedback. See Table \\@ref(tab:informativetab) for the text of all informative statements. Numbers are the number of participants in that category."}
hagenutils::ggxtabs(xtabs(~Informative + vignette, d), cell_counts = T) + xlab('')
```

```{r informativetab}
d %>% 
  dplyr::select(Informative, vignette, Feedback_original) %>% 
  dplyr::filter(!is.na(Informative)) %>% 
  arrange(Informative, vignette) %>% 
  kable(caption="Informative feedback provided by participants after T3 (the end of the study). Categorization required some judgement calls.") %>% 
  kable_styling()
```

## Regression tables

```{r regressiontables}
tmp <- models %>% 
  dplyr::filter(Name %in% c('m1', 'm2', 'm3', 'm4', 'm21', 'm24', 'm25'))

regressiontable(tmp$Model, caption = 'Regression coefficients')
```

## ANOVA tables

ANOVA tables for generalized linear regression models (quasibinomial family).

```{r anovatables}

bind_rows(tmp$TidyANOVA, .id = 'Model') %>%
  dplyr::rename(Variable = term, `LR Chisq` = statistic) %>% 
  gt::gt(groupname_col = "Model", caption = "ANOVA tables") %>% 
  gt::cols_label(Variable = "") %>% 
  gt::fmt_number(columns=3, decimals = 1) %>% 
  gt::fmt_scientific(5, decimals = 1) %>% 
  gt::tab_style(style = gt::cell_text(indent = gt::px(40)), 
                locations = gt::cells_body(columns = "Variable"))
```

## Preregistered OLS models

```{r preregbelief, fig.width=8, fig.height=8, fig.cap="The effect of the signals on *T2 Belief* (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively. Estimated using OLS linear regression models (gaussian)."}
p1 <- effect_plot(fit=models_prereg$Model$m1, xvar='signal',ylbl='T2 Belief', data=d)

p2 <- effect_plot(fit=models_prereg$Model$m3, xvar='signal', ylbl='T2 Action', data=d)

(p1 + theme(axis.text.x = element_blank()))/p2 + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

```{r preregbeliefB, fig.width=8, fig.height=8, fig.cap="The effect of the signals on *T2 Belief* (top) and Action (bottom) controlling for T1 values of Belief and Action, respectively, with an interaction between T1 values and the signal. Estimated using OLS linear regression models (gaussian)."}
p1 <- effect_plot(fit=models_prereg$Model$m1b, xvar='signal', by='T1Belief', ylbl='T2 Belief', data=d)

p2 <- effect_plot(fit=models_prereg$Model$m3b, xvar='signal', by='T1Action', ylbl='T2 Action', data=d)

(p1 + theme(axis.text.x = element_blank()))/p2 + plot_layout(ncol = 1) & theme(plot.margin = unit(c(0,0,0,0), units = 'mm'))
```

### OLS regression tables

```{r preregtables}
regressiontable(models_prereg$Model, caption = 'OLS regression coefficients for preregistered linear regression models.')
```

### OLS Anova tables

```{r prepreganova}
bind_rows(models_prereg$TidyANOVA, .id = 'Model') %>%
  dplyr::rename(Variable = term, `LR Chisq` = statistic) %>% 
  gt::gt(groupname_col = "Model", caption = "OLS ANOVA tables") %>% 
  gt::cols_label(Variable = "") %>% 
  gt::fmt_number(columns=c(3,5), decimals = 1) %>% 
  gt::fmt_scientific(6, decimals = 1) %>% 
  gt::fmt_missing(columns=5:6, rows = everything(), missing_text = "---") %>% 
  gt::tab_style(style = gt::cell_text(indent = gt::px(40)), 
                locations = gt::cells_body(columns = "Variable"))

```

## Comparison of estimates and confidence intervals of GLM, Fractional, and bootstrapped models

```{r}
m1 <- models$Model[['m1']]
m1_robust <- lmtest::coeftest(m1, vcov = sandwich::vcovHC(m1, type="HC"))

m1_quasi <- models$TidyModel$m1
m1_diff <- (m1_boot[,'boot 97.5%'] - m1_boot[,'boot 2.5%'])/(m1_quasi$conf.high - m1_quasi$conf.low)

m3 <- models$Model[['m3']]
m3_robust <- lmtest::coeftest(m3, vcov = sandwich::vcovHC(m3, type="HC"))

m3_quasi <- models$TidyModel$m3
m3_diff <- (m3_boot[,'boot 97.5%'] - m3_boot[,'boot 2.5%'])/(m3_quasi$conf.high - m3_quasi$conf.low)

diff_sum <- summary(c(m1_diff, m3_diff))
```

```{r modelcompareplots, fig.width=11, fig.cap="Comparison of the coefficients and 95% CIs of the quasibinomial glm, fractional, and bootstrapped logistic regression models of *T2 Belief* and *T2 Action* (for details, see the Statistical methods section)."}

belief_compare_plot + action_compare_plot + plot_layout(guides = 'collect')

```

# References
